### 2021-07-14 paper/referee_response.tex
initial release

\documentclass[letterpaper, 12pt]{article}

% Imports
\usepackage{amsmath, amssymb}
\usepackage{mathtools}
\usepackage{graphicx}

% Layout
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\setlength{\parindent}{0em}
\setlength{\parskip}{1em}

% Body font
\usepackage{soul, color}
\usepackage[colorlinks, allcolors=blue]{hyperref}
\usepackage{newtxtext}

% Reviews
\usepackage{framed}
\newenvironment{review}{\vspace{1em}\begin{leftbar}}{\end{leftbar}}

\begin{document}

Dear Professor Tobias,

We thank the referees for constructive reports. We have made all of the changes they suggested. In addition, we added section 3c detailing an asymptotic confirmation of the 3D to 2D transition reported numerically in section 3a to at least three significant figures. If this section is too much of an addition at this stage, we can relegate it to an appendix or remove it.

Please see our responses in line below.

\begin{review}
Comments to the Author(s)
This is an interesting manuscript that reveals a new and important result regarding the magnetorotational instability (MRI) in a radially confined geometry such as a Taylor-Couette experiment between differentially rotating cylinders. The authors show that 3D (non-axisymmetric) modes can be preferred to 2D (axisymmetric) ones, especially near the onset of the 2D instability. This might be important for the outcome of MRI experiments such as the one at Princeton, and it also has ramifications for dynamo action. The authors obtain good agreement, where expected between their numerical calculations of unstable modes and their asymptotic analysis.

The manuscript is mostly well written and the figures are well presented (although some are too small in the current format - in particular, Fig. 3 is hard to read).

I recommend publication in Proceedings A provided the following matters can be addressed.

1. It would be helpful to explain that "three-dimensional" means "non-axisymmetric" and "two-dimensional" means "axisymmetric" (but with three-component velocity and magnetic fields). (Of course, "axisymmetric" is to be interpreted as "y-independent" in this local model.)
\end{review}
We have fixed this.

\begin{review}  
2. Section 1: Perhaps clarify that the "finite-channel cutoff" is due to the finite width, not height.
\end{review}
Fixed.

\begin{review}
3. Section 1: What does "weak-shear" mean here: weak compared to what?  
\end{review}

Fixed

\begin{review}
4. Section 2: The notation could be defined more clearly by saying what v, b and j represent. We certainly need to know that f = 2 Omega.
\end{review}

We are unclear what the referee refers to here; these are already in the text.

\begin{review}
5. Section 2: The channel is periodic in z, but it seems that the periodicity length is unspecified, so that any value of the wavenumber $k_z$ is permissible. Please clarify this.
\end{review}

Fixed.

\begin{review}
6. Section 2: Personally, I find it confusing to have the symbol q denoting the Rossby number when (as said in the footnote) in most of the MRI literature q denotes twice the Rossby number. Would it not be better to use a different symbol for the Rossby number?
\end{review}

Fixed

\begin{review}
7. Section 2: I believe that q > 1 (rather than q = 1, as stated) corresponds to Rayleigh-unstable shear. (Are we to assume throughout that q > 0, i.e. S < 0?)
\end{review}
Fixed

\begin{review}
8. Around eq. 2.6: Please clarify the meaning of "in two dimensions" (i.e. axisymmetric, or y-independent, perturbations). Is there a convenient reference for eq. 2.6, as the result is not obvious? I believe it comes from taking the limit $k_z d \ll 1$, so the periodicity length in z has to be very long.
\end{review}

Fixed.

\begin{review}
9. Section 2: An important point is that the system of units has not been clarified and the relevant dimensionless parameters have not all been identified. Everything is expressed in dimensional terms until we are told that nu = eta = 1e-5. In what units are these values expressed?
  \end{review}

We have clarified this point while simultaneously addressing point 16.

\begin{review}
10. Section 2: I thought initially that there was a missing, and unspecified, dimensionless parameter such as $B_0 / f d$, which is a measure of the field strength. I see now that this can be determined from the given values of $q$ and $S/S_c$, but it might be better to explain this in the text.
\end{review}

We have done so.

\begin{review}
11. Figure 1 caption: Presumably "cray" should be "grey".  
\end{review}

Fixed

\begin{review}
12. Before eq. 3.1: Presumably "in when" should be "when".  
\end{review}

Fixed

\begin{review}
13. Before eq. 3.9: "2nd order" should probably be "2nd order in x".  
\end{review}

Fixed

\begin{review}
14. Eq. 3.9: Why are $k_z$ and $k_x$ renamed $k$ and $l$ here?  
\end{review}

We have corrected this.

\begin{review}
15. Section 3(c): It seems to me that the prevalence of axisymmetric modes in previous numerical simulations may have more to do with the radial boundary conditions than with the degree of supercriticality.
\end{review}

A note was added to this effect.

\begin{review}
16. Section 3(e): It would have been better to define Re and Rm back in Section 2. Again, the units in which eta and nu are given are not specified. There are a couple of typos in the first sentence: "as series" should be "a series", and "with the" should be "with".
\end{review}

Typos fixed and Reynolds numbers moved upwards.

\begin{review}
17. I disagree that there is only one previous study of non-turbulent MRI dynamos; indeed, reference [15] and related papers provide examples of this.  
\end{review}

This is completely correct; our original paper was in error. We meant to express that there are no other \emph{linear} MRI dynamos, not \emph{laminar} ones!

\begin{review}
Referee: 2

Comments to the Author(s)
Thank you very much for sending your article! I find the subject, three-dimensional (3d) modes of the magneto-rotational instability (MRI), a very interesting aspect that so far has received much less attention than the corresponding two-dimensional (2d) modes. Your analysis is very well suited to the task, the results are interesting and presented clearly and in detail and set into the context of conditions in stars and lab experiments . Hence I recommend the article for publication. I have, however, a few minor points to
raise. The primary one is a question regarding the range of applicability of the 3d results, in particular in stars. You vary the ratio between the shear, S, and the critical shear value, $S_c$, given in equation (2.6), from values below unity to a maximum of 4. You note that for $S/S_c >~ 2.05$, the linear dynamics should be dominated by 2d
modes. $S_c$ involves, among other quantities, the square of the magnetic field strength, which can vary by (or is uncertain to) quite large factors. Thus, even moderate variations of the field may change $S_c$ such that $S/S_c$ no longer favours 3d modes. My question is, thus, if this apparent need for what might be considered fine-tuning limits the applicability of the results to, e.g., stars.
\end{review}
Thank you for pointing this out. We have added a cautionary note regarding this point.

\begin{review}
Apart from that, I have only small points:
- Section 3(a): the text mentions $S/S_c = 1.02$, but Fig. 1 has 1.002 and 1.2. Also, at the end of the paragraph: "does not intercept the x-axis"; one might put "$k_z$-axis" to avoid confusion with the axes of the simulation box.
- Section 3(f), after the equation: "rms is denotes...": remove one of the two verbs.
- Also 3(f): does the integral of the EMF over x vanish for reasons of symmetry?
\end{review}

We have fixed all of these.

\vspace{1em}

Sincerely, \\
Jeffrey S. Oishi, Geoffrey M. Vasil, Morgan Baxter, Andrew Swan, Keaton J. Burns, Daniel Lecoanet, \& Benjamin P. Brown
\end{document}

%%% Local Variables:
%%% ispell-local-dictionary: "british"
%%% End:

%  LocalWords:  Tobias magnetorotational Couette axisymmetric Rossby
%  LocalWords:  wavenumber eq nd supercriticality rms Oishi Vasil
%  LocalWords:  Lecoanet

### 2021-07-14 paper/paper.tex
initial release

\documentclass{rsproca_new}%%%%where rsproca is the template name
\newcommand{\SSC}{S/S_{c}}
\newcommand{\Prm}{\mbox{\textit{Pm}}}
\newcommand{\mnras}{Monthly Notices of the Royal Astronomical Society}
\newcommand{\apj}{The Astrophysical Journal}
\newcommand{\apjs}{The Astrophysical Journal Supplement Series}
\newcommand{\apjl}{Astrophysical Journal Letters}
\newcommand{\nat}{Nature}
\newcommand{\prl}{Physical Review Letters}

\newcommand\Beq{\begin{eqnarray}} 
\newcommand\Eeq{\end{eqnarray}}

\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Eqs}[2]{Eqs.~(\ref{#1})--(\ref{#2})}
\newcommand{\eq}[1]{eq.~(\ref{#1})}
\newcommand{\eqs}[2]{eqs.~(\ref{#1})--(\ref{#2})}

\newcommand{\Prob}[2]{\mathbb{P}[\, #1 \, | \, #2 \, ] }
\newcommand{\Pro}[1]{\mathbb{P}[\, #1 \, ] }

\newcommand{\inv}[1]{\frac{1}{#1}}
\newcommand{\pd}[1]{\partial_{#1}}

\newcommand{\Reyn}{\mathrm{Re}}
\newcommand{\Reym}{\mathrm{Rm}}
\newcommand{\Ra}{R}
\newcommand{\Ro}{\mathrm{Ro}}
\newcommand{\Ve}{\mbox{\textit{Ve}}}
\newcommand{\eps}{\varepsilon}
\newcommand{\del}{\delta}
\newcommand{\e}[1]{\eps^{#1}}
\newcommand{\f}{2\Omega}

\newcommand{\n}{\hat{n}}

\newcommand{\s}{\text{s}}
\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}

\newcommand{\dat}{\cdot}
\newcommand{\dd}[1]{\,\mathrm{d}{#1}}
\newcommand{\cc}{\mathrm{c.c.}}
\renewcommand{\vec}[1]{\boldsymbol{#1}}
\newcommand{\grad}{\nabla}
\renewcommand{\div}{\nabla \dat}
\newcommand{\Tr}{\text{Tr}}

\newcommand{\pren}[1]{ \left(#1\right) }
\newcommand{\brac}[1]{\left[\,#1\,\right]}

\newcommand{\ave}[1]{\big<#1\big>}

\newcommand{\txt}[1]{\quad \text{#1} \quad }

\usepackage{graphicx}
\usepackage{amsmath}
%\usepackage{natbib}

\begin{document}

\title{The Magnetorotational Instability Prefers Three Dimensions}

\author{%%%% Author details
Jeffrey~S.~Oishi$^{1}$, Geoffrey~M.~Vasil$^{2}$, Morgan Baxter$^{1}$,
Andrew Swan$^{3}$, Keaton~J.~Burns$^{4,5}$, Daniel~Lecoanet$^6$, Benjamin~P.~Brown$^7$}

%%%%%%%%% Insert author address here
\address{$^{1}$Bates College, Lewiston, ME 04240, USA\\
$^{2}$University of Sydney, Sydney, NSW, Australia\\
$^{3}$University of Cambridge, Cambridge,
UK\\
$^{4}$Flatiron Institute, New York, NY 10010, USA\\
$^{5}$MIT, Cambridge, MA 02139, USA\\
$^{6}$Princeton University, Princeton, NJ 08544, USA\\
$^{7}$University of Colorado, CO 80309, USA}
\corres{Jeffrey~S.~Oishi\\
\email{joishi@bates.edu}}

\begin{abstract}
The magnetorotational instability (MRI) occurs when a weak magnetic field destabilises a rotating, electrically conducting fluid with inwardly increasing angular velocity.
The MRI is essential to astrophysical disk theory where the shear is typically Keplerian.
Internal shear layers in stars may also be MRI unstable, and they take a wide range of profiles, including near-critical.
We show that the fastest growing modes of an ideal magnetofluid are three-dimensional provided the shear rate, $S$, is near the two-dimensional onset value, $S_c$.
For a Keplerian shear, three-dimensional modes are unstable above $S\approx0.10S_c$, and dominate the two-dimensional modes until $S\approx2.05S_{c}$.
These three-dimensional modes dominate for shear profiles relevant to stars and at magnetic Prandtl numbers relevant to liquid-metal laboratory experiments.
Significant numbers of rapidly growing three-dimensional modes remain well past $2.05S_{c}$. 
These finding are significant in three ways. 
First, weakly nonlinear theory suggests that the MRI saturates by pushing the shear rate to its critical value. 
This can happen for systems, like stars and laboratory experiments, that can rearrange their angular velocity profiles.
Second, the non-normal character and large transient growth of MRI modes should be important whenever three-dimensionality exists.
Finally, three-dimensional growth suggests direct dynamo action driven from the linear instability.
\end{abstract}

\maketitle

\section{Introduction}
The magnetorotational instability (MRI) is extremely important in astrophysical fluid dynamics.
A weak magnetic field catalyses turbulence in a Keplerian shear by changing the stability criterion for differentially rotating flows from a negative angular \emph{momentum} gradient to a negative angular \emph{velocity} gradient \cite{1998RvMP...70....1B,2010RSPTA.368.1607J}.
This discovery explained the ubiquitous accretion onto compact objects at rates compatible with observations, and may also influence the formation of planets \cite{2007Natur.448.1022J}.
In disks, the gravitational field dominates the local plasma dynamics, and thus the MRI cannot significantly affect the background shear; it must saturate by other means \cite{2018MNRAS.474.3451X}.
However, stars and liquid metal Taylor-Couette experiments have differential rotation profiles driven by much weaker stresses.
Where the MRI is active in these flows, it saturates by pushing the background shear close to critical \cite{2015RSPSA.47140699V,2017ApJ...841....1C,2017ApJ...841....2C}, analogous to convection mixing entropy.
Stellar interiors have extremely high fluid and magnetic Reynolds numbers, but can operate at or near the critical shear rate for the MRI.
This finite critical shear results from a finite-channel cutoff in the radial direction.
Despite the extensive literature on accretion disks (strong shear, low dissipation) and liquid-metal experiments (weak shear, large dissipation), even the linear MRI is not well understood in the weak-shear (e.g.\ $S \gtrsim S_c$), low-dissipation regime.

Here, we investigate the stability of three-dimensional perturbations near the two-dimensional critical shear rate $S_{c}$ for a nearly inviscid, ideal MHD flow.
Throughout this paper, we use ``three-dimensional'' to refer to non-axisymmetric perturbations (and their local Cartesian equivalents), while ``two-dimensional'' refers to axisymmetric perturbations.
In both cases, we retain all three components of velocity and magnetic fields.
We find that the first destabilised modes are three-dimensional, and thus could act as a dynamo even in the absence of secondary instability.
These results also suggest that the non-normal growth of the MRI is always important, even when axisymmetric modes dominate.

\section{Methods}
\label{sec:methods}

We numerically solve the linearized magnetohydrodynamic equations in rotating plane Couette geometry.
This corresponds to a Cartesian frame rotating with angular frequency $\Omega$ and a linear background shear, $V_{y}=Sx$ \cite{2015RSPSA.47140699V}; it is also the narrow-gap limit of the Taylor-Couette geometry.
We cast the Navier-Stokes equation in the form,
\begin{equation}\label{eq:mhd}
\frac{D \boldsymbol{v}}{Dt}+f \boldsymbol{\hat{z}}\times\boldsymbol{v}+{S}v_{x}\,\boldsymbol{\hat{y}}+\boldsymbol{\nabla}{p}+\nu\,\boldsymbol{\nabla}\times\boldsymbol{\omega}=B_{0}\partial_{z}\boldsymbol{b},
\end{equation}
where
\begin{equation}
\boldsymbol{\omega}=\boldsymbol{\nabla}\times\boldsymbol{v},\quad\text{and}\quad\frac{D}{Dt}=\partial_{t}+{S}x\,\partial_{y}\end{equation}

We write the induction equation in terms of the $x$-component of the magnetic field,
\begin{equation}\label{eq:Bx}
\frac{Db_{x}}{Dt}+\eta(\partial_{y}j_{z}-\partial_{z}j_{y})=B_{0}\partial_{z}v_{x},
\end{equation}
and the $x$-component of the current density ($j_{x}=\partial_{y}b_{z}-\partial_{z}b_{y}$),
\begin{equation}\label{eq:Jx}
\frac{Dj_{x}}{Dt}-\eta\nabla^{2}j_{x}=B_{0}\partial_{z}\omega_{x}-S\partial_{z}b_{x}.
\end{equation}
We explicitly enforce divergence-free velocity and magnetic field,
\begin{equation}\label{eq:divu}
 \boldsymbol{\nabla}\cdot\boldsymbol{v}=\boldsymbol{\nabla}\cdot\boldsymbol{b}=0.
\end{equation}
The spatial domain is a doubly periodic channel in $y,\,z$ with width $-d/2\le{x}\le d/2$.
The boundary conditions are impenetrable stress-free and perfectly conducting; $v_{x}=\omega_{y}=\omega_{z}=b_{x}=\partial_{x}j_{x}=0$ at $x=\pm{d/2}$.

The main input parameters are the Coriolis parameter, $f=2 \Omega$; the background shear rate, $S=dV_{y}/dx<0$; and the vertical magnetic field $B_{0}$ (in Alfv\'{e}n units $\mu_{0}\rho_{0}=1$).
Accretion-disk modelling usually considers the Rossby number $\Ro=-S/f<1$. 
The regime $\Ro\ge1$ corresponds to purely hydrodynamical Rayleigh unstable shear.
Unless otherwise stated $\Ro=3/4$ (Keplerian).
The solution also depends on the viscosity $\nu$ and resistivity $\eta$; in our non-dimensionalization, these are equivalent to the inverse Reynolds and Magnetic Reynolds numbers, respectively.
That is, $\Reyn = 1/\nu$ and $\Reym = 1/\eta$.

The MRI is a weak-field instability; in the inviscid, ideal case the critical shear rate for axisymmetric instability (i.e.\ in \emph{two dimensions}) is
\begin{equation}\label{eq:Sc}
  S_{c}=-\frac{\pi^{2}B_{0}^2}{fd^2}
\end{equation}
\cite{2015RSPSA.47140699V}.

We use $\SSC$ as our instability control parameter.
Because $\SSC$ serves as a ratio of the dimensionless magnetic field strength $B_0^2/f d$ to the length scale $d$, this serves to specify the background field strength.

We assume harmonic perturbations in $y$ and $z$, (e.g. pressure) $p=\hat{p}(x)e^{i(k_{y}y+k_{z}z)+\sigma{t}}$. 
We use a complex-valued growth rate $\sigma=\gamma+i\omega$ with $\gamma,\,\omega$ both real. 
The system reduces to a $\sigma$-eigenvalue problem of 10 first-order ODEs in $x$ with Dirichlet boundary conditions.
We pose and solve equations~(\ref{eq:mhd}-\ref{eq:divu}) using the \emph{Dedalus} framework \cite{2019arXiv190510388B}.

Our main interest is in ideal ($\eta=0$), inviscid ($\nu=0$) conditions.
However, we set $\eta=\nu=10^{-5}$ to avoid critical layers in the \textit{stable} solution branch.
We confirmed our results for unstable solutions are insensitive to small diffusion. 
For each $(k_{y},k_{z})$ pair, we solve the eigenvalue problem using $n_{x}=128$ modes; all our results are identical at double the resolution \footnote{See \protect\url{github.com/jsoishi/mri_prefers_3d} for all code used in this paper.}

For both the ideal and non-ideal MHD equations, we solve the eigenvalue problem in the $x$ direction using the \texttt{EigenValueProblem} solver in \emph{Dedalus} for a grid of $N_y \times N_z$ modes in the $y$ and $z$ directions.
For most of our runs, we use a targeted, sparse eigenvalue solver to find the 15 modes closest to a guess for the maximum growth rate.
In run 3, we have confirmed that dense solvers retrieve identical results.
For nearly ideal runs, we use the ideal 2D growth rate as input for the smallest $k_{y} > 0$ mode at each $k_{z}$ and then use the output from each previous $k_{y}$ as an input guess for the next mode.
For those runs that have significant $\eta$, we instead use a dense solve for the $k_{y} = 0$ modes and then step forward in $k_{y}$ at each $k_{z}$ as before.
Our solver is embarrassingly parallelised over the $k_{z}$ modes.
For the spectrum in figure 1 of the main text, we used a dense eigenvalue solver at $(k_{y}^{max},k_{z}^{max})$ for $\SSC= 1.002$.

In order to ensure our results are converged, we have repeated runs at $n_x=256$ Chebyshev modes as well as doubling the number of modes in $y$ and $z$.
\begin{table}
\begin{tabular}{cccccccccccccc}
\textbf{Run} & \textbf{$\SSC$} & \textbf{$\nu$} & \textbf{$\eta$} & \textbf{$\Ro$} & \textbf{$N_x$} & \textbf{$N_{k_{y}}$} & \textbf{$N_{k_{z}}$} & \textbf{Sparse/Dense}& \\
1  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & resolution study\\
2  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 256 & 200 & 200 & sparse & \\
3  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & dense  & \\
4  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 512 & 512 & sparse & \\
5  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 100 & 100 & sparse & \\
6  &   0.2  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & $\SSC$ variation\\
7  &   0.3  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
8  &   0.4  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
9  &   0.5  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
10 &   0.64 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
11 & 1.002  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
12 & 1.002  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & dense & \\
13 & 1.44   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
14 & 1.75   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
15 & 1.891  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
16 & 2.     & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
17 & 2.01   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
18 & 2.015  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
19 & 2.031  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
20 & 2.05   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
21 & 2.1    & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
22 & 2.25   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
23 & 2.5    & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
24 & 4      & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
25 & 1.02   & $10^{-6}$ & $10^{-6}$ & 0.75 & 128 & 200 & 200 & sparse & Reynolds number study\\
26 & 1.02   & $10^{-4}$ & $10^{-4}$ & 0.75 & 128 & 200 & 200 & sparse & \\
27 & 1.02   & $10^{-3}$ & $10^{-3}$ & 0.75 & 128 & 200 & 200 & sparse & \\
28 & 1.02   & $10^{-2}$ & $10^{-2}$ & 0.75 & 128 & 200 & 200 & sparse & \\
29 & 1.02   & $10^{-5}$ & $10^{-5}$ & 0.1  & 128 & 200 & 200 & sparse & Low Rossby\\ 
30 & 1.02   & $10^{-6}$ & $10^{-2}$ & 0.1  & 128 & 200 & 200 & sparse & Liquid metal case\\ 
\end{tabular}
\caption{Eigenvalue calculations performed.}
\label{tab:runs}
\end{table}
%\label{simulations}

\begin{figure*}[ht]
  \includegraphics[width=\textwidth]{fig_1.pdf}
  \caption{Growth rates for three-dimensional MRI modes. 
  a) growth rates $\gamma$ over a grid of $k_{y}$ and $k_{z}$ for four values of supercriticality $\SSC$. 
  The black contours are equally spaced between zero and the maximum growth rate.
  The grey contour highlights $\gamma=0$; the red dots indicate the fastest growing mode. 
Two dimensional modes occur at $k_{y} = 0$ on the bottom of each figure.
  At $\SSC=0.640$, there are no two dimensional modes. 
  b) Growth rate vs $k_{y}$ for $k_{z}=0.259$ at $\SSC=1.002$ (highlighted by the blue line in a). 
  The orange line gives the asymptotic result from equation~(\ref{eq:asymp}) and the blue line are the numerical results. 
  The asymptotic form is valid as $\SSC\to1$ and $k_{y}\ll1$.) c) The full discrete spectrum of the MRI for $(k_{y},k_{z})\simeq(0.263,0.447)$. 
  The unstable mode is plotted in orange, its stable complex conjugate is blue, and all other modes are grey. 
  d) The phase angle, $\arctan(k_{y}/k_{z})$, as a function of $\SSC$ showing the exact transition to purely two-dimensional fastest modes ($\phi=0$) at $\SSC\gtrsim2.05$. }
  \label{fig:growth_rate}
\end{figure*}

\section{Results}
\label{sec:results}

\subsection{Growth Rates and Three-dimensionality}
\label{sec:growth}

Our two main results are: 
(I) When the first two-dimensional mode becomes unstable, there already exist three-dimensional modes with positive growth rate. 
(II) At sufficiently large criticality, the fastest-growing mode becomes purely two-dimensional. 
These results are pertinent in two ways. 
First, the MRI contains a substantial ``Goldilocks regime'' with possible direct dynamo action, and this regime likely applies to stellar interiors and laboratory experiments. 
Second, our results accord with well-established results for accretion disks that expect two-dimensional primary modes.
Figure~\ref{fig:growth_rate}a shows the growth rates for four values of $\SSC$. 
At $\SSC=1.002$, the maximum growth rate ($\gamma_{\max}/|S|=0.093$) happens at $(k_{y},k_{z})\simeq(0.263,0.447)$.
Two-dimensional instability occurs for modes with finite growth rates at $k_{y} = 0$ along the bottom of the figure.
In all cases in figure~\ref{fig:growth_rate}a, the maximum growth rate does not occur along this line, indicating the dominance of three-dimensional modes.
The first panel of figure~\ref{fig:growth_rate}a shows $\SSC=0.640$, which contains only three-dimensional instability: the zero growth rate contour (highlighted in grey) does not intercept the $k_{z}$-axis.

\subsection{Asymptotic Calculation}
\label{sec:asymp}
The MRI's preference for three-dimensional modes can be predicted analytically.
We analyse equations~(\ref{eq:mhd}-\ref{eq:divu}) asymptotically close to $S_{c}$, and compute the leading-order correction to the growth rate from three-dimensional effects assuming $\SSC=1+\epsilon^{2}R$, $k_{z}\sim\mathcal{O}(\epsilon)$, $\sigma\sim{k_{y}}\sim\mathcal{O}(\epsilon^{2})$. 
Here, we outline the asymptotic calculation for the leading order correction to the 2D growth rates when accounting for three dimensional effects via finite $k_{y}$.
Being in the rotating frame means the shear has no net
\Beq
\int_{-d/2}^{d/2} V_{0,y}(x) \text{d} x  \ = \ 0 \quad \implies  \quad V_{0,y}(x) \ = \ S\, x
\Eeq
The full linear ideal equations are:
\Beq\label{u-eq}
\pd{t} v_{x} + S x \pd{y} v_{x} - f v_{y} + \pd{x} p - B_{0} \pd{z} b_{x} &=& 0\\
\pd{t} v_{y} + S x \pd{y} v_{y} + (f+S) v_{x} + \pd{y} p - B_{0} \pd{z} b_{y} &=& 0\label{v-eq} \\
\pd{t} v_{z} + S x \pd{y} v_{z} + \pd{z} p - B_{0} \pd{z} b_{z} &=& 0 \\
\pd{x} v_{x} + \pd{y} v_{y} + \pd{z} v_{z}  &=& 0\\
\pd{t} b_{x} + S x \pd{y} b_{x} - B_{0} \pd{z} v_{x} &=& 0\\
\pd{t} b_{y} + S x \pd{y} b_{y} - B_{0} \pd{z} v_{y} - S b_{x}  &=& 0\\
\pd{t} b_{z} + S x \pd{y} b_{z} - B_{0} \pd{z} v_{z} &=& 0 \label{bz-eq}
\Eeq
The divergence-free condition for the velocity implies the same for the magnetic field.
The system is 2nd order in $x$ because $\pd{x}$ only appears in two places.
All variables take the form
\Beq
p = \hat{P}(x) e^{ i ( \omega t + k_{z} z + k_{y} y ) },
\Eeq
where we note that for this section, we use the complex frequency $\omega = -i \sigma$ for the time dependence.

To make analytical progress, define the frequency ``parameters'' 
\Beq
\omega_{S}(x) \ \equiv \ \omega + S x k_{y}, \quad \omega_{A}  \ \equiv \ B_{0} k.
\Eeq
We can use equations (\ref{v-eq}--\ref{bz-eq}) to find all amplitudes in terms of $\hat{V}_{x}(x)$ and $\hat{V}_{x}'(x)$. 
\Beq
\hat{V}_{y}(x) &=& \frac{i \left(k_{y} 
   \hat{V}_{x}'(x) + k_{z}^2 \frac{ \left((f+S) \omega
   _S^2-S \omega _A^2\right)}{\omega_{S}
   \left(\omega_{S}^2-\omega _A^2\right)} \hat{V}_{x}(x)\right)}{k_{z}^2+k_{y} ^2} \\
\hat{V}_{z}(x) &=& \frac{i k_{z} \left(\hat{V}_{x}'(x)- \frac{ \left((f+S) \omega
   _S^2-S \omega _A^2\right)}{\omega_{S}
   \left(\omega_{S}^2-\omega _A^2\right)}  k_{y} 
   \hat{V}_{x}(x)\right)}{k_{z}^2+k_{y} ^2} \\
   \hat{P}(x) &=& \frac{i \left(\omega _A^2-\omega_{S}^2\right)
   \left(\hat{V}_{x}'(x) -  k_{y} \frac{ \left((f+S) \omega
   _S^2-S \omega _A^2\right)}{\omega_{S}
   \left(\omega_{S}^2-\omega _A^2\right)}
   \hat{V}_{x}(x)\right)}{\left(k_{z}^2+k_{y} ^2\right)
   \omega_{S}} \\
   \hat{B}_{x}(x) &=& \frac{\omega _A \hat{V}_{x}(x)}{\omega_{S}} \\
   \hat{B}_{y}(x) &=&\frac{\omega _A \hat{V}_{y}(x)}{\omega_{S}} -\frac{i S \omega _A \hat{V}_{x}(x)}{\omega_{S}^2} \\
   \hat{B}_{z}(x) &=& \frac{\omega _A \hat{V}_{z}(x)}{\omega_{S}}
   \Eeq
One can substitute everything into equation (\ref{u-eq}) to get a second-order equation for $\hat{V}_{x}(x)$ of the general form 
\Beq
 \hat{V}_{x}''(x) + 2 C_{1}(x) \hat{V}_{x}'(x) + C_{0}(x) \hat{V}_{x}(x) \ = \ 0,
\Eeq
where
\Beq
   C_{1}(x) \ = \  \frac{ S k_{y}  \omega _A^2}{\omega_{S} ( \omega_{S}^2-\omega _A^2 )} , \quad C_{0}(x) \ = \   \frac{\frac{f^{2} k_{z}^2 \omega_{S}^{2}}{\omega_{S}^{2}-\omega
   _A^2}+f  S  k_{z}^2 -\frac{2 S^2  k_{y}^2 \omega _A^2}{\omega_{S}^{2}}}{\omega_{S}^2-\omega _A^2}  -\left(k_{y}^{2} + k_{z}^2 \right) .
\Eeq
We can eliminate the first-order term via
\Beq
\hat{V}_{x}(x) \ = \ \chi(x)\, \psi(x) \quad \text{where} \quad \frac{\chi'(x)}{\chi(x)} \  = \ - C_{1} (x).
\Eeq
Therefore 
\Beq
-\psi''(x) + \Phi(x) \psi(x) \ = \ 0 \label{Schr-eq}, \quad \text{where}\quad
\Phi(x)  \ = \ C_{1}'(x) + C_{1} (x)^{2} - C_{0}(x).
\Eeq
Finally, by putting everything together 
\Beq
\Phi(x) \ = \ \frac{S  \left(f k_{z}^2-S k_{y}
   ^2\right) \omega _A^2 -f k_{z}^2 (f+S) \omega
   _S^2}{\left(\omega _A^2-\omega
   _S^2\right)^2} + k_{y}^{2} + k_{z}^2 .
\Eeq
The denominator of $\Phi(x)$ cannot vanish if $\omega_{S}=-i \gamma+Sxk_{y}$, with $Re(\gamma)\neq0$. 
We want to solve the Schr\"{o}dinger-type equation~(\ref{Schr-eq}) with the boundary conditions 
\Beq
\psi(x=\pm d/2) = 0.
\Eeq
The boundary conditions for $\psi(x)$ are the same as $V_{x}(x)$ because 
\Beq
\chi(x) \ \propto \ \frac{\omega_{S}(x)}{\sqrt{\omega_{S}(x)^2-\omega _A^2}} \  \ne \ 0 \quad \text{for} \quad \gamma \ \ne \ 0.
\Eeq
We solve perturbatively near the critical values for the 2D instability. We rescale each parameter in terms of a bookkeeping parameter, $\eps$.
\Beq
S  \ \to \  - \frac{\pi ^2 B_{0}^2}{d^2 f} ( 1 + \eps^{2} R), \quad \omega \ \to \ \eps^{2} \omega , \quad k_{z} \ \to \ \eps k_{z}, \quad k_{y} \ \to \ \eps^{2} k_{y},
\Eeq 
keeping in mind
\Beq
\Ro  \ \equiv  \ \frac{\pi ^2 B_{0}^2}{d^2 f^{2}}.
\Eeq
Expanding equation~(\ref{Schr-eq}) to $\mathcal{O}(\eps^{2})$, we arrive at
\Beq
\Phi(x) \ = \ -\frac{\pi^{2}}{d^{2}} - \frac{\pi^{2} \eps^{2}}{d^{2}}\left[R - \frac{d^2 }{\pi^2}k_z^2 + \frac{\Ro^2 B_0^2  k_y^2+(1+\Ro) (\omega + f \Ro\, x k_y) ^2}{ \Ro\, B_0^2
   k_z^2} \right].
\Eeq
The leading order balance is
\Beq
\psi_{0}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{0}(x) \ = \ 0  \quad \implies \quad \psi_{0}(x) \ = \ \cos(\pi x / d),
\Eeq
and next order is
\Beq
\psi_{2}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{2}(x) \ = \  \Phi_{2}(x)\psi_{0}(x) .
\Eeq
The solvability condition for $\psi_{2}(x)$ is 
\Beq
\int_{-d/2}^{d/2} \Phi_{2}(x)\psi_{0}(x)^{2} \text{d} x \ = \ 0.
\Eeq
This yields our final result
\begin{equation}\label{eq:asymp}
\sigma^{2}\ = \ - \omega^{2} \ = \  \frac{S_{c}^{2}}{1+\Ro} \frac{d^{2}}{\pi^{2}} \left[ k_{z}^{2}\left(R-\frac{d^2}{\pi^{2}}k_{z}^{2}\right)+\Upsilon\,k_{y}^{2}\right]+\ldots,
\end{equation}
where
\begin{equation}
\Upsilon \ = \  \Ro + (1+ \Ro) \frac{\pi^{2}-6}{12}  \ \approx \ 1.31\quad\text{for}\quad{\Ro} \ = \ 0.75,
\end{equation}
and we have switched back to $\sigma$ for time dependence.
The first term in equation~(\ref{eq:asymp}) results from the two-dimensional calculation.
The second term is positive definite: it \emph{always} leads to enhanced growth rates when $k_{y}\neq0$ and ultraviolet divergence in the absence of higher-order effects.

Figure~\ref{fig:growth_rate}b compares the numerical growth rates for $\SSC=1.002$ between $0\le{k_{y}}\le0.2$ to the asymptotic approximation, showing good agreement where the latter is valid.
Figure~\ref{fig:growth_rate}c shows the full spectrum for $\SSC=1.002$.
The plot shows a purely growing/decaying complex-conjugate pair (orange/blue dots on the real axis) consistent with equation~(\ref{eq:asymp}).
The other stable modes (grey dots) are rotationally modified Alfv\'{e}n waves found in left- and right-going pairs, consistent with the analytic predictions for two-dimensional stability calculations.
We reject spurious eigenvalues by using \texttt{eigentools} \footnote{\protect\url{https://github.com/dedalusproject/eigentools}} to solve equations~(\ref{eq:mhd}-\ref{eq:divu}) at two resolutions and retain only pairs equal to within one part in $10^{-6}$.

\subsection{Analysis of the transition to 2D}
\label{sec:2D_transition}

We can understand the transition from three- to two-dimensional instability semi-analytically. We do this by conducting a similar analysis as the asymptotic calculation in section~\ref{sec:asymp} near the fastest growing 2D modes, looking for values for $\SSC$ where the analogous value of $\Upsilon(\SSC)$ equals zero.

The growth rate can be written as a function of  $k_{y},k_{z}$ and $\SSC$ near $k_{y} \approx0$.  We therefore expand the whole problem in a power-series 
\Beq
\gamma(\SSC,k_{y},k_{z}) \ = \   \sum_{n\ge 0} \gamma_{n}(\SSC,k_{z}) k_{y}^{n}, \quad \psi(x) \ = \ \sum_{n\ge 0} \psi_{n}(x) k_{y}^{n},
\Eeq
and use the same asymptotic techniques as before to obtain each progressive correction,
\Beq
\psi_{0}'' - \Phi_{0} \psi_{0} & = & 0 \label{eq:psi0} \\
\psi_{1}'' - \Phi_{0} \psi_{1} & = & \Phi_{1} \psi_{0} \label{eq:psi1} \\
\psi_{2}'' - \Phi_{0} \psi_{2} & = & \Phi_{2} \psi_{0} + \Phi_{1} \psi_{1}.
\label{eq:psi2}  
\Eeq
At leading order, we making use of the fact that all non-constant coefficients vanish for 2D modes. The leading-order eigenfunction is $\psi_{0} = \cos(\pi x / d)$, just as before. The following dispersion relations determine the maximum growth rate and the corresponding wavenumber,
\Beq
\mathfrak{D}_{0}(k_{z}) &=& \frac{f k_z^2 \left(B_0^2 S k_z^2+\gamma _0^2 (f+S)\right)}{\left(B_0^2 k_z^2+\gamma _0^2\right){}^2}+\frac{\pi ^2}{d^2}+k_z^2 \ = \ 0 \\
\frac{\mathfrak{D}_{0}'(k_{z})}{2k_{z}} &=& \frac{\gamma _0^2 f \left(B_0^2 k_{z}^2 (S-f)+\gamma _0^2 (f+S)\right)}{\left(B_0^2 k_{z}^2+\gamma _0^2\right)^3}+1 \ = \ 0.
\Eeq
We transform these relations into polynomials in terms of non-dimensional variables,
\Beq
\label{dispersion}
\y^2 \s^2 \Ro^2 +  \y \, \s  \, ( \y \, \s \, \Ro^2 + (2  - \s)  \, \Ro + \s ) \x + ( 2 \y \,\s \,  \Ro - \s  + 1 ) \x^2 +\x^3 &=& 0 \\ 
\y \, \s  \, (\y \, \s \, \Ro^2 + (2  - \s)  \, \Ro + \s    )  + 2 ( 2 \y \,\s \,  \Ro - \s  + 1 ) \x +3\x^2 &=& 0,
\label{critical-dispersion}
\Eeq
where
\Beq
\x  \ = \ \frac{d^2 k_z^2}{\pi ^2}, \quad  \quad \y \ = \ \frac{\gamma_{0}^{2}}{S^{2}}, \quad \quad \s  \ = \ \frac{S}{S_{c}}.
\Eeq
These can be computed from the parameters reported in figure~\ref{fig:growth_rate} and table~\ref{tab:runs}.
The two polynomials are straightforward to solve numerically for $\x,\y$ given $\s, \Ro$. However, we want to find critical values for $\s$, where the fastest growing modes are two dimensional.

We find the higher-order $\gamma_{n}(\s,k_{z})$ coefficients by applying order-by-order solvability conditions. We first find that $\gamma_{1}(\s, k_{z}) = 0$ identically. This renders the first-order correction \eq{eq:psi1} solvable. The 1st-order eigenfunction takes the form 
\Beq
\psi_{1}(x) \ = \ A_{1} F(\pi x/d) \quad \text{where} \quad  F(\vartheta) = \vartheta (\vartheta \sin (\vartheta)+\cos (\vartheta))-\frac{\pi^{2}}{4} \sin (\vartheta).
\Eeq
and
\Beq
A_{1} \ = \ \frac{i \gamma _0 d^3 f S k_z^2 \left(B_0^2 (f-S) k_z^2-\gamma _0^2
   (f+S)\right)}{2 \pi ^3 \left(B_0^2 k_z^2+\gamma _0^2\right){}^{\!3}}.
\Eeq
The structure of the growth rate is therefore 
\Beq
\gamma(\s,k_{y},k_{z}) \ = \ \gamma_{0}(\s,\x) + \gamma_{2}(\s,\x) k_{y}^{2} \ + \ \ldots 
\Eeq
This is the same structure as the asymptotic calculation in section~\ref{sec:asymp}. Here, $\Upsilon \propto \gamma_{2}$ effectively. The case with dominant 2D modes corresponds to $\gamma_{2}(\s,\x)=0$.  The final result derives from the 2nd-order solvability condition 
\Beq
\int_{-d/2}^{d/2} ( \Phi_{2} \psi_{0} + \Phi_{1} \psi_{1} ) \psi_{0} \dd{x}  \ = \ 0,
\Eeq
which produces the degree-6 polynomial expression  
\Beq
& \y^{6} \s^6\Ro^6  \ + \
 \y^4 \s^5 \Ro^4   \left( (3 \xi  \s+6 \y-1)\Ro -3 \xi  \s \right) \, \x +
  \nonumber \\  
& \frac{ \y^3 \s^4 \Ro^2  }{4}  \left( (60 \y -4 \xi  (\s-8) \s -16 + 3 \s^2)\Ro^{2} + 2 \s (4 \xi  (\s+1)-3 \s) \Ro - (4 \xi -3 ) \s^2 \right) \, \x^{2} + \nonumber \\ 
& \frac{\y^2  \s^3 \Ro }{2}  \left( (3 \s^2-4 \xi  (\s-3) \s+40 \y-12) \Ro^{2} + 24 \xi \s \Ro + (4 \xi -3) s^2 \right) \, \x^{3} + 
\nonumber \\ 
&  \frac{\y \,\s^2}{4} \left( (60 \y -(4 \xi -3) \s^2 - 16)\Ro^{2} + 2\s(3 \s-4 \xi  (\s-3)) \Ro - (4 \xi -3) \s^2 \right)\, \x^{4} +\nonumber \\ 
 &  \s\, (( 6 \y -\xi  \s-1) \Ro- \xi  \s) \, \x^{5}  \ + \   
   \x^{6} \ = \ 0.
   \label{2D-condition}
\Eeq
The new coefficient is
\Beq
\xi \ = \ \frac{\pi^{2}-6}{12} \ = \ \frac{\int_{-d/2}^{d/2} x^2 \cos ^2\left(\frac{\pi  x}{d}\right) \dd{x}}{\frac{d^{2}}{\pi^{2}}\int_{-d/2}^{d/2}  \cos ^2\left(\frac{\pi  x}{d}\right) \dd{x}} \ \approx \ 0.322467.
\Eeq
This results from projecting the squared shear onto the leading-order eigenmodes, just as in the previous section.

We solve \eq{dispersion}, \eq{critical-dispersion}, and  \eq{2D-condition} for $\x,\y,\s$ using Newton's method for a variety of $\Ro$, yielding

\begin{center}
\begin{tabular}{cccccccccccccc}
\textbf{$\Ro$} & \textbf{$k_{z} d/\pi$}  & \textbf{$\gamma_{\max}/|S_{2D}|$}  &  \textbf{$S_{2D}/S_{c}$} \\
                   1.00  & 0.689610 & 0.208727 &   2.09694 \\ 
                   0.75 & 0.682819 & 0.214895 & 2.05136 \\
                   0.50 & 0.676062 & 0.221309 & 2.00061 \\
                   0.25 & 0.671528 & 0.228551 & 1.94963 \\
                   0.10 & 0.673164 & 0.234376 & 1.92692 \\
                   0.00 & 0.462324 & 0.240212 & 1.92465
\end{tabular}
\end{center}
The value of $\SSC = 2.05136$ at the 2D to 3D transition when $\Ro = 0.75$ matches the numerical results shown in Figure~\ref{fig:growth_rate} two all reported significant figures in the numerical calculation.

We conclude this section with some speculations concerning the higher order corrections. Because of symmetry, $\gamma_{3}(\s,\x) = 0$ identically. This means that to leading order 
\Beq
\gamma(\s,k_{y},k_{z}) \ = \ \gamma_{0}(\s,\x) + \gamma_{2}(\s,\x) k_{y}^{2}  + \gamma_{4}(\s,\x) k_{y}^{4} \ + \ \ldots
\Eeq 
For $\SSC > S_{2D}/S_{c}$, $\gamma_{2} > 0$. This is sufficient to determine the most unstable 3D wavenumber near the transition to dominant 2D modes, 
\Beq
k_{y}^{2} \ \approx \ -\frac{\gamma _2}{2 \gamma_4} \  \propto \   S_{2D}/S_{c} - \SSC   . 
\Eeq 
In general, $\gamma_{4} \ne 0$. We speculate that $\gamma_{4} < 0$, though the calculation is outside the scope of this work. If there is a negative 4th-order feedback, then 
\Beq
k_{y} \ \sim \ \sqrt{S_{2D}/S_{c} - \SSC}, 
\Eeq
which is consistent with Figure~\ref{fig:growth_rate}d near the critical point. However, a scenario where $\gamma_{4} > 0$ would require behaviour that is inconsistent with our numerical calculations.

\subsection{Consistency with previous simulations}
\label{sec:prior_work}

Numerical simulations of the MRI in accretion disks consistently show axisymmetric modes dominating the early evolution of the MRI before breaking down into 3D MHD turbulence \cite{1995ApJ...440..742H,2018ApJ...853..174H,2019ApJS..241...26D}. 
These ``channel modes'' are exact non-linear solutions for $x$-shearing-periodic domains, and can only saturate via parasitic shear instabilities \cite{1994ApJ...432..213G}.
Impenetrable and stress-free boundary conditions are more applicable to stellar interiors. 
But even in disks, any kind of finite radial extent will cutoff the unbounded growth of channel modes. 
We reconcile our results with earlier simulations by showing that the MRI indeed prefers two dimensions for the larger values of criticality found in disk simulations. 
Figure~\ref{fig:growth_rate}d shows the phase angle $\phi=\arctan(k_{y}/k_{z})$ of the fastest growing mode as function of $\SSC$.
The overall critical shear for three-dimensional modes is $S_{c,\text{3D}}\simeq0.102S_c$.
Above $\SSC\gtrsim2.05$, $\phi$ becomes zero, indicating that axisymmetric modes have the fastest growth rates.
For the fiducial run in \cite{1996ApJ...464..690H}, $\SSC\simeq4.84$ (most works use similar values).
Thus, our theory predicts axisymmetric modes should dominate the linear dynamics for parameters studied in prior numerical simulations.
However, it may be that those simulations saw a predominance of axisymmetric modes because of the radial boundary conditions and not their large shear rates.

Even for $\SSC> 2.05$ there are significant swaths of unstable three-dimensional modes with growth rates comparable to the maximum (figure~\ref{fig:growth_rate}a).
Also, non-normality generically accompanies non-axisymmetry in shearing systems \cite{1992MNRAS.255P..25K}.
In non-normal systems, transient amplification can occur for stable modes, and can even cause turbulence.
Therefore significant three dimensionality likely implies important non-normal behaviour near onset.

\subsection{Eigenvectors}
\label{sec:eigenvectors}

\begin{figure}[h!]
  \centering
  \includegraphics[width=\columnwidth]{eigvecs_xy_run_11_ideal_single_mode.pdf}
  \caption{\textit{Ideal} eigenvectors of the velocity and magnetic field perturbations for the most unstable mode when $\SSC=1.02$, and $\eta=\nu=0$; calculations at $\eta=\nu=10^{-5}$ are indistinguishable.
  The top row shows $v_{x}$, $v_{y}$, $v_{z}$ (red is negative, blue positive); the bottom row shows $b_{x}$, $b_{y}$, $b_{z}$ (purple is negative, orange positive). 
The amplitudes are arbitrary because of linearity.}
  \label{fig:eigvec}
\end{figure}

Figure~\ref{fig:eigvec} shows the eigenvectors for the most unstable mode at $\SSC=1.02$ using ideal MHD. 
No critical layers can form for $\gamma\ne0$.
We therefore solve equations~(\ref{eq:mhd}--\ref{eq:divu}) without dissipation as a 2nd-order system in $\partial_{x}$ and impose $v_{x}=0$ at $x=\pm{d/2}$.
The eigenfunctions are indistinguishable from those at finite dissipation, lending additional confidence to our other results.
The tilted structures of $v_{x}$ and $v_{y}$ as well as $b_{x}$ and $b_{y}$, imply non-trivial Reynolds and Maxwell stresses.

\subsection{Reynolds Number and Departure from Ideal MHD}
\label{sec:reyn}

In order to ensure that our calculations are probing the ideal MHD regime we are interested in, we have performed a series of runs (24-27) with $\eta = \nu$ ranging from $10^{-6}$ to $10^{-2}$, equivalent to varying the Reynolds and Magnetic Reynolds numbers $\Reyn = \Reym = 10^2$ -- $10^6$.
\begin{figure}[h!]
  \centering
  \includegraphics[width=\textwidth]{re_plots.pdf}
  \caption{Growth rates at $\SSC = 1.02$ for varying diffusivities. Note that the top three panels are essentially identical, suggesting that above $\Reyn \gtrsim 10^4$ we are reasonably close to ideal conditions.}
  \label{fig:reynolds}
\end{figure}
Figure~\pageref{fig:reynolds} shows the growth rates at $\SSC=1.05$ for five values of $\Reyn = \Reym$.
For $\Reyn \gtrsim 10^4$, the maximum growth rate is located at the same position and the same growth rate; the unstable region expands, as expected.
This suggests that our results at $\Reyn = 10^5$ are reasonably close to ideal MHD for the purposes of understanding the MRI's preference for 3D modes near shear onset.

\subsection{Mean electromotive forces}
\label{sec:emf}
We should expect the generation of a mean electromotive force (EMF) in the linear regime, \textit{i.e}
$\left<\boldsymbol{v}\times\boldsymbol{b}\right>\sim\,\alpha\!\left<\boldsymbol{B}_{0}\right>$.
This is significant because of the possibility of direct laminar dynamo action in regions of weak shear at large Reynolds number.
\begin{figure}[h!]
  \includegraphics[width=\columnwidth]{mean_field_alpha_run_11_output.pdf}
  \caption{Velocity-magnetic field correlations relevant to the mean electromotive force for the eigenvectors of the fastest growing mode at $\SSC = 1.02$.}
  \label{fig:correlation}
\end{figure}
Figure~\ref{fig:correlation} shows the correlation (averaged over $y,\,z$) as a function of $x$ over the domain for $\SSC = 1.02$,
\begin{equation}
\text{EMF}\propto\frac{\left<\boldsymbol{\hat{z}}\cdot(\boldsymbol{v}\times\boldsymbol{b})\right>_{\!y,z}}{\boldsymbol{v}_{\text{rms}}\boldsymbol{b}_{\text{rms}}},
\end{equation}
where ``rms'' is denotes the average over the whole domain. 
The correlation is independent of the arbitrary normalisation of the linear eigenvectors.
The non-zero correlation shows the tendency for back reaction on the mean magnetic field.

MRI dynamos have been studied in a number of different contexts \cite{2007PhRvL..98y4502R,2011ApJ...740...18O,2015PhRvL.114h5002S}, but all except one focused on non-linear, usually turbulent, dynamos. 
The only exception we are aware of, \cite{2016MNRAS.462..818B}, found exponential growth of mean magnetic fields during the linear growth phase of the MRI far from stability in numerical simulations.
Our work explains this result in terms of purely linear dynamics:
non-axisymmetric MRI unstable modes drive the dynamo growth of magnetic fields.

\subsection{Extensions to stellar and experimental conditions}
\label{sec:extensions}

Finally, we vary two important parameters:
$\Ro$ and the magnetic Prandtl number $\Prm=\nu/\eta$.
\begin{figure}[h!]
  \centering
  \includegraphics[width=\columnwidth]{low_rossby_liquid_metal_growth_rates.pdf}
  \caption{(a) Growth rates for $\Ro=1/10$ relevant to the stellar interiors. 
  (b) Growth rates for MRI near onset at liquid metal-like diffusivities, $\nu=10^{-6}$, $\eta=0.1$.
The lack of two dimensional instability is evident by the fact that the zero contour does not intercept the x-axis.}
  \label{fig:other_params}
\end{figure}
Figure~\ref{fig:other_params} shows the growth rates for $\SSC=1.02$.
The upper panel shows $\Ro=1/10$ (addressing a possible range within stellar interiors) holding all other parameters equal to their fiducial values.
The growth rates show similar behaviour to the $\Ro=3/4$ cases, \textit{i.e.}, dominated by three-dimensional modes with $\gamma_{\max}/|S|\simeq0.1$.
The lower panel shows $\nu=10^{-6}$ and $\eta=0.1$ with all other parameters equal to their fiducial values. 
This case is relevant to liquid metals e.g., the Princeton MRI experiment \cite{2002JFM...462..365G}
at $\Prm=10^{-5}$. 
Our results are for the rotating plane Couette geometry (Taylor-Couette in the small-gap limit $R_{2}\approx{R}_{1}$) with highly idealised boundary conditions.
It is nevertheless quite interesting that, near onset, the low-$\Prm$ MRI is \emph{only} unstable to three-dimensional modes.

\section{Conclusion}
\label{sec:conclusion}

Our results show that three-dimensional modes grow faster than two-dimensional modes whenever the MRI is near its critical shear values.
While we mainly focus on the ideal Keplerian case ($\Ro=3/4$), we also demonstrate robustness for low Rossby ($\Ro=1/10$) and magnetic Prandtl numbers ($\Prm=10^{-5}$).
There are several important future directions this work suggests.
First, the Sun possesses two internal shear layers with inwardly increasing shear, the high-latitude tachocline, and the near-surface shear layer (NSSL). 
Past work has already pointed out the possibility of the small-scale MRI in the Sun using local analysis \cite{2007ApJ...667L.207P,2011MNRAS.411L..26M,2014ApJ...787...21K}.
The NSSL, in particular, may also host slower MRI-driven dynamics, despite containing small-scale convection.
The NSSL contains the strongest shear anywhere within the solar interior;  $\Omega(r) \ \propto \ 1/r$, implying $\Ro = 1/2$.
It is therefore crucial to further elucidate the nonlinear saturation of the 3D MRI, along with its robustness to convection.
An important uncertainty with regard to stellar applications is the fact that $\SSC$ depends on the square of the magnetic field strength, a quantity that is rather uncertain.
Second, our low-$\Prm$ results suggest that three-dimensional MRI modes may be the easiest to excite in liquid metal experiments.
Determining possible non-axisymmetric signatures in Taylor-Couette experiments requires followup work using more realistic boundary conditions and geometry. 
Our prior work on the axisymmetric MRI \cite{2017ApJ...841....1C,2017ApJ...841....2C} shows only small differences between rotating plane and cylindrical Taylor-Couette geometries. 
We fully expect the general three-dimensional features of the MRI to persist in more complex applications.

%\ethics{Insert ethics text here.}

\dataccess{All code used for this project is available at \url{https://github.com/jsoishi/mri\_prefers\_3D}}

\aucontribute{J\@.S\@.O and G\@.M.\@V\@. led the project.
J\@.S\@.O prepared all figures.
G\@.M\@.V performed the asymptotic calculation.
M.B.\ ran the bulk of the eigenvalue calculations and prepared several preliminary figures.
A\@.S\@. performed preliminary eigenvalue calculations which led to the fiducial parameter choices used here.
J\@.S\@.O, G\@.M.V\@., K\@.J\@.B., D\@.L\@., and B.P\@.B developed \emph{Dedalus}, which is used for all calculations in this paper.
All authors contributed to the writing of the manuscript and physical interpretation of the results.
}

\competing{We declare we have no competing interests.}

\funding{Oishi, Baxter, and Brown acknowledge support from NASA LWS grant No. NNX16AC92G.
Oishi also acknowledges support from Research Corporation Scialog Collaborative Award (TDA) ID \#24231.}

\ack{Computations were performed on the \emph{Leavitt} cluster at the Bates College High Performance Computing Centre.}

%\disclaimer{Insert disclaimer text here.}

\bibliographystyle{RS}
\bibliography{mri}

\end{document}

%%% Local Variables:
%%% ispell-local-dictionary: "british"
%%% End:

%  LocalWords:  Magnetorotational Oishi Lewiston Vasil NSW Lecoanet
%  LocalWords:  magnetorotational Keplerian magnetofluid Prandtl MHD
%  LocalWords:  dimensionality Couette axisymmetric linearized RSPSA
%  LocalWords:  magnetohydrodynamic Navier Dt Dj dV dx Alfv Rossby fd
%  LocalWords:  hydrodynamical ODEs Dedalus arXiv supercriticality nd
%  LocalWords:  criticality eigenvectors fiducial axisymmetry NSSL Sx
%  LocalWords:  eigenfunctions eigenmodes diffusivities tachocline eq
%  LocalWords:  Tobias JSO LWS NNX Scialog TDA Leavitt mri DPMMS rms
%  LocalWords:  eigentools EigenValueProblem parallelised Chebyshev
%  LocalWords:  cccccccccccccc analytical Schr dinger perturbatively
%  LocalWords:  solvability Center resistivity dimensionalization Sxk
%  LocalWords:  eigenfunction wavenumber th

### 2021-07-14 paper/submitted_main_text.tex
initial release

% PRL Length limits
% Displayed Math 	The word equivalent for displayed math is 16 words per row for single-column equations. Two-column equations count as 32 words per row.
% Figures 	To estimate the word equivalent for figures use the figure’s aspect ratio (width / height). The estimate is [(150 / aspect ratio) + 20 words] for single-column figures, and [300 / (0.5 * aspect ratio)] + 40 words for double-column figures.

% Fig 1: A=w/h=2. Double column, words=340
% Fig 2: A=4/3. Single column, words=133
% Fig 3: A=1. Single column, words=170
% Fig 4: A=0.5 Single column, words=320
% equations: 8 * 16=128

% current word count: 2057 + 128 + 340 + 133 + 170 + 320
% limit: 3750
\documentclass[aps,prl,reprint,superscriptaddress]{revtex4-1}
\newcommand{\SSC}{S/S_{c}}
\newcommand{\Prm}{\mbox{\textit{Pm}}}
\newcommand{\mnras}{Monthly Notices of the Royal Astronomical Society}
\newcommand{\apjs}{The Astrophysical Journal Supplement Series}
\newcommand{\apjl}{Astrophysical Journal Letters}
\usepackage{graphicx}
\usepackage{amsmath}

\begin{document}

\title{The Magnetorotational Instability Prefers Three Dimensions}

\author{Jeffrey~S.~Oishi}
\email[]{joishi@bates.edu}
\affiliation{Bates College Department of Physics and Astronomy, Lewiston, ME 04240}

\author{Geoffrey~M.~Vasil}
\affiliation{University of Sydney School of Mathematics and Statistics, Sydney, NSW, Australia}
\author{Morgan Baxter}
\affiliation{Bates College Department of Physics and Astronomy, Lewiston, ME 04240}
\author{Andrew Swan}
\affiliation{Statistical Laboratory, DPMMS, University of Cambridge, Cambridge,
United Kingdom}
\author{Keaton~J.~Burns}
\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010}
\affiliation{Massachusetts Institute of Technology Department of Physics, Cambridge, MA 02139}
\author{Daniel~Lecoanet}
\affiliation{Princeton University Department of Astrophysical Sciences, Princeton, NJ 08544}
\author{Benjamin~P.~Brown}
\affiliation{University of Colorado Laboratory for Atmospheric and Space Physics and Department of Astrophysical and Planetary Sciences, Boulder, CO 80309}

\date{\today}

\begin{abstract}
The magnetorotational instability (MRI) occurs when a weak magnetic field destabilizes a rotating electrically conducting fluid with an inwardly increasing angular velocity shear.
The MRI is considered essential to the operation of astrophysical accretion disks, with the central object supplying a Keplerian rotation profile.
Internal shear layers in stars may also be unstable to the MRI.
Stellar-interior shears can take a wide range of profiles, including near-critical, but with negligible dissipation. 
Here, we show that the fastest growing unstable modes of an ideal magnetofluid are three dimensional provided the shear rate, $S$, is close to the two-dimensional onset value, $S_c$.
For a Keplerian shear, three-dimensional modes are unstable down to $S\approx0.10S_c$, and dominate the two-dimensional modes until $S\approx2.05S_{c}$.
Significant numbers of rapidly growing three-dimensional modes remain well past $2.05S_{c}$. 
These finding are significant in three ways. 
First, evidence from weakly nonlinear theory suggests that the MRI saturates by pushing the shear rate to its critical value. 
This can happen for systems, like stars and laboratory experiments, that can rearrange their angular velocity profiles.
Second, the non-normal character and large transient growth of MRI modes should be important anytime strong three-dimensionality exists.
Finally, three-dimensional growth suggest the possibility of direct dynamo action driven from the linear instability itself.
We also demonstrate that three-dimensional modes dominate for shear profiles relevant to stars and at the tiny magnetic Prandtl numbers relevant to liquid-metal laboratory experiments.
\end{abstract}

\pacs{}
\maketitle

The magnetorotational instability (MRI) is extremely important in astrophysical fluid dynamics.
A weak magnetic field catalyzes turbulence in a Keplerian shear by changing the stability criterion for differentially rotating flows from a negative angular \emph{momentum} gradient to a negative angular \emph{velocity} gradient \citep[e.g.][]{1998RvMP...70....1B,2010RSPTA.368.1607J}.
This discovery explained the ubiquitous accretion onto compact objects at rates compatible with observations, and may also influence the formation of planets \citep[e.g.][]{2007Natur.448.1022J}.
In disks, the gravitational field dominates the local plasma dynamics, and thus the MRI cannot significantly affect the background shear; it must saturate by other means \citep{2018MNRAS.474.3451X}.
However, stars and liquid metal Taylor-Couette experiments have differential rotation profiles driven by much weaker stresses.
Where the MRI is active in these flows, it saturates by pushing the background shear close to critical \citep{2015RSPSA.47140699V,2017ApJ...841....1C,2017ApJ...841....2C}, analogous to convection mixing entropy.
Stellar interiors have extremely high fluid and magnetic Reynolds numbers, but can operate at or near the critical shear rate for the MRI.
This finite critical shear results from a finite-channel cutoff.
Despite the extensive literature on accretion disks (strong shear, low dissipation) and liquid-metal experiments (weak shear, large dissipation), even the linear MRI is not well understood in the weak-shear, low-dissipation regime.

Here, we investigate the stability of three-dimensional perturbations near the two-dimensional critical shear rate $S_{c}$ for a nearly inviscid, ideal MHD flow.
We find that the first destabilized modes are three-dimensional, and thus could act as a dynamo even in the absence of secondary instability.
These results also suggest that the non-normal growth of the MRI is always important, even when axisymmetric modes dominate.

We numerically solve the linearized magnetohydrodynamic equations in rotating plane Couette geometry.
This corresponds to a Cartesian frame rotating with angular frequency $\Omega$ and a linear background shear, $V_{y}=Sx$; \citep[see][]{2015RSPSA.47140699V}.
We cast the Navier-Stokes equation in the form,
\begin{equation}\label{eq:mhd}
\frac{D \boldsymbol{v}}{Dt}+f \boldsymbol{\hat{z}}\times\boldsymbol{v}+{S}v_{x}\,\boldsymbol{\hat{y}}+\boldsymbol{\nabla}{p}+\nu\,\boldsymbol{\nabla}\times\boldsymbol{\omega}=B_{0}\partial_{z}\boldsymbol{b},
\end{equation}
where
\begin{equation}
\boldsymbol{\omega}=\boldsymbol{\nabla}\times\boldsymbol{v},\quad\text{and}\quad\frac{D}{Dt}=\partial_{t}+{S}x\,\partial_{y}\end{equation}

We write the induction equation in terms of the $x$-component magnetic field,
\begin{equation}\label{eq:Bx}
\frac{Db_{x}}{Dt}+\eta(\partial_{y}j_{z}-\partial_{z}j_{y})=B_{0}\partial_{z}v_{x},
\end{equation}
and the $x$-component current density ($j_{x}=\partial_{y}b_{z}-\partial_{z}b_{y}$),
\begin{equation}\label{eq:Jx}
\frac{Dj_{x}}{Dt}-\eta\nabla^{2}j_{x}=B_{0}\partial_{z}\omega_{x}-S\partial_{z}b_{x}.
\end{equation}
We explicitly enforce divergence-free velocity and magnetic field,
\begin{equation}\label{eq:divu}
 \boldsymbol{\nabla}\cdot\boldsymbol{v}=\boldsymbol{\nabla}\cdot\boldsymbol{b}=0.
\end{equation}
The spatial domain is a doubly periodic channel in $y,\,z$ with width $-d/2\le{x}\le d/2$.
The boundary conditions are impenetrable stress-free and perfectly conducting; $v_{x}=\omega_{y}=\omega_{z}=b_{x}=\partial_{x}j_{x}=0$ at $x=\pm{d/2}$.

The main input parameters are the Coriolis parameter, $f=2 \Omega$; the background shear rate, $S=dV_{y}/dx<0$;  the vertical magnetic field $B_{0}$ (in Alfv\'{e}n units $\mu_{0}\rho_{0}=1$).
Accretion-disk modeling usually considers the Rossby number $q=-S/f<1$; $q=1$ corresponds to purely hydrodynamical Rayleigh unstable shear
\footnote{The astrophysical disk community often uses $q' = -S/\Omega$ to represent the same effective quantity; Keplerian profiles have $q' = 3/2$ in this case.}.
Unless otherwise stated $q=3/4$ (Keplerian).

The MRI is a weak-field instability; in the inviscid, ideal case the critical shear rate for instability in \textit{two dimensions} is
\begin{equation}\label{eq:Sc}
  S_{c}=-\frac{\pi^{2}B_{0}^2}{fd^2}.
\end{equation}
We use $\SSC$ as our instability control parameter.

We assume harmonic perturbations in $y$ and $z$, (e.g. pressure) $p=\hat{p}(x)e^{i(k_{y}y+k_{z}z)+\sigma{t}}$. 
We use a complex-valued growth rate $\sigma=\gamma+i\omega$ with $\gamma,\,\omega$ both real. 
The system reduces to a $\sigma$-eigenvalue problem of 10 first-order ODEs in $x$ with Dirichlet boundary conditions.
We pose and solve equations~(\ref{eq:mhd}-\ref{eq:divu}) using the \emph{Dedalus} framework \citep{2019arXiv190510388B}.

Our main interest is in ideal ($\eta=0$), inviscid ($\nu=0$) conditions.
However, we set $\eta=\nu=10^{-5}$ to avoid critical layers in the \textit{stable} solution branch.
We confirmed our results for unstable solutions are insensitive to small diffusion. 
For each $(k_{y},k_{z})$ pair, we solve the eigenvalue problem using $n_{x}=128$ modes; all our results are identical at double the resolution \footnote{see supplemental material for details}; see \protect\url{github.com/jsoishi/mri_prefers_3d} for all code used in this paper.
\begin{figure*}[ht]
  \includegraphics[width=\textwidth]{fig_1.pdf}
  \caption{Growth rates for three-dimensional MRI modes. 
  a) growth rates $\gamma$ over a grid of $k_{y}$ and $k_{z}$ for four values of supercriticality $\SSC$. 
  The black contours are equally spaced between zero and the maximum growth rate.
  The gray contour highlights $\gamma=0$; the red dots indicate the fastest growing mode. 
Two dimensional modes occur at $k_y = 0$ on the bottom of each figure.
  At $\SSC=0.640$, there are no two dimensional modes. 
  b) Growth rate vs $k_{y}$ for $k_{z}=0.259$ at $\SSC=1.002$ (highlighted by the blue line in a). 
  The orange line gives the asymptotic result from equation~(\ref{eq:asymp}) and the blue line are the numerical results. 
  The asymptotic form is valid as $\SSC\to1$ and $k_{y}\ll1$.) c) The full discrete spectrum of the MRI for $(k_{y},k_{z})\simeq(0.263,0.447)$. 
  The unstable mode is plotted in orange, its stable complex conjugate is blue, and all other modes are gray. 
  d) The phase angle, $\arctan(k_{y}/k_{z})$, as a function of $\SSC$ showing the exact transition to purely two-dimensional fastest modes ($\phi=0$) at $\SSC\gtrsim2.05$. }
  \label{fig:growth_rate}
\end{figure*}

Our two main results are: 
(I) When the first two-dimensional mode becomes unstable, there already exist three-dimensional modes with positive growth rate. 
(II) At sufficiently large criticality, the fastest-growing mode becomes purely two-dimensional. 
These results are pertinent in two ways. 
First, the MRI contains a substantial ``Goldilocks regime" with possible direct dynamo action, and this regime likely applies to stellar interiors and laboratory experiments. 
Second, our results accord with well-established results for accretion disks that expect two-dimensional primary modes.
Figure~\ref{fig:growth_rate}a shows the growth rates for four values of $\SSC$. 
At $\SSC=1.02$, the maximum growth rate ($\gamma_{\max}/|S|=0.093$) happens at $(k_{y},k_{z})\simeq(0.263,0.447)$.
Two-dimensional instability occurs for modes with finite growth rates at $k_y = 0$ along the bottom of the figure.
In all cases in figure~\ref{fig:growth_rate}a, the maximum growth rate does not occur along this line, indicating the dominance of three-dimensional modes.
The first panel of figure~\ref{fig:growth_rate}a shows $\SSC=0.640$, which contains only three-dimensional instability: the zero growth rate contour (highlighted in gray) does not intercept the $x$-axis.

The MRI's preference for three-dimensional modes can be predicted analytically.
We analyze equations~(\ref{eq:mhd}-\ref{eq:divu}) asymptotically close to $S_{c}$, and compute the leading-order correction to the growth rate from three-dimensional effects assuming $\SSC=1+\epsilon^{2}R$, $k_{z}\sim\mathcal{O}(\epsilon)$, $\sigma\sim{k_{y}}\sim\mathcal{O}(\epsilon^{2})$, 
\begin{equation}\label{eq:asymp}
\sigma^{2}=\frac{q\,B_{0}^{2}}{(q+1)}\left[ k_{z}^{2}\left(R-\frac{d^2}{\pi^{2}}k_{z}^{2}\right)+\Upsilon\,k_{y}^{2}\right]+\ldots,
\end{equation}
where
\begin{equation}
\Upsilon=\frac{\left(\pi^{2}+6\right) q+\pi^{2}-6}{12}\approx1.31\quad\text{for}\quad{q}=\frac{3}{4}.
\end{equation}
The first term in equation~(\ref{eq:asymp}) results from the two-dimensional calculation.
The second term is positive definite: it \emph{always} leads to enhanced growth rates when $k_{y}\neq0$ and ultraviolet divergence in the absence of higher-order effects.

Figure~\ref{fig:growth_rate}b compares the numerical growth rates for $\SSC=1.002$ between $0\le{k_{y}}\le0.2$ to the asymptotic approximation, showing good agreement where the latter is valid.
Figure~\ref{fig:growth_rate}c shows the full spectrum for $\SSC=1.02$.
The plot shows a purely growing/decaying complex-conjugate pair (orange/blue dots on the real axis) consistent with equation~(\ref{eq:asymp}).
The other stable modes (gray dots) are rotationally modified Alfv\'{e}n waves found in left- and right-going pairs, consistent with the analytic predictions for two-dimensional stability calculations.
We reject spurious eigenvalues by using \texttt{eigentools} \footnote{\protect\url{https://bitbucket.org/jsoishi/eigentools}} to solve equations~(\ref{eq:mhd}-\ref{eq:divu}) at two resolutions and retain only pairs equal to within one part in $10^{-6}$.

\begin{figure}[h!]
  \centering
  \includegraphics[width=\columnwidth]{eigvecs_xy_run_11_ideal_single_mode.pdf}
  \caption{\textit{Ideal} eigenvectors of the velocity and magnetic field perturbations for the most unstable mode when $\SSC=1.02$, and $\eta=\nu=0$; calculations at $\eta=\nu=10^{-5}$ are indistinguishable.
  The top row shows $v_{x}$, $v_{y}$, $v_{z}$ (red is negative, blue positive); the bottom row shows $b_{x}$, $b_{y}$, $b_{z}$ (purple is negative, orange positive). 
The amplitudes are arbitrary because of linearity.}
  \label{fig:eigvec}
\end{figure}

Numerical simulations of the MRI in accretion disks consistently show axisymmetric modes dominating the early evolution of the MRI before breaking down into 3D MHD turbulence \citep{1995ApJ...440..742H,2018ApJ...853..174H,2019ApJS..241...26D}. 
These ``channel modes'' are exact non-linear solutions for $x$-shearing-periodic domains, and can only saturate via parasitic shear instabilities \citep{1994ApJ...432..213G}.
Impenetrable and stress-free boundary conditions are more applicable to stellar interiors. 
But even in disks, any kind of finite radial extent will cutoff the unbounded growth of channel modes. 
We reconcile our results with earlier simulations by showing that the MRI indeed prefers two dimensions for the larger values of criticality found in disk simulations. 
Figure~\ref{fig:growth_rate}d shows the phase angle $\phi=\arctan(k_{y}/k_{z})$ of the fastest growing mode as function of $\SSC$.
The overall critical shear for three-dimensional modes is $S_{c,\text{3D}}\simeq0.102S_c$.
Above $\SSC\gtrsim2.05$, $\phi$ becomes zero, indicating that axisymmetric modes have the fastest growth rates.
For the fiducial run in \citet{1996ApJ...464..690H}, $\SSC\simeq4.84$ (most works use similar values).
Thus, our theory predicts axisymmetric modes should dominate the linear dynamics for parameters studied in prior numerical simulations.

\begin{figure}[h!]
  \includegraphics[width=\columnwidth]{{mean_field_alpha_run_11_output.h5}.pdf}
  \caption{Velocity-magnetic field correlations relevant to the mean electromotive force for the eigenvectors of the fastest growing mode at $\SSC = 1.02$.}
  \label{fig:correlation}
\end{figure}

Even for $\SSC> 2.05$ there are significant swaths of unstable three-dimensional modes with growth rates comparable to the maximum (figure~\ref{fig:growth_rate}a).
Also, non-normality generically accompanies non-axisymmetry in shearing systems \citep[see][]{1992MNRAS.255P..25K}.
In non-normal systems, transient amplification can occur for stable modes, and can even cause turbulence.
Therefore significant three dimensionality likely implies important non-normal behavior near onset.

Figure~\ref{fig:eigvec} shows the eigenvectors for the most unstable mode at $\SSC=1.02$ using ideal MHD. 
No critical layers can form for $\gamma\ne0$.
We therefore solve equations~(\ref{eq:mhd}--\ref{eq:divu}) without dissipation as a 2nd-order system in $\partial_{x}$ and impose $v_{x}=0$ at $x=\pm{d/2}$.
The eigenfunctions are indistinguishable from those at finite dissipation, lending additional confidence to our other results.
The tilted structures of $v_{x}$ and $v_{y}$ as well as $b_{x}$ and $b_{y}$, imply non-trivial Reynolds and Maxwell stresses.

We should expect the generation of a mean electromotive force (EMF) in the linear regime, \textit{i.e}
$\left<\boldsymbol{v}\times\boldsymbol{b}\right>\sim\,\alpha\!\left<\boldsymbol{B}_{0}\right>$.
This is significant because of the possibility of direct laminar dynamo action in regions of weak shear at large Reynolds number.
Figure~\ref{fig:correlation} shows the correlation (averaged over $y,\,z$) as a function of $x$ over the domain for $\SSC = 1.02$,
\begin{equation}
\text{EMF}\propto\frac{\left<\boldsymbol{\hat{z}}\cdot(\boldsymbol{v}\times\boldsymbol{b})\right>_{\!y,z}}{\boldsymbol{v}_{\text{rms}}\boldsymbol{b}_{\text{rms}}},
\end{equation}
where ``rms'' is denotes the average over the whole domain. 
The correlation is independent of the arbitrary normalization of the linear eigenvectors.
The non-zero correlation shows the tendency for back reaction on the mean magnetic field.

MRI dynamos have been studied in a number of different contexts \citep{2007PhRvL..98y4502R,2011ApJ...740...18O,2015PhRvL.114h5002S}, but all except one focused on turbulent dynamos. 
The lone exception, \citet{2016MNRAS.462..818B}, numerically found the exponential growth of mean magnetic fields during the linear growth phase of the MRI far from stability.
Our work explains this result in terms of purely linear dynamics:
non-axisymmetric MRI unstable modes drive the dynamo growth of magnetic fields.

\begin{figure}[h!]
  \centering
  \includegraphics[width=\columnwidth]{low_rossby_liquid_metal_growth_rates.pdf}
  \caption{(a) Growth rates for $q=1/10$ relevant to the stellar interiors. 
  (b) Growth rates for MRI near onset at liquid metal-like diffusivities, $\nu=10^{-6}$, $\eta=0.1$.
The lack of two dimensional instability is evident by the fact that the zero contour does not intercept the x-axis.}
  \label{fig:other_params}
\end{figure}
Finally, we vary two important parameters:
$q$ and the magnetic Prandtl number $\Prm=\nu/\eta$.
Figure~\ref{fig:other_params} shows the growth rates for $\SSC=1.02$.
The upper panel shows $q=1/10$ (addressing a possible range within stellar interiors) holding all other parameters equal to their fiducial values.
The growth rates show similar behavior to the $q=3/4$ cases, \textit{i.e.}, dominated by three-dimensional modes with $\gamma_{\max}/|S|\simeq0.1$.
The lower panel shows $\nu=10^{-6}$ and $\eta=0.1$ with all other parameters equal to their fiducial values. 
This case is relevant to liquid metals e.g., the Princeton MRI experiment \citep{2002JFM...462..365G}
at $\Prm=10^{-5}$. 
Our results are for the rotating plane Couette geometry (Taylor-Couette in the small-gap limit $R_{2}\approx{R}_{1}$) with highly idealized boundary conditions.
It is nevertheless quite interesting that, near onset, the low-$\Prm$ MRI is \emph{only} unstable to three-dimensional modes.

Our results show that three-dimensional modes grow faster than two-dimensional modes whenever the MRI is near its critical shear values.
While we mainly focus on the ideal Keplerian case ($q=3/4$), we also demonstrate robustness for low Rossby ($q=1/10$) and Prandtl numbers ($\Prm=10^{-5}$).
There are several important future directions this work suggests.
First, the Sun possesses two internal shear layers with inwardly increasing shear, the high-latitude tachocline, and the near-surface shear layer (NSSL). 
Past work has already pointed out the possibility of the small-scale MRI in the Sun using local analysis \citep{2007ApJ...667L.207P,2011MNRAS.411L..26M,2014ApJ...787...21K}.
The NSSL, in particular, may also host slower MRI-driven dynamics; despite containing small-scale convection.
It is therefore crucial to further elucidate the nonlinear saturation of the 3D MRI, along with its robustness to convection.
Second, our low-$\Prm$ results suggest that three-dimensional MRI modes may be the easiest to excite in liquid metal experiments.
Determining possible non-axisymmetric signatures in Taylor-Couette experiments requires followup work using more realistic boundary conditions and geometry. 
Our prior work on the axisymmetric MRI \citep{2017ApJ...841....1C,2017ApJ...841....2C} shows only small differences between rotating plane and cylindrical Taylor-Couette geometries. 
We fully expect the general three-dimensional features of the MRI to persist in more complex applications.

\begin{acknowledgments}
We thank Steve Tobias for helpful conversations on this work.
Oishi, Baxter, and Brown acknowledge support from NASA LWS grant No. NNX16AC92G.
Oishi also acknowledges support from Research Corporation Scialog Collaborative Award (TDA) ID\#24231. 
Computations were performed on the \emph{Leavitt} cluster at the Bates College High Performance Computing Center.

\end{acknowledgments}

%merlin.mbs apsrev4-1.bst 2010-07-25 4.21a (PWD, AO, DPC) hacked
%Control: key (0)
%Control: author (8) initials jnrlst
%Control: editor formatted (1) identically to author
%Control: production of article title (-1) disabled
%Control: page (0) single
%Control: year (1) truncated
%Control: production of eprint (0) enabled
\begin{thebibliography}{25}%
\makeatletter
\providecommand \@ifxundefined [1]{%
 \@ifx{#1\undefined}
}%
\providecommand \@ifnum [1]{%
 \ifnum #1\expandafter \@firstoftwo
 \else \expandafter \@secondoftwo
 \fi
}%
\providecommand \@ifx [1]{%
 \ifx #1\expandafter \@firstoftwo
 \else \expandafter \@secondoftwo
 \fi
}%
\providecommand \natexlab [1]{#1}%
\providecommand \enquote  [1]{``#1''}%
\providecommand \bibnamefont  [1]{#1}%
\providecommand \bibfnamefont [1]{#1}%
\providecommand \citenamefont [1]{#1}%
\providecommand \href@noop [0]{\@secondoftwo}%
\providecommand \href [0]{\begingroup \@sanitize@url \@href}%
\providecommand \@href[1]{\@@startlink{#1}\@@href}%
\providecommand \@@href[1]{\endgroup#1\@@endlink}%
\providecommand \@sanitize@url [0]{\catcode `\\12\catcode `\$12\catcode
  `\&12\catcode `\#12\catcode `\^12\catcode `\_12\catcode `\%12\relax}%
\providecommand \@@startlink[1]{}%
\providecommand \@@endlink[0]{}%
\providecommand \url  [0]{\begingroup\@sanitize@url \@url }%
\providecommand \@url [1]{\endgroup\@href {#1}{\urlprefix }}%
\providecommand \urlprefix  [0]{URL }%
\providecommand \Eprint [0]{\href }%
\providecommand \doibase [0]{http://dx.doi.org/}%
\providecommand \selectlanguage [0]{\@gobble}%
\providecommand \bibinfo  [0]{\@secondoftwo}%
\providecommand \bibfield  [0]{\@secondoftwo}%
\providecommand \translation [1]{[#1]}%
\providecommand \BibitemOpen [0]{}%
\providecommand \bibitemStop [0]{}%
\providecommand \bibitemNoStop [0]{.\EOS\space}%
\providecommand \EOS [0]{\spacefactor3000\relax}%
\providecommand \BibitemShut  [1]{\csname bibitem#1\endcsname}%
\let\auto@bib@innerbib\@empty
%</preamble>
\bibitem [{\citenamefont {{Balbus}}\ and\ \citenamefont
  {{Hawley}}(1998)}]{1998RvMP...70....1B}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.~A.}\ \bibnamefont
  {{Balbus}}}\ and\ \bibinfo {author} {\bibfnamefont {J.~F.}\ \bibnamefont
  {{Hawley}}},\ }\href {\doibase 10.1103/RevModPhys.70.1} {\bibfield  {journal}
  {\bibinfo  {journal} {Reviews of Modern Physics}\ }\textbf {\bibinfo {volume}
  {70}},\ \bibinfo {pages} {1} (\bibinfo {year} {1998})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {{Julien}}\ and\ \citenamefont
  {{Knobloch}}(2010)}]{2010RSPTA.368.1607J}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {K.}~\bibnamefont
  {{Julien}}}\ and\ \bibinfo {author} {\bibfnamefont {E.}~\bibnamefont
  {{Knobloch}}},\ }\href {\doibase 10.1098/rsta.2009.0251} {\bibfield
  {journal} {\bibinfo  {journal} {Philosophical Transactions of the Royal
  Society of London Series A}\ }\textbf {\bibinfo {volume} {368}},\ \bibinfo
  {pages} {1607} (\bibinfo {year} {2010})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {{Johansen}}\ \emph {et~al.}(2007)\citenamefont
  {{Johansen}}, \citenamefont {{Oishi}}, \citenamefont {{Mac Low}},
  \citenamefont {{Klahr}}, \citenamefont {{Henning}},\ and\ \citenamefont
  {{Youdin}}}]{2007Natur.448.1022J}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
  {{Johansen}}}, \bibinfo {author} {\bibfnamefont {J.~S.}\ \bibnamefont
  {{Oishi}}}, \bibinfo {author} {\bibfnamefont {M.-M.}\ \bibnamefont {{Mac
  Low}}}, \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {{Klahr}}},
  \bibinfo {author} {\bibfnamefont {T.}~\bibnamefont {{Henning}}}, \ and\
  \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont {{Youdin}}},\ }\href
  {\doibase 10.1038/nature06086} {\bibfield  {journal} {\bibinfo  {journal}
  {\nat}\ }\textbf {\bibinfo {volume} {448}},\ \bibinfo {pages} {1022}
  (\bibinfo {year} {2007})},\ \Eprint {http://arxiv.org/abs/0708.3890}
  {arXiv:0708.3890 [astro-ph]} \BibitemShut {NoStop}%
\bibitem [{\citenamefont {{Xie}}\ \emph {et~al.}(2018)\citenamefont {{Xie}},
  \citenamefont {{Julien}},\ and\ \citenamefont
  {{Knobloch}}}]{2018MNRAS.474.3451X}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.-H.}\ \bibnamefont
  {{Xie}}}, \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont {{Julien}}}, \
  and\ \bibinfo {author} {\bibfnamefont {E.}~\bibnamefont {{Knobloch}}},\
  }\href {\doibase 10.1093/mnras/stx2977} {\bibfield  {journal} {\bibinfo
  {journal} {\mnras}\ }\textbf {\bibinfo {volume} {474}},\ \bibinfo {pages}
  {3451} (\bibinfo {year} {2018})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {{Vasil}}(2015)}]{2015RSPSA.47140699V}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {G.~M.}\ \bibnamefont
  {{Vasil}}},\ }\href {\doibase 10.1098/rspa.2014.0699} {\bibfield  {journal}
  {\bibinfo  {journal} {Proceedings of the Royal Society of London Series A}\
  }\textbf {\bibinfo {volume} {471}},\ \bibinfo {pages} {20140699} (\bibinfo
  {year} {2015})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {{Clark}}\ and\ \citenamefont
  {{Oishi}}(2017{\natexlab{a}})}]{2017ApJ...841....1C}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.~E.}\ \bibnamefont
  {{Clark}}}\ and\ \bibinfo {author} {\bibfnamefont {J.~S.}\ \bibnamefont
  {{Oishi}}},\ }\href {\doibase 10.3847/1538-4357/aa6ff1} {\bibfield  {journal}
  {\bibinfo  {journal} {\apj}\ }\textbf {\bibinfo {volume} {841}},\ \bibinfo
  {eid} {1} (\bibinfo {year} {2017}{\natexlab{a}})},\ \Eprint
  {http://arxiv.org/abs/1610.01616} {arXiv:1610.01616 [astro-ph.HE]}
  \BibitemShut {NoStop}%
\bibitem [{\citenamefont {{Clark}}\ and\ \citenamefont
  {{Oishi}}(2017{\natexlab{b}})}]{2017ApJ...841....2C}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {S.~E.}\ \bibnamefont
  {{Clark}}}\ and\ \bibinfo {author} {\bibfnamefont {J.~S.}\ \bibnamefont
  {{Oishi}}},\ }\href {\doibase 10.3847/1538-4357/aa6ff6} {\bibfield  {journal}
  {\bibinfo  {journal} {\apj}\ }\textbf {\bibinfo {volume} {841}},\ \bibinfo
  {eid} {2} (\bibinfo {year} {2017}{\natexlab{b}})},\ \Eprint
  {http://arxiv.org/abs/1610.01603} {arXiv:1610.01603 [astro-ph.HE]}
  \BibitemShut {NoStop}%
\bibitem [{Note1()}]{Note1}%
  \BibitemOpen
  \bibinfo {note} {The astrophysical disk community often uses $q' = -S/\Omega
  $ to represent the same effective quantity; Keplerian profiles have $q' =
  3/2$ in this case.}\BibitemShut {Stop}%
\bibitem [{\citenamefont {{Burns}}\ \emph {et~al.}(2019)\citenamefont
  {{Burns}}, \citenamefont {{Vasil}}, \citenamefont {{Oishi}}, \citenamefont
  {{Lecoanet}},\ and\ \citenamefont {{Brown}}}]{2019arXiv190510388B}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {K.~J.}\ \bibnamefont
  {{Burns}}}, \bibinfo {author} {\bibfnamefont {G.~M.}\ \bibnamefont
  {{Vasil}}}, \bibinfo {author} {\bibfnamefont {J.~S.}\ \bibnamefont
  {{Oishi}}}, \bibinfo {author} {\bibfnamefont {D.}~\bibnamefont {{Lecoanet}}},
  \ and\ \bibinfo {author} {\bibfnamefont {B.~P.}\ \bibnamefont {{Brown}}},\
  }\href@noop {} {\bibfield  {journal} {\bibinfo  {journal} {Physical Review
  Research, submitted}\ ,\ \bibinfo {eid} {arXiv:1905.10388}} (\bibinfo {year}
  {2019})},\ \Eprint {http://arxiv.org/abs/1905.10388} {arXiv:1905.10388
  [astro-ph.IM]} \BibitemShut {NoStop}%
\bibitem [{Note2()}]{Note2}%
  \BibitemOpen
  \bibinfo {note} {See supplemental material for details}\BibitemShut {NoStop}%
\bibitem [{Note3()}]{Note3}%
  \BibitemOpen
  \bibinfo {note} {Available at \protect \url
  {https://bitbucket.org/jsoishi/eigentools}}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {{Hawley}}\ \emph {et~al.}(1995)\citenamefont
  {{Hawley}}, \citenamefont {{Gammie}},\ and\ \citenamefont
  {{Balbus}}}]{1995ApJ...440..742H}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~F.}\ \bibnamefont
  {{Hawley}}}, \bibinfo {author} {\bibfnamefont {C.~F.}\ \bibnamefont
  {{Gammie}}}, \ and\ \bibinfo {author} {\bibfnamefont {S.~A.}\ \bibnamefont
  {{Balbus}}},\ }\href {\doibase 10.1086/175311} {\bibfield  {journal}
  {\bibinfo  {journal} {\apj}\ }\textbf {\bibinfo {volume} {440}},\ \bibinfo
  {pages} {742} (\bibinfo {year} {1995})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {{Hirai}}\ \emph {et~al.}(2018)\citenamefont
  {{Hirai}}, \citenamefont {{Katoh}}, \citenamefont {{Terada}},\ and\
  \citenamefont {{Kawai}}}]{2018ApJ...853..174H}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {K.}~\bibnamefont
  {{Hirai}}}, \bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont {{Katoh}}},
  \bibinfo {author} {\bibfnamefont {N.}~\bibnamefont {{Terada}}}, \ and\
  \bibinfo {author} {\bibfnamefont {S.}~\bibnamefont {{Kawai}}},\ }\href
  {\doibase 10.3847/1538-4357/aaa5b2} {\bibfield  {journal} {\bibinfo
  {journal} {\apj}\ }\textbf {\bibinfo {volume} {853}},\ \bibinfo {eid} {174}
  (\bibinfo {year} {2018})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {{Deng}}\ \emph {et~al.}(2019)\citenamefont {{Deng}},
  \citenamefont {{Mayer}}, \citenamefont {{Latter}}, \citenamefont
  {{Hopkins}},\ and\ \citenamefont {{Bai}}}]{2019ApJS..241...26D}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {H.}~\bibnamefont
  {{Deng}}}, \bibinfo {author} {\bibfnamefont {L.}~\bibnamefont {{Mayer}}},
  \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont {{Latter}}}, \bibinfo
  {author} {\bibfnamefont {P.~F.}\ \bibnamefont {{Hopkins}}}, \ and\ \bibinfo
  {author} {\bibfnamefont {X.-N.}\ \bibnamefont {{Bai}}},\ }\href {\doibase
  10.3847/1538-4365/ab0957} {\bibfield  {journal} {\bibinfo  {journal} {\apjs}\
  }\textbf {\bibinfo {volume} {241}},\ \bibinfo {eid} {26} (\bibinfo {year}
  {2019})},\ \Eprint {http://arxiv.org/abs/1901.05190} {arXiv:1901.05190
  [astro-ph.EP]} \BibitemShut {NoStop}%
\bibitem [{\citenamefont {{Goodman}}\ and\ \citenamefont
  {{Xu}}(1994)}]{1994ApJ...432..213G}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
  {{Goodman}}}\ and\ \bibinfo {author} {\bibfnamefont {G.}~\bibnamefont
  {{Xu}}},\ }\href {\doibase 10.1086/174562} {\bibfield  {journal} {\bibinfo
  {journal} {\apj}\ }\textbf {\bibinfo {volume} {432}},\ \bibinfo {pages} {213}
  (\bibinfo {year} {1994})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {{Hawley}}\ \emph {et~al.}(1996)\citenamefont
  {{Hawley}}, \citenamefont {{Gammie}},\ and\ \citenamefont
  {{Balbus}}}]{1996ApJ...464..690H}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~F.}\ \bibnamefont
  {{Hawley}}}, \bibinfo {author} {\bibfnamefont {C.~F.}\ \bibnamefont
  {{Gammie}}}, \ and\ \bibinfo {author} {\bibfnamefont {S.~A.}\ \bibnamefont
  {{Balbus}}},\ }\href {\doibase 10.1086/177356} {\bibfield  {journal}
  {\bibinfo  {journal} {\apj}\ }\textbf {\bibinfo {volume} {464}},\ \bibinfo
  {pages} {690} (\bibinfo {year} {1996})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {{Knobloch}}(1992)}]{1992MNRAS.255P..25K}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {E.}~\bibnamefont
  {{Knobloch}}},\ }\href {\doibase 10.1093/mnras/255.1.25P} {\bibfield
  {journal} {\bibinfo  {journal} {\mnras}\ }\textbf {\bibinfo {volume} {255}},\
  \bibinfo {pages} {25P} (\bibinfo {year} {1992})}\BibitemShut {NoStop}%
\bibitem [{\citenamefont {{Rincon}}\ \emph {et~al.}(2007)\citenamefont
  {{Rincon}}, \citenamefont {{Ogilvie}},\ and\ \citenamefont
  {{Proctor}}}]{2007PhRvL..98y4502R}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {F.}~\bibnamefont
  {{Rincon}}}, \bibinfo {author} {\bibfnamefont {G.~I.}\ \bibnamefont
  {{Ogilvie}}}, \ and\ \bibinfo {author} {\bibfnamefont {M.~R.~E.}\
  \bibnamefont {{Proctor}}},\ }\href {\doibase 10.1103/PhysRevLett.98.254502}
  {\bibfield  {journal} {\bibinfo  {journal} {\prl}\ }\textbf {\bibinfo
  {volume} {98}},\ \bibinfo {eid} {254502} (\bibinfo {year} {2007})},\ \Eprint
  {http://arxiv.org/abs/0705.2814} {arXiv:0705.2814 [astro-ph]} \BibitemShut
  {NoStop}%
\bibitem [{\citenamefont {{Oishi}}\ and\ \citenamefont {{Mac
  Low}}(2011)}]{2011ApJ...740...18O}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.~S.}\ \bibnamefont
  {{Oishi}}}\ and\ \bibinfo {author} {\bibfnamefont {M.-M.}\ \bibnamefont {{Mac
  Low}}},\ }\href {\doibase 10.1088/0004-637X/740/1/18} {\bibfield  {journal}
  {\bibinfo  {journal} {\apj}\ }\textbf {\bibinfo {volume} {740}},\ \bibinfo
  {eid} {18} (\bibinfo {year} {2011})},\ \Eprint
  {http://arxiv.org/abs/1102.5093} {arXiv:1102.5093 [astro-ph.SR]} \BibitemShut
  {NoStop}%
\bibitem [{\citenamefont {{Squire}}\ and\ \citenamefont
  {{Bhattacharjee}}(2015)}]{2015PhRvL.114h5002S}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
  {{Squire}}}\ and\ \bibinfo {author} {\bibfnamefont {A.}~\bibnamefont
  {{Bhattacharjee}}},\ }\href {\doibase 10.1103/PhysRevLett.114.085002}
  {\bibfield  {journal} {\bibinfo  {journal} {\prl}\ }\textbf {\bibinfo
  {volume} {114}},\ \bibinfo {eid} {085002} (\bibinfo {year} {2015})},\ \Eprint
  {http://arxiv.org/abs/1408.6793} {arXiv:1408.6793 [astro-ph.HE]} \BibitemShut
  {NoStop}%
\bibitem [{\citenamefont {{Bhat}}\ \emph {et~al.}(2016)\citenamefont {{Bhat}},
  \citenamefont {{Ebrahimi}},\ and\ \citenamefont
  {{Blackman}}}]{2016MNRAS.462..818B}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {P.}~\bibnamefont
  {{Bhat}}}, \bibinfo {author} {\bibfnamefont {F.}~\bibnamefont {{Ebrahimi}}},
  \ and\ \bibinfo {author} {\bibfnamefont {E.~G.}\ \bibnamefont {{Blackman}}},\
  }\href {\doibase 10.1093/mnras/stw1619} {\bibfield  {journal} {\bibinfo
  {journal} {\mnras}\ }\textbf {\bibinfo {volume} {462}},\ \bibinfo {pages}
  {818} (\bibinfo {year} {2016})},\ \Eprint {http://arxiv.org/abs/1605.02433}
  {arXiv:1605.02433 [astro-ph.HE]} \BibitemShut {NoStop}%
\bibitem [{\citenamefont {{Goodman}}\ and\ \citenamefont
  {{Ji}}(2002)}]{2002JFM...462..365G}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {J.}~\bibnamefont
  {{Goodman}}}\ and\ \bibinfo {author} {\bibfnamefont {H.}~\bibnamefont
  {{Ji}}},\ }\href {\doibase 10.1017/S0022112002008704} {\bibfield  {journal}
  {\bibinfo  {journal} {Journal of Fluid Mechanics}\ }\textbf {\bibinfo
  {volume} {462}},\ \bibinfo {pages} {365} (\bibinfo {year} {2002})},\ \Eprint
  {http://arxiv.org/abs/astro-ph/0104206} {arXiv:astro-ph/0104206 [astro-ph]}
  \BibitemShut {NoStop}%
\bibitem [{\citenamefont {{Parfrey}}\ and\ \citenamefont
  {{Menou}}(2007)}]{2007ApJ...667L.207P}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {K.~P.}\ \bibnamefont
  {{Parfrey}}}\ and\ \bibinfo {author} {\bibfnamefont {K.}~\bibnamefont
  {{Menou}}},\ }\href {\doibase 10.1086/522426} {\bibfield  {journal} {\bibinfo
   {journal} {\apjl}\ }\textbf {\bibinfo {volume} {667}},\ \bibinfo {pages}
  {L207} (\bibinfo {year} {2007})},\ \Eprint {http://arxiv.org/abs/0708.3675}
  {arXiv:0708.3675} \BibitemShut {NoStop}%
\bibitem [{\citenamefont {{Masada}}(2011)}]{2011MNRAS.411L..26M}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {Y.}~\bibnamefont
  {{Masada}}},\ }\href {\doibase 10.1111/j.1745-3933.2010.00987.x} {\bibfield
  {journal} {\bibinfo  {journal} {Monthly Notices of the Royal Astronomical
  Society}\ }\textbf {\bibinfo {volume} {411}},\ \bibinfo {pages} {L26}
  (\bibinfo {year} {2011})},\ \Eprint {http://arxiv.org/abs/1011.2681}
  {arXiv:1011.2681 [astro-ph.SR]} \BibitemShut {NoStop}%
\bibitem [{\citenamefont {{Kagan}}\ and\ \citenamefont
  {{Wheeler}}(2014)}]{2014ApJ...787...21K}%
  \BibitemOpen
  \bibfield  {author} {\bibinfo {author} {\bibfnamefont {D.}~\bibnamefont
  {{Kagan}}}\ and\ \bibinfo {author} {\bibfnamefont {J.~C.}\ \bibnamefont
  {{Wheeler}}},\ }\href {\doibase 10.1088/0004-637X/787/1/21} {\bibfield
  {journal} {\bibinfo  {journal} {\apj}\ }\textbf {\bibinfo {volume} {787}},\
  \bibinfo {eid} {21} (\bibinfo {year} {2014})},\ \Eprint
  {http://arxiv.org/abs/1407.4654} {arXiv:1407.4654} \BibitemShut {NoStop}%
\end{thebibliography}%

\end{document}

%  LocalWords:  Magnetorotational Oishi Lewiston Vasil NSW Lecoanet
%  LocalWords:  magnetorotational Keplerian magnetofluid Prandtl MHD
%  LocalWords:  dimensionality Couette axisymmetric linearized RSPSA
%  LocalWords:  magnetohydrodynamic Navier Dt Dj dV dx Alfv Rossby fd
%  LocalWords:  hydrodynamical ODEs Dedalus arXiv supercriticality nd
%  LocalWords:  criticality eigenvectors fiducial axisymmetry NSSL Sx
%  LocalWords:  eigenfunctions eigenmodes diffusivities tachocline
%  LocalWords:  Tobias JSO LWS NNX Scialog TDA Leavitt mri DPMMS rms
%  LocalWords:  eigentools

### 2021-07-14 paper/supplemental.tex
initial release

\documentclass[aps,prl,preprint
,superscriptaddress]{revtex4-1}
\newcommand{\SSC}{S/S_{c}}
\newcommand{\mnras}{Monthly Notices of the Royal Astronomical Society}
\newcommand{\apjs}{The Astrophysical Journal Supplement Series}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}

% You should use BibTeX and apsrev.bst for references
% Choosing a journal automatically selects the correct APS
% BibTeX style file (bst file), so only uncomment the line
% below if necessary.
%\bibliographystyle{apsrev4-1}

\newcommand\Beq{\begin{eqnarray}} 
\newcommand\Eeq{\end{eqnarray}}

\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Eqs}[2]{Eqs.~(\ref{#1})--(\ref{#2})}
\newcommand{\eq}[1]{eq.~(\ref{#1})}
\newcommand{\eqs}[2]{eqs.~(\ref{#1})--(\ref{#2})}

\newcommand{\Prob}[2]{\mathbb{P}[\, #1 \, | \, #2 \, ] }
\newcommand{\Pro}[1]{\mathbb{P}[\, #1 \, ] }

\newcommand{\inv}[1]{\frac{1}{#1}}
\newcommand{\pd}[1]{\partial_{#1}}

\newcommand{\Reyn}{\mathrm{Re}}
\newcommand{\Reym}{\mathrm{Rm}}
\newcommand{\Ra}{R}
\newcommand{\Ve}{\mbox{\textit{Ve}}}
\newcommand{\eps}{\varepsilon}
\newcommand{\del}{\delta}
\newcommand{\e}[1]{\eps^{#1}}
\newcommand{\f}{2\Omega}

\newcommand{\n}{\hat{n}}

\newcommand{\dat}{\cdot}
\newcommand{\dd}[1]{\,\mathrm{d}{#1}}
\newcommand{\cc}{\mathrm{c.c.}}
\renewcommand{\vec}[1]{\boldsymbol{#1}}
\newcommand{\grad}{\nabla}
\renewcommand{\div}{\nabla \dat}
\newcommand{\Tr}{\text{Tr}}

\newcommand{\pren}[1]{ \left(#1\right) }
\newcommand{\brac}[1]{\left[\,#1\,\right]}

\newcommand{\ave}[1]{\big<#1\big>}

\renewcommand{\l}{\ell}

\newcommand{\txt}[1]{\quad \text{#1} \quad }

\begin{document}

% Use the \preprint command to place your local institutional report
% number in the upper righthand corner of the title page in preprint mode.
% Multiple \preprint commands are allowed.
% Use the 'preprintnumbers' class option to override journal defaults
% to display numbers if necessary
%\preprint{}

%Title of paper
\title{Supplemental Information for: The Magnetorotational Instability Prefers Three Dimensions}

\maketitle
\section{Numerical Methods and Resolution Studies}
\label{sec:methods}

For both the ideal and non-ideal MHD equations, we solve the eigenvalue problem in the $x$ direction using the \texttt{EigenValueProblem} solver in \emph{Dedalus} for a grid of $N_y \times N_z$ modes in the $y$ and $z$ directions.
For most of our runs, we use a targeted, sparse eigenvalue solver to find the 15 modes closest to a guess for the maximum growth rate.
In run 3, we have confirmed that dense solvers retrieve identical results.
For nearly ideal runs, we use the ideal 2D growth rate as input for the smallest $k_y > 0$ mode at each $k_z$ and then use the output from each previous $k_y$ as an input guess for the next mode.
For those runs that have significant $\eta$, we instead use a dense solve for the $k_y = 0$ modes and then step forward in $k_y$ at each $k_z$ as before.
Our solver is embarrassingly parallelized over the $k_z$ modes.
For the spectrum in figure 1 of the main text, we used a dense eigenvalue solver at $(k_y^{max},k_z^{max})$ for $\SSC= 1.02$.

In order to ensure our results are converged, we have repeated runs at $n_x=256$ Chebyshev modes as well as doubling the number of modes in $y$ and $z$.

\begin{tabular}{cccccccccccccc}
\textbf{Run} & \textbf{$\SSC$} & \textbf{$\nu$} & \textbf{$\eta$} & \textbf{$q$} & \textbf{$N_x$} & \textbf{$N_{k_y}$} & \textbf{$N_{k_z}$} & \textbf{Sparse/Dense}& \\
1  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & resolution study\\
2  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 256 & 200 & 200 & sparse & \\
3  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & dense  & \\
4  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 512 & 512 & sparse & \\
5  &   1.02 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 100 & 100 & sparse & \\
6  &   0.2  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & $\SSC$ variation\\
7  &   0.3  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
8  &   0.4  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
9  &   0.5  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
10 &   0.64 & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
11 & 1.020  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
12 & 1.44   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
13 & 1.75   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
14 & 1.891  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
15 & 2.     & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
16 & 2.01   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
17 & 2.015  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
18 & 2.031  & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
19 & 2.05   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
20 & 2.1    & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
21 & 2.25   & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
22 & 2.5    & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
23 & 4      & $10^{-5}$ & $10^{-5}$ & 0.75 & 128 & 200 & 200 & sparse & \\
24 & 1.02   & $10^{-6}$ & $10^{-6}$ & 0.75 & 128 & 200 & 200 & sparse & Reynolds number study\\
25 & 1.02   & $10^{-4}$ & $10^{-4}$ & 0.75 & 128 & 200 & 200 & sparse & \\
26 & 1.02   & $10^{-3}$ & $10^{-3}$ & 0.75 & 128 & 200 & 200 & sparse & \\
27 & 1.02   & $10^{-2}$ & $10^{-2}$ & 0.75 & 128 & 200 & 200 & sparse & \\
28 & 1.02   & $10^{-5}$ & $10^{-5}$ & 0.1  & 128 & 200 & 200 & sparse & Low Rossby\\ 
29 & 1.02   & $10^{-6}$ & $10^{-2}$ & 0.1  & 128 & 200 & 200 & sparse & Liquid metal case\\ 
\end{tabular}
%\label{simulations}

\section{1st-order formulation of the dissipative equations}
\label{sec:formulation}

We formulate the system in terms of the 10 problem \texttt{variables} 
\Beq
X \ = \ \left[\,p,v_{x},v_{y},v_{z},\omega_{y},\omega_{z},b_{x},b_{y},b_{z},j_{xx}\,\right]^{\top} \nonumber
\Eeq
We use string-based \texttt{substitutions} to expedite the input of the full the system of equations: 
\Beq
``\omega_{x}" &=& ``\pd{y} v_{z} - \pd{z} v_{y}" \nonumber \\
``j_{x}" &=& ``\pd{y} b_{z} - \pd{z} b_{y}" \nonumber \\
``j_{y}" &=& ``\pd{z} b_{x} - \pd{x} b_{z}" \nonumber \\
``j_{z}" &=& ``\pd{x} b_{y} - \pd{y} b_{x}" \nonumber 
\Eeq
The expression for $\omega_{x},\,j_{x},\,j_{y},\,j_{z}$ are not variables that \textit{Dedalus} solves for. 
They are short-hand aliases that are directly replaced by the expressions on the right-hand side of the definition, which are themselves defined strictly in terms of problem variables.

We also define the following operator substitutions (with general argument $A$)
\Beq
``\pd{y} A" &=& ``i k_{y} A" \nonumber \\
``\pd{z} A" &=& ``i k_{z} A" \nonumber \\
``D_{t} A" &=& ``\sigma A +S x \pd{y} A " \nonumber \\
``\Delta A" &=& ``\pd{y}^{2} A + \pd{z}^{2}A" \nonumber
\Eeq
where $\sigma$ is the problem \texttt{eigenvalue}, and $k_{y},\,k_{z}$ are problem \texttt{parameters}, along with $f,\,S,\,B_{0},\, \nu,\, \eta$.

The system of \texttt{equations} is:
\Beq\label{u-eq}
\pd{x} v_{x} + \pd{y} v_{y} + \pd{z} v_{z}  &=& 0\\
D_{t} v_{x}  - f v_{y} + \pd{x} p - B_{0} \pd{z} b_{x}  + \nu ( \pd{y} \omega_{z} - \pd{z} \omega_{y}) &=& 0\\
D_{t} v_{y}  + (f+S) v_{x} + \pd{y} p - B_{0} \pd{z} b_{y} + \nu ( \pd{z} \omega_{x} - \pd{x} \omega_{z})&=& 0\label{v-eq} \\
D_{t} v_{z} + \pd{z} p - B_{0} \pd{z} b_{z} + \nu ( \pd{x} \omega_{y} - \pd{y} \omega_{x})&=& 0 \\
\omega_{y} - \pd{z} v_{x} + \pd{x} v_{z} &=& 0 \\
\omega_{z} - \pd{x} v_{y} + \pd{y} v_{x} &=& 0 \\
\pd{x} b_{x} + \pd{y} b_{y} + \pd{z} b_{z}   &=& 0\\
D_{t} b_{x} - B_{0} \pd{z} v_{x} + \eta ( \pd{y} j_{z} - \pd{z} j_{y})&=& 0\\
D_{t} j_{x}  - \eta( \pd{x}j_{xx} +  \Delta j_{x} )  - B_{0} \pd{z} \omega_{x} + S \pd{z} b_{x} &=& 0\\
j_{xx} - \pd{x} j_{x} &=& 0
\Eeq
We can see there are the same number of equations as variables, each equations has only 1st-order derivatives in $x$, and every variable has at least one $\pd{x}$. 
It appears (for example) that there are no $\pd{x}$ terms in the $b_{x}$ equation, but this due to the substitutions, 
\Beq
\pd{y} j_{z} - \pd{z} j_{y} =  \pd{y}(\pd{x} b_{y} - \pd{y} b_{x}) - \pd{z}(\pd{z} b_{x} - \pd{x} b_{z})  =  \pd{y}\pd{x}b_{y} + \pd{z}\pd{x}b_{z} - (\pd{y}^{2} + \pd{z}^{2})b_{x}.
\Eeq
Also,
\Beq
\pd{x}j_{x} \ = \ \pd{y} \pd{x} b_{z} - \pd{z} \pd{x}b_{y}.
\Eeq
The entire system is formulated and solved as a generalized eigenvalue problem with non-constant coefficients 
\Beq
\sigma \, M \cdot X + L_{0}(x) \cdot X + L_{1}\cdot \pd{x} X \ = \ 0.
\Eeq
The problem is ``generalized" because $M$ is not full rank. The system is 10th order because $L_{1}$ is full rank,
\Beq
\text{det}(L_{1}) \ = \ -\eta ^2 \nu ^2 \left(k_y^2+k_z^2\right).
\Eeq
We therefore impose the 10 \texttt{boundary conditions} 
\Beq
v_{x} \ = \ \omega_{y} \ = \ \omega_{z} \ = \ b_{x} \ = \ j_{xx} \ = \ 0 \quad \text{at} \quad x \ = \ \pm d/2
\Eeq
We could choose any other set of boundary conditions constructed out of a linearly independent combinations of the problem variables. 
We choose impenetrable stress-free and perfectly conducting because it conforms with the purely impenetrable conditions for an ideal calculation, thus minimizing the arbitrary influence of boundary layers.

Respectively $M,\,L_{0}(x),\,L_{1}$:
\Beq
&
\tiny \nonumber
\left[
\begin{array}{cccccccccc}
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & -i k_z & i k_y & 0 \\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
\end{array}
\right]\\
&
\tiny \nonumber
\left[
\begin{array}{cccccccccc}
 0 & 0 & i k_y & i k_z & 0 & 0 & 0 & 0 & 0 & 0 \\
 0 & i S x k_y & -f & 0 & -i \nu  k_z & i \nu  k_y
   & -i B_{0} k_z & 0 & 0 & 0 \\
 i k_y & f+S & \nu  k_z^2+i S x k_y & -\nu  k_y
   k_z & 0 & 0 & 0 & -i B_{0} k_z & 0 & 0 \\
 i k_z & 0 & -\nu  k_y k_z & k_y \left(\nu  k_y+i
   S x\right) & 0 & 0 & 0 & 0 & -i B_{0} k_z & 0 \\
 0 & -i k_z & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
 0 & i k_y & 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 \\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & i k_y & i k_z & 0 \\
 0 & -i B_{0} k_z & 0 & 0 & 0 & 0 & \eta  |k|^2+i S x
   k_y & 0 & 0 & 0 \\
 0 & 0 & -B_{0} k_z^2 & B_{0} k_y k_z & 0 & 0 & i S k_z &
   k_z \left(S x k_y-i \eta  |k|^2\right) & k_y
   \left(-S x k_y+i \eta  |k|^2\right) & 0 \\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 \\
\end{array}
\right]\\
&
\tiny \nonumber
\left[
\begin{array}{cccccccccc}
 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
 0 & 0 & 0 & 0 & 0 & -\nu  & 0 & 0 & 0 & 0 \\
 0 & 0 & 0 & 0 & \nu  & 0 & 0 & 0 & 0 & 0 \\
 0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
 0 & 0 & -1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
 0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & i \eta  k_y & i \eta 
   k_z & 0 \\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & -\eta  \\
 0 & 0 & 0 & 0 & 0 & 0 & 0 & i k_z & -i k_y & 0 \\
\end{array}
\right]
\Eeq

\section{Reynolds Number and Departure from Ideal MHD}
\label{sec:reyn}

In order to ensure that our calculations are probing the ideal MHD regime we are interested in, we have performed as series of runs (24-27) with the $\eta = \nu$ ranging from $10^{-6}$ to $10^{-2}$.
In our units, this is equivalent to varying the Reynolds and Magnetic Reynolds numbers $\Reyn = \Reym = 10^2 - 10^6$.
Figure~\pageref{fig:reynolds} shows the growth rates at $\SSC=1.05$ for five values of $\Reyn = \Reym$.
For $\Reyn \gtrsim 10^4$, the maximum growth rate is located at the same position and the same growth rate; the unstable region expands, as expected.
This suggests that our results at $\Reyn = 10^5$ are reasonably close to ideal MHD for the purposes of understanding the MRI's preference for 3D modes near shear onset.
\begin{figure}[h!]
  \centering
  \includegraphics[width=\textwidth]{re_plots.pdf}
  \caption{Growth rates at $\SSC = 1.02$ for varying diffusivities. Note that the top three panels are essentially identical, suggesting that above $\Reyn \gtrsim 10^4$ we are reasonably close to ideal conditions.}
  \label{fig:reynolds}
\end{figure}

\section{Asymptotic Calculation}
\label{sec:asymp}
Here, we outline the asymptotic calculation for the leading order correction to the 2D growth rates in when accounting for three dimensional effects via finite $k_y$.
Define the domain as $ - d/2 < x < d/2$. Being in the rotating frame means the shear has no net
\Beq
\int_{-d/2}^{d/2} V_{0,y}(x) \text{d} x  \ = \ 0 \quad \implies  \quad V_{0,y}(x) \ = \ S\, x
\Eeq
The full linear ideal equations are:
\Beq\label{u-eq}
\pd{t} v_{x} + S x \pd{y} v_{x} - f v_{y} + \pd{x} p - B_{0} \pd{z} b_{x} &=& 0\\
\pd{t} v_{y} + S x \pd{y} v_{y} + (f+S) v_{x} + \pd{y} p - B_{0} \pd{z} b_{y} &=& 0\label{v-eq} \\
\pd{t} v_{z} + S x \pd{y} v_{z} + \pd{z} p - B_{0} \pd{z} b_{z} &=& 0 \\
\pd{x} v_{x} + \pd{y} v_{y} + \pd{z} v_{z}  &=& 0\\
\pd{t} b_{x} + S x \pd{y} b_{x} - B_{0} \pd{z} v_{x} &=& 0\\
\pd{t} b_{y} + S x \pd{y} b_{y} - B_{0} \pd{z} v_{y} - S b_{x}  &=& 0\\
\pd{t} b_{z} + S x \pd{y} b_{z} - B_{0} \pd{z} v_{z} &=& 0 \label{bz-eq}
\Eeq
The divergence-free condition for the velocity implies the same for the magnetic field.
The system is 2nd order because $\pd{x}$ only appears in two places.
All variables take the form
\Beq
p = \hat{P}(x) e^{ i ( \omega t + k z + \ell y ) } 
\Eeq
We solve the ideal eigenvalue problem in \textit{Dedalus} in the above formulation. To make analytical progress, define the frequency ``parameters'' 
\Beq
\omega_{S}(x) \ \equiv \ \omega + S x \ell, \quad \omega_{A}  \ \equiv \ B_{0} k.
\Eeq
We can use equations (\ref{v-eq}--\ref{bz-eq}) to find all amplitudes in terms of $\hat{V}_{x}(x)$ and $\hat{V}_{x}'(x)$. 
\Beq
\hat{V}_{y}(x) &=& \frac{i \left(\ell 
   \hat{V}_{x}'(x) + k^{2} \frac{ \left((f+S) \omega
   _S^2-S \omega _A^2\right)}{\omega _S
   \left(\omega _S^2-\omega _A^2\right)} \hat{V}_{x}(x)\right)}{k^2+\ell ^2} \\
\hat{V}_{z}(x) &=& \frac{i k \left(\hat{V}_{x}'(x)- \frac{ \left((f+S) \omega
   _S^2-S \omega _A^2\right)}{\omega _S
   \left(\omega _S^2-\omega _A^2\right)}  \ell 
   \hat{V}_{x}(x)\right)}{k^2+\ell ^2} \\
   \hat{P}(x) &=& \frac{i \left(\omega _A^2-\omega _S^2\right)
   \left(\hat{V}_{x}'(x) -  \ell \frac{ \left((f+S) \omega
   _S^2-S \omega _A^2\right)}{\omega _S
   \left(\omega _S^2-\omega _A^2\right)}
   \hat{V}_{x}(x)\right)}{\left(k^2+\ell ^2\right)
   \omega _S} \\
   \hat{B}_{x}(x) &=& \frac{\omega _A \hat{V}_{x}(x)}{\omega _S} \\
   \hat{B}_{y}(x) &=&\frac{\omega _A \hat{V}_{y}(x)}{\omega _S} -\frac{i S \omega _A \hat{V}_{x}(x)}{\omega _S^2} \\
   \hat{B}_{z}(x) &=& \frac{\omega _A \hat{V}_{z}(x)}{\omega _S}
   \Eeq
Substitute everything into equation (\ref{u-eq}) to get a second-order equation for $\hat{V}_{x}(x)$ of the general form 
\Beq
\alpha(x) \hat{V}_{x}''(x) + \beta(x) \hat{V}_{x}'(x) + \gamma(x) \hat{V}_{x}(x) \ = \ 0
\Eeq
where
\Beq
\alpha(x) &=& \omega _S^2-\omega _A^2 \\ 
   \beta(x) &=& \frac{2 S \ell  \omega _A^2}{\omega _S} \\
   \gamma(x) &=& \omega _A^2 \left(\frac{f^2 k^2}{\omega
   _S^2-\omega _A^2}-\frac{2 S^2 \ell
   ^2}{\omega _S^2}+k^2+\ell ^2\right)+f k^2
   (f+S)-\left(k^2+\ell ^2\right) \omega _S^2
\Eeq

We can eliminate the first-order term via
\Beq
\hat{V}_{x}(x) \ = \ \chi(x) \psi(x) \quad \text{where} \quad \chi'(x) \  = \ -\frac{\beta (x) \chi (x)}{2 \alpha (x)}
\Eeq
\Beq
-\psi''(x) + \Phi(x) \psi(x) \ = \ 0 \label{Schr-eq}
\Eeq
\Beq
\Phi(x)  \ = \ \frac{-2 \beta (x) \alpha '(x)+2 \alpha (x)
   \left(\beta '(x)-2 \gamma (x)\right)+\beta
   (x)^2}{4 \alpha (x)^2}
\Eeq
Putting everything together 
\Beq
\Phi(x) \ = \ \frac{S  \left(f k^2-S \ell
   ^2\right) \omega _A^2 -f k^2 (f+S) \omega
   _S^2}{\left(\omega _A^2-\omega
   _S^2\right){}^2} + k^{2} + \ell^{2}
\Eeq
If $\omega_{S}=-i \gamma+Sx\ell$, with $Re(\gamma)\neq0$ then the denominator of $\Phi(x)$ cannot vanish. 
We want to solve the Schr\"{o}dinger-type equation~(\ref{Schr-eq}) with the boundary conditions 
\Beq
\psi(x=\pm d/2) = 0.
\Eeq
We solve perturbatively near the critical values for the 2D  instability. We rescale each parameters in terms of a bookkeeping parameter, $\eps$.
\Beq
S  \ \to \  - \frac{\pi ^2 B_{0}^2}{d^2 f} ( 1 + \eps^{2} R), \quad \omega \ \to \ \eps^{2} \omega , \quad k \ \to \ \eps k, \quad \ell \ \to \ \eps^{2} \ell
\Eeq 
Also 
\Beq
q  \ \equiv  \ \frac{\pi ^2 B_{0}^2}{d^2 f^{2}}
\Eeq
Expanding to $\mathcal{O}(\eps^{2})$, 
\Beq
\Phi(x) \ = \ \frac{\pi^{2}}{d^{2}} + \eps^{2} \frac{B_{0}^2 f^2 q \left(k^2 R+q
   \ell ^2\right) - B_{0}^4 k^4 +f^2 (q+1) (\omega -f q x \ell
   )^2  }{B_{0}^4 k^2}
\Eeq
For leading order 
\Beq
\psi_{0}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{0}(x) \ = \ 0  \quad \implies \quad \psi_{0}(x) \ = \ \cos(\pi x / d).
\Eeq
The next order is
\Beq
\psi_{2}''(x) + \frac{\pi^{2}}{d^{2}}  \psi_{2}(x) + \Phi_{2}(x)\psi_{0}(x)\ = \ 0.
\Eeq
The solvability condition for $\psi_{2}(x)$ is 
\Beq
\int_{-d/2}^{d/2} \Phi_{2}(x)\psi_{0}(x)^{2} \text{d} x \ = \ 0.
\Eeq
This implies 
\Beq
-\omega^{2} \ = \ \frac{q\,B_{0}^{2}}{(q+1)} \left[ k_z^2 \left(R-\frac{d^2}{\pi^{2}} k_z^2\right) \ + \ \frac{\left(6+\pi ^2\right) q+\pi
   ^2-6}{12 } \, k_{y}^{2}  \right] \ + \ ...
\Eeq

\end{document}

%  LocalWords:  Magnetorotational MHD EigenValueProblem Dedalus th nd
%  LocalWords:  parallelized Chebyshev cccccccccccccc Rossby det Sx
%  LocalWords:  cccccccccc diffusivities analytical Schr dinger
%  LocalWords:  perturbatively rescale solvability

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2021-07-14 python/calc_asymptotic_approx.py

@@ -1,0 +1,12 @@
+"""Use leading order approximation to compute growth rate
+
+"""
+
+import numpy as np
+
+def calc_asymptotic_growth(R, q, B, d, ky, kz):
+    Rprime = R**2 - 1
+    omega2 = q*B**2/(q+1)*(kz**2*(d**2/np.pi**2 * kz**2 - Rprime) - ((6 + np.pi**2)*q + np.pi**2 - 6)*ky**2/12.)
+    omega = np.sqrt(-omega2)
+
+    return omega, omega2

### 2021-07-14 python/make_eigenvalue_plot.py

@@ -1,0 +1,72 @@
+"""
+This script plots both the largest growth rates of the MRI data
+passed to the script. 
+
+Usage:
+    make_eigenvector_plots.py [--vmin=<vmin> --vmax=<vmax> --q=<q> --no-contours --label=<label>] <filename>
+
+Options:
+    --vmin=<vmin>    min for colorbar [default: 0]
+    --vmax=<vmax>    max for colorbar [default: 0.701]
+    --q=<q>          q = S/f [default: 0.75]
+    --no-contours    turn off contours
+    --label=<label>  run label
+"""
+
+import h5py
+from docopt import docopt
+from pathlib import Path
+import numpy as np
+import matplotlib.pyplot as plt
+
+plt.style.use('prl')
+
+# Parses filename passed to script
+args = docopt(__doc__)
+
+filename = Path(args['<filename>'])
+vmin = float(args['--vmin'])
+vmax = float(args['--vmax'])
+q = float(args['--q'])
+contours = not args['--no-contours']
+outbase = Path("plots")
+label = args['--label']
+
+# Plot growth rates
+datafile = h5py.File(filename,'r')
+gamma_global = datafile['gamma'][:]
+gamma_r = gamma_global.real/q # scale by S...gamma is already multiplied by f in script
+
+#gamma_r[np.where(gamma_r<0)] = 0.0
+#max_gamma = datafile['gamma'].attrs['max growth rate']
+ky_global    = datafile['ky']
+kz_global    = datafile['kz']
+print(ky_global.shape)
+print(gamma_r.shape)
+try:
+    max_ky = datafile['gamma'].attrs['max ky']
+    max_kz = datafile['gamma'].attrs['max kz']
+except KeyError:
+    ky_index, kz_index = np.where(gamma_r == gamma_r.max())
+    max_ky = ky_global[ky_index[0]]
+    max_kz = kz_global[kz_index[0]]
+
+print("max growth rate on grid = {}".format(gamma_r.max()))
+
+contour_levels = np.linspace(0,gamma_r.max(),10)
+PCM = plt.imshow(gamma_r, extent=[kz_global[:].min(),kz_global[:].max(),ky_global[:].min(),ky_global[:].max()], vmin=vmin,vmax=vmax, origin='lower')
+if contours:
[…]

### 2021-07-14 python/mri.py

@@ -1,0 +1,306 @@
+"""
+The magnetorotational instability prefers three dimensions.
+    
+Dedalus script for calculating the maximum growth rates for the
+3D MRI over a range of wavenumbers.
+
+This script can be ran serially or in parallel, and produces an h5py file
+that contains maximum eigenvalues and the corresponding eigenvectors.
+
+Pass it a .cfg file to specify parameters for the particular solve.
+
+To run using 4 processes, you would use:
+    $ mpiexec -n 4 python3 mri.py mri_params.cfg
+
+"""
+
+import time
+import configparser
+from configparser import ConfigParser
+import argparse
+from pathlib import Path
+import numpy as np
+import scipy.optimize as opt
+import os
+import h5py
+import matplotlib.pyplot as plt
+import dedalus.public as de
+from mpi4py import MPI
+CW = MPI.COMM_WORLD
+import logging
+logger = logging.getLogger(__name__)
+
+# Parses .cfg filename passed to script
+parser = argparse.ArgumentParser(description='Passes filename')
+parser.add_argument('filename', metavar='Rc', type=str, help='.h5 file to plot eigenvectors for maximum eigenvalue')
+args = parser.parse_args()
+filename = Path(vars(args)['filename'])
+outbase = Path("data")
+
+# Parse .cfg file to set global parameters for script
+config = ConfigParser()
+config.read(str(filename))
+
+logger.info('Running mri.py with the following parameters:')
+logger.info(config.items('parameters'))
+
+try:
+    dense = config.getboolean('solver','dense')
+except:
+    dense = False
+if dense:
+    logger.info("Using dense solver.")
+    dense_threshold = config.getfloat('solver','dense_threshold')
+else:
+    logger.info("Using sparse solver.")
+
+Nx = config.getint('parameters','Nx')
+Lx = eval(config.get('parameters','Lx'))
+B = config.getfloat('parameters','B')
+
+Nmodes = config.getint('parameters','Nmodes')
+
+R      =  config.getfloat('parameters','R')
+q      =  config.getfloat('parameters','q')
+
[…]

### 2021-07-14 python/mri_single_yz_mode.py

@@ -1,0 +1,238 @@
+"""
+The magnetorotational instability prefers three dimensions.
+
+Single mode calculation. Returns spectrum.
+
+Usage:
+    mri_single_yz_mode.py [--ideal] <config_file> <ky> <kz>
+
+Options:
+    --ideal    Use Ideal MHD
+    --hardwall Use experimental boundary conditions
+"""
+
+from docopt import docopt
+import time
+from configparser import ConfigParser
+from pathlib import Path
+import numpy as np
+import os
+import h5py
+from eigentools import Eigenproblem
+import dedalus.public as de
+from mpi4py import MPI
+CW = MPI.COMM_WORLD
+import logging
+logger = logging.getLogger(__name__)
+
+args = docopt(__doc__)
+ky = float(args['<ky>'])
+kz = float(args['<kz>'])
+ideal = args['--ideal']
+hardwall = args['--hardwall']
+filename = Path(args['<config_file>'])
+outbase = Path("data")
+
+logger.info("Solving for mode ky = {}, kz = {}".format(ky, kz))
+# Parse .cfg file to set global parameters for script
+config = ConfigParser()
+config.read(str(filename))
+
+logger.info('Running mri.py with the following parameters:')
+logger.info(config.items('parameters'))
+try:
+    dense = config.getboolean('solver','dense')
+except:
+    dense = False
+if dense:
+    logger.info("Using dense solver.")
+    dense_threshold = config.getfloat('solver','dense_threshold')
+else:
+    logger.info("Using sparse solver.")
+
+Nx = config.getint('parameters','Nx')
+Lx = eval(config.get('parameters','Lx'))
+B = config.getfloat('parameters','B')
+
+Nmodes = config.getint('parameters','Nmodes')
+
+R      =  config.getfloat('parameters','R')
+q      =  config.getfloat('parameters','q')
+
+kymin = config.getfloat('parameters','kymin')
+kymax = config.getfloat('parameters','kymax')
+Nky = config.getint('parameters','Nky')
+
+kzmin = config.getfloat('parameters','kzmin')
[…]

### 2021-07-14 python/plot_3D_eigenvalues.py

@@ -1,0 +1,125 @@
+"""
+constructs 3D eigenvector structures; plots them in z = 0 plane.
+
+python3 path/to/datafile.h5
+"""
+import h5py
+import argparse
+from pathlib import Path
+import numpy as np
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1 import ImageGrid
+
+# Parses filename passed to script
+parser = argparse.ArgumentParser(description='Passes filename')
+parser.add_argument('filename', metavar='Rc', type=str, help='.h5 file to plot eigenvectors for maximum eigenvalue')
+args = parser.parse_args()
+filename = Path(vars(args)['filename'])
+outbase = Path("plots")
+
+eigenvector_vars = ['p','vx','vy','vz','ωy','ωz','bx','by','bz','jxx']
+
+# Import .h5 file passed to script 
+datafile = h5py.File(filename,'r')
+eigvec = datafile['eigvec'][:,:,:,:]
+x = datafile['x'][:]
+gamma_global = datafile['gamma'][:,:]
+
+# Find grid point with largest growth rate to plot corresponding eigenvectors
+gamma_r = gamma_global.real
+gamma_r[np.where(gamma_r<0)] = 0.0
+max_point = np.amax(gamma_r)
+z_max = np.where(gamma_r == max_point)[1][0]
+y_max = np.where(gamma_r == max_point)[0][0]
+# zv = kz_global[z_max] #max x value
+# yv = ky_global[y_max] #max y value
+# maxvalues[0,i] = zv
+# maxvalues[1,i] = yv
+# print(z_max)
+# print(y_max)
+
+ky = datafile['gamma'].attrs['max ky']
+kz = datafile['gamma'].attrs['max kz']
+
+Ly = 2*np.pi/ky
+Lz = 2*np.pi/kz
+
+n_yz = 128
+y = np.linspace(0, Ly, n_yz, endpoint=False)
+z = np.linspace(0, Lz, n_yz, endpoint=False)
+
+zz,yy,xx = np.meshgrid(z,y,x,indexing='ij')
+fig = plt.figure(figsize=(6,6))
+
+yz_dep = np.exp(1j*(ky*yy +kz*zz))
+
+p = (eigvec[y_max, z_max,0,:]*yz_dep).real
[…]

### 2021-07-14 python/plot_fig1.py

@@ -1,0 +1,186 @@
+import h5py
+from pathlib import Path
+import numpy as np
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1 import AxesGrid
+from calc_asymptotic_approx import calc_asymptotic_growth
+from basic_units import radians
+plt.style.use('prl')
+
+# params
+filebase = 'data/run_{:d}_output.h5'
+runs = [15, 39, 12, 13]
+SSC = [0.64, 1.002001, 1.2, 2.]
+q = 0.75
+index = 0
+
+fig = plt.figure(figsize=(16,8))
+# grid = AxesGrid(fig, (0.125, 0.5,0.8,0.35),
+#                 nrows_ncols=(1, 4),
+#                 axes_pad=0.1,
+#                 cbar_mode='single',
+#                 cbar_location='right',
+#                 cbar_pad=0.1
+#                 )
+xgutter = 0.1
+pad = 0.01
+w = (1. - xgutter*2 - pad*3)/4
+print(w)
+h = 0.4
+grid = [fig.add_axes([xgutter+i*(w+pad),0.55,w,h]) for i in range(4)]
+for i,r in enumerate(runs):
+    ax = grid[i]
+    print("loading data from {}".format(filebase.format(r)))
+    datafile = h5py.File(filebase.format(r),'r')
+    ky_global    = datafile['ky']
+    kz_global    = datafile['kz']
+    gamma_global = datafile['gamma'][:]
+    gamma_r = gamma_global.real/q # scale by S
+    #gamma_r[np.where(gamma_r<0)] = 0.0
+    max_gamma = datafile['gamma'].attrs['max growth rate']
+    max_ky = datafile['gamma'].attrs['max ky']
+    max_kz = datafile['gamma'].attrs['max kz']
+    contour_levels = np.linspace(0,gamma_r.max(),5)
+    vmin = 0.
+    vmax = 0.1
+    print("max gamma_r/S = {}".format(gamma_r.max()))
[…]

### 2021-07-14 python/plot_mean_field_alpha.py

@@ -1,0 +1,34 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import h5py
+
+plt.style.use('prl')
+
+dfname = "run_11_output.h5"
+df = h5py.File(dfname,"r")
+x = df['x_grid'][:,:,:]
+u = df['u'][:,:,:]
+v = df['v'][:,:,:]
+w = df['w'][:,:,:]
+Bx = df['Bx'][:,:,:]
+By = df['By'][:,:,:]
+Bz = df['Bz'][:,:,:]
+cheby_factor = np.sqrt(np.pi**2 - x**2)
+
+u_mean = np.sqrt(np.mean((u**2 + v**2 + w**2)*cheby_factor)/np.mean(cheby_factor))
+B_mean = np.sqrt(np.mean((Bx**2 + By**2 + Bz**2)*cheby_factor)/np.mean(cheby_factor))
+norm = u_mean*B_mean
+alpha = (u*By - v*Bx)
+alpha_yz = alpha.mean(axis=0).mean(axis=0)/norm
+
+plt.plot(x[0,0,:], alpha_yz)
+plt.xlabel(r"$x$")
+plt.ylabel(r"$\langle v_x b_y - v_y b_x \rangle/|v||b|$")
+
+ax =plt.gca()
+ax.set_xlim(-np.pi/2,np.pi/2)
+ax.set_xticks([-np.pi/2, 0, np.pi/2])
+ax.set_xticklabels([r'$-\pi/2$', 0, r'$\pi/2$'])
+plt.tight_layout()
+plt.savefig('plots/mean_field_alpha_{}.pdf'.format(dfname))
+

### 2021-07-14 python/plot_phi.py

@@ -1,0 +1,136 @@
+"""
+phi = arctan(ky/kz)
+
+"""
+
+import numpy as np
+import numpy.polynomial.polynomial as poly
+import h5py
+import matplotlib.pyplot as plt
+from pathlib import Path
+from basic_units import radians
+
+plt.style.use('prl')
+
+use_opt = False
+
+def calc_phi():
+    basedir = Path("data/")
+    runs = [34, 33, 31, 21, 15, 11, 12, 16, 30, 13, 27, 28, 29, 26, 24, 17, 14, 18]
+    R = [np.sqrt(0.2), np.sqrt(0.3), np.sqrt(0.4), np.sqrt(0.5), 0.8, 1.01, 1.2, np.sqrt(1.75), 1.375, np.sqrt(2.0), np.sqrt(2.01), np.sqrt(2.015), 1.425, np.sqrt(2.05), np.sqrt(2.1), 1.5, np.sqrt(2.5), 2.0]
+    ssc = np.array(R)**2
+    phi = []
+
+    growth = []
+    dk = []
+    for run in runs:
+        dfname = Path("run_{:d}_output.h5".format(run))
+        df = h5py.File(basedir/dfname, "r")
+        if use_opt:
+            #phi.append(np.arctan2(df['gamma'].attrs['max ky'], df['gamma'].attrs['max kz']))
+            ky = df['gamma'].attrs['max ky']
+            kx = df['gamma'].attrs['max kz']
+            gamma_max = df['gamma'].attrs['max growth rate']
+        else:
+            gamma_global = df['gamma'][:,:]
+            max_where = np.unravel_index(gamma_global.argmax(), gamma_global.shape)
+            ky = df['ky'][max_where[0]] 
+            kz = df['kz'][max_where[1]]
+            gamma_max = gamma_global.real.max()
+        dk.append(df['ky'][1]-df['ky'][0])
+        kk = np.sqrt(ky**2 + kz**2)
+        phi.append(np.arcsin(ky/kk))
+        growth.append(gamma_max)
+        df.close()
+    phi = np.array(phi)
+    ssc = np.array(ssc)
+    dk = np.array(dk)
+
+    return phi, ssc, dk, growth
+
+if __name__ == "__main__":
[…]

### 2021-07-14 python/plot_single_mode_evecs.py

@@ -1,0 +1,57 @@
+import matplotlib.pyplot as plt
+import h5py
+import sys
+import numpy as np
+from pathlib import Path
+
+plt.style.use('prl')
+filename = Path(sys.argv[-1])
+
+rows = 2
+cols = 3
+names = [r'$v_x$',r'$v_y$',r'$v_z$',r'$b_x$',r'$b_y$',r'$b_z$']
+cmaps = ['RdBu','RdBu','RdBu','PuOr','PuOr','PuOr']
+plt.figure(figsize=(12,6))
+with h5py.File(filename, 'r') as df:
+    x = df['eigvecs'].attrs['x']
+    for i in range(1,7):
+        plt.subplot(rows,cols,i)
+        plt.plot(x, df['eigvecs'][i,:].real, label='real')
+        plt.plot(x, df['eigvecs'][i,:].imag, label='imag')
+        if i == 1:
+            plt.legend()
+        plt.xlabel(r'$x$')
+        plt.ylabel(names[i-1])
+
+    plt.tight_layout()
+    plt.savefig('plots/' + filename.stem + 'eigvecs')
+
+    # 2D plots
+    n_yz = 128
+    ky = df['eigvals'].attrs['ky']
+    kz = df['eigvals'].attrs['kz']
+    Ly = 2*np.pi/ky
+    Lz = 2*np.pi/kz
+
+    y = np.linspace(0, Ly, n_yz, endpoint=False)
+
+    yyy, xxx = np.meshgrid(y,x,indexing='ij')
+    y_dep = np.exp(1j*ky*yyy)
+    print(ky)
+    plt.figure(figsize=(10,10))
+    eigvec = df['eigvecs']
+    for i in range(1,7):
+        ax = plt.subplot(rows,cols,i)
+        pcol = plt.pcolormesh(xxx,yyy,(eigvec[i,:]*y_dep).real,linewidth=0,rasterized=True, cmap=cmaps[i-1])
+        #pcol.set_edgecolor('face')
+        plt.title(names[i-1])
+        if i !=4:
+            ax.axes.get_xaxis().set_visible(False)
+            ax.axes.get_yaxis().set_visible(False)
+        else:
+            plt.xlabel(r'$x$')
+            plt.ylabel(r'$y$')
+            ax.set_xticklabels(["$-\pi/2$", 0, "$\pi/2$"])
+            ax.set_xticks([-np.pi/2, 0, np.pi/2])
+
+    plt.savefig('plots/eigvecs_xy_' + filename.stem)

### 2021-07-14 python/plot_spectrum.py

@@ -1,0 +1,56 @@
+"""plot_spectrum.py [--asymptotic] <filename>
+
+Usage:
+    plot_spectrum.py [--asymptotic] <filename>
+
+"""
+
+import numpy as np
+import sys
+from pathlib import Path
+import h5py
+import matplotlib.pyplot as plt
+from calc_asymptotic_approx import calc_asymptotic_growth
+
+from docopt import docopt
+
+plt.style.use('prl')
+
+# parse arguments
+args = docopt(__doc__)
+filename = args['<filename>']
+asymptotic = args['--asymptotic']
+
+outbase = Path("plots")
+filename = Path(filename)
+data = h5py.File(filename, 'r')
+spectrum = data['eigvals'][:]
+R = data['eigvals'].attrs['R']
+B = data['eigvals'].attrs['B']
+q = data['eigvals'].attrs['q']
+d = data['eigvals'].attrs['d']
+ky = data['eigvals'].attrs['ky']
+kz = data['eigvals'].attrs['kz']
+# check asymptotic calculation
+if asymptotic:
+    omega = calc_asymptotic_growth(R, q, B, d, ky, kz)
+    print("asymptotic omega = {}".format(omega))
+
+thresh = 2
+
+
+spec_thresh = spectrum[np.abs(spectrum) < thresh]
+plt.scatter(spec_thresh[spec_thresh.real <= 0].real, spec_thresh[spec_thresh.real <= 0].imag)
+plt.scatter(spec_thresh[spec_thresh.real > 0].real, spec_thresh[spec_thresh.real > 0].imag)
+plt.gca().set_prop_cycle(None)
+if asymptotic:
+    plt.plot(-omega, 0., marker='^')
+    plt.plot(omega, 0., marker='^')
+plt.xlim(-0.25,0.25)
+plt.ylim(-0.8,0.8)
+plt.xlabel("real")
+plt.ylabel("imag")
+
+plot_file_name = Path(filename.stem + '_spectrum.png')
+plt.tight_layout()
+plt.savefig(outbase/plot_file_name, dpi=300)

### 2021-07-14 python/run_mri.sh

@@ -1,0 +1,9 @@
+#!/usr/bin/bash
+#SBATCH --nodes=4
+#SBATCH --partition=faculty
+
+#conda activate dedalus
+
+date
+mpiexec -np 28 python3 mri.py runs/run_1.cfg
+date

### 2021-07-14 python/single_mode.py

@@ -1,0 +1,244 @@
+"""
+The magnetorotational instability prefers three dimensions.
+
+Single mode calculation. Returns spectrum.
+
+Usage:
+    mri_single_yz_mode.py [--ideal] <config_file> <ky> <kz>
+
+Options:
+    --ideal    Use Ideal MHD
+    --hardwall Use experimental boundary conditions
+"""
+
+from docopt import docopt
+import time
+from configparser import ConfigParser
+from pathlib import Path
+import numpy as np
+import os
+import h5py
+from eigentools import Eigenproblem
+import dedalus.public as de
+from mpi4py import MPI
+CW = MPI.COMM_WORLD
+import logging
+logger = logging.getLogger(__name__)
+
+# args = docopt(__doc__)
+# ky = float(args['<ky>'])
+# kz = float(args['<kz>'])
+# ideal = args['--ideal']
+# hardwall = args['--hardwall']
+# filename = Path(args['<config_file>'])
+# args = docopt(__doc__)
+ky = 0.1
+kz = 0.5
+ideal = True
+hardwall = False
+filename = Path('mri_options.cfg')
+outbase = Path("data")
+
+logger.info("Solving for mode ky = {}, kz = {}".format(ky, kz))
+# Parse .cfg file to set global parameters for script
+config = ConfigParser()
+config.read(str(filename))
+
+logger.info('Running mri.py with the following parameters:')
+logger.info(config.items('parameters'))
+try:
+    dense = config.getboolean('solver','dense')
+except:
+    dense = False
+if dense:
+    logger.info("Using dense solver.")
+    dense_threshold = config.getfloat('solver','dense_threshold')
+else:
+    logger.info("Using sparse solver.")
+
+Nx = config.getint('parameters','Nx')
+Lx = eval(config.get('parameters','Lx'))
+B = config.getfloat('parameters','B')
+
+Nmodes = config.getint('parameters','Nmodes')
+
+R      =  config.getfloat('parameters','R')
+q      =  config.getfloat('parameters','q')
+
+kymin = config.getfloat('parameters','kymin')
[…]

### 2021-08-12 evan_ic_stuff.py
migrate local

@@ -1,0 +1,35 @@
+def filter_field(field, frac=0.25):
+    """
+    Filter a field in coefficient space by cutting off all coefficient above
+    a given threshold.  This is accomplished by changing the scale of a field,
+    forcing it into coefficient space at that small scale, then coming back to
+    the original scale.
+    Inputs:
+        field   - The dedalus field to filter
+        frac    - The fraction of coefficients to KEEP POWER IN.  If frac=0.25,
+                    The upper 75% of coefficients are set to 0.
+    """
+    dom = field.domain
+    logger.info("filtering field {} with frac={} using a set-scales approach".format(field.name,frac))
+    orig_scale = field.scales
+    field.set_scales(frac, keep_data=True)
+    field['c']
+    field['g']
+    field.set_scales(orig_scale, keep_data=True)
+    
+def global_noise(domain, seed=42, **kwargs):
+    """
+    Create a field fielled with random noise of order 1.  Modify seed to
+    get varying noise, keep seed the same to directly compare runs.
+    """
+    # Random perturbations, initialized globally for same results in parallel
+    gshape = domain.dist.grid_layout.global_shape(scales=domain.dealias)
+    slices = domain.dist.grid_layout.slices(scales=domain.dealias)
+    rand = np.random.RandomState(seed=seed)
+    noise = rand.standard_normal(gshape)[slices]
+    # filter in k-space
+    noise_field = domain.new_field()
+    noise_field.set_scales(domain.dealias, keep_data=False)
+    noise_field['g'] = noise
+    filter_field(noise_field, **kwargs)
+    return noise_field

### 2021-08-12 make_video.py

@@ -8,11 +8,11 @@ img_array = []
 test = os.getcwd()
 
 # directory = "/nobackup22/loconno2/mri_simulations/frames_wedtwo/"
-directory = path = os.path.dirname(os.path.abspath(__file__)) + "/frames_friz/"
+directory = path = os.path.dirname(os.path.abspath(__file__)) + "/frames_sliced/"
 fileType = "*.png"
-videoName = 'frames_friz.avi'
+videoName = 'frames_sliced_midz.avi'
 files = natsorted(glob.glob(directory + fileType))
-files = files[::]
+files = files[:150:]
 nFrames = len(files)
 
 index = 0

### 2021-08-12 mri_nonideal.sh

@@ -1,7 +1,7 @@
 #PBS -S /bin/bash
-#PBS -N mri_nonideal
-#PBS -l select=19:ncpus=28:mpiprocs=512:model=bro
-#PBS -l walltime=16:00:00
+#PBS -N mri_nonideal_diff2en4
+#PBS -l select=10:ncpus=28:mpiprocs=256:model=bro
+#PBS -l walltime=24:00:00
 #PBS -j oe
 module load mpi-sgi/mpt
 module load comp-intel
@@ -14,8 +14,12 @@ conda activate dedalus
 # support lots of text output to stdio for analysis
 export MPI_UNBUFFERED_STDIO=true
 cd ~/scratch/dedalus/mri_simulations/
+source png2mp4.sh
 
-mpiexec_mpt -np 512 python3 mri_nonideal_bc.py
-mpiexec_mpt -np 512 python3 -m dedalus merge_procs checkpoints_mri_non
-mpiexec_mpt -np 512 python3 -m dedalus merge_procs slicepoints_mri_non
-mpiexec_mpt -np 512 python3 plot_slices.py checkpoints_mri_wed/*.h5 --output=frames_non
+mpiexec_mpt -np 256 python3 mri_nonideal_bc.py
+mpiexec_mpt -np 256 python3 -m dedalus merge_procs slicepoints_diff2en4
+mpiexec_mpt -np 256 python3 plot_slicepoints.py slicepoints_diff2en4/*.h5 --output=frames_diff2en4
+png2mp4 frames_diff2en4/ mri_diff2en4.mp4 60
+mpiexec_mpt -np 1 python3 plot_ke.py slicepoints_diff2en4/*.h5
+mpiexec_mpt -np 1 python3 plot_be.py slicepoints_diff2en4/*.h5
+mpiexec_mpt -np 256 python3 -m dedalus merge_procs checkpoints_diff2en4

### 2021-08-12 mri_nonideal_bc.py

@@ -19,6 +19,43 @@ import logging
 import pathlib
 logger = logging.getLogger(__name__)
 
+def filter_field(field, frac=0.25):
+    """
+    Filter a field in coefficient space by cutting off all coefficient above
+    a given threshold.  This is accomplished by changing the scale of a field,
+    forcing it into coefficient space at that small scale, then coming back to
+    the original scale.
+    Inputs:
+        field   - The dedalus field to filter
+        frac    - The fraction of coefficients to KEEP POWER IN.  If frac=0.25,
+                    The upper 75% of coefficients are set to 0.
+    """
+    dom = field.domain
+    logger.info("filtering field {} with frac={} using a set-scales approach".format(field.name,frac))
+    orig_scale = field.scales
+    field.set_scales(frac, keep_data=True)
+    field['c']
+    field['g']
+    field.set_scales(orig_scale, keep_data=True)
+    
+def global_noise(domain, seed=42, **kwargs):
+    """
+    Create a field fielled with random noise of order 1.  Modify seed to
+    get varying noise, keep seed the same to directly compare runs.
+    """
+    # Random perturbations, initialized globally for same results in parallel
+    gshape = domain.dist.grid_layout.global_shape(scales=domain.dealias)
+    slices = domain.dist.grid_layout.slices(scales=domain.dealias)
+    rand = np.random.RandomState(seed=seed)
+    noise = rand.standard_normal(gshape)[slices]
+    # filter in k-space
+    noise_field = domain.new_field()
+    noise_field.set_scales(domain.dealias, keep_data=False)
+    noise_field['g'] = noise
+    filter_field(noise_field, **kwargs)
+    return noise_field
+
+
 hardwall = False
 filename = Path('mri_options.cfg')
 outbase = Path("data")
@@ -165,7 +202,7 @@ problem.add_bc("right(bx) = 0", condition="(ny != 0) or (nz != 0)")
 problem.add_bc("right(jxx) = 0", condition="(ny != 0) or (nz != 0)")
 
 # setup
-dt = 1e-6
+dt = 1e-4
[…]

### 2021-08-12 plot_be.py

@@ -1,0 +1,84 @@
+"""
+Plot planes from joint analysis files.
+
+Usage:
+    plot_slices.py <files>... [--output=<dir>] 
+
+Options:
+    --output=<dir>  Output directory [default: ./frames]
+
+"""
+
+import h5py
+import pickle
+import os
+import sys
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+plt.ioff()
+from dedalus.extras import plot_tools
+import publication_settings
+
+matplotlib.rcParams.update(publication_settings.params)
+plt.rcParams.update({'figure.autolayout': True})
+plt.gcf().subplots_adjust(left=0.15)
+
+fig = plt.figure()
+
+def main(filename, start, count, output):
+    """Save plot of specified tasks for given range of analysis writes."""
+
+    # Plot writes
+    with h5py.File(filename, mode='r') as file:
+        be_data = pickle.load(open(path + '/be_data_mri.pick', 'rb'))
+        be_ar = be_data['be_ar']
+        # ke_max_ar = be_data['ke_max_ar']
+        sim_times_ar = be_data['sim_times_ar']
+        for index in range(start, start+count):
+            ke_avg = file['tasks']['be'][index][0, 0, 0]
+            be_ar.append(ke_avg)
+            sim_times_ar.append(file['scales/sim_time'][index])
+        be_data['be_ar'] = be_ar
+        be_data['sim_times_ar'] = sim_times_ar
+        pickle.dump(be_data, open(path + '/be_data_mri.pick', 'wb'))
+        
+
+
+if __name__ == "__main__":
+
+    import pathlib
+    from docopt import docopt
+    from dedalus.tools import logging
+    from dedalus.tools import post
+    from dedalus.tools.parallel import Sync
+
+    global path 
+    path = os.path.dirname(os.path.abspath(__file__))
+    write_data = True
+    plot = True
+    if (write_data):
+        args = docopt(__doc__)
+        output_path = pathlib.Path(args['--output']).absolute()
+        be_data = {'be_ar' : [], 'sim_times_ar' : []}
+        pickle.dump(be_data, open(path + '/be_data_mri.pick', 'wb'))
+
+        # Create output directory if needed
+        with Sync() as sync:
+            if sync.comm.rank == 0:
+                if not output_path.exists():
[…]

### 2021-08-12 plot_ke.py

@@ -1,0 +1,85 @@
+"""
+Plot planes from joint analysis files.
+
+Usage:
+    plot_slices.py <files>... [--output=<dir>] 
+
+Options:
+    --output=<dir>  Output directory [default: ./frames]
+
+"""
+
+import h5py
+import pickle
+import os
+import sys
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+plt.ioff()
+from dedalus.extras import plot_tools
+import publication_settings
+
+matplotlib.rcParams.update(publication_settings.params)
+plt.rcParams.update({'figure.autolayout': True})
+plt.gcf().subplots_adjust(left=0.15)
+
+fig = plt.figure()
+
+def main(filename, start, count, output):
+    """Save plot of specified tasks for given range of analysis writes."""
+
+    # Plot writes
+    with h5py.File(filename, mode='r') as file:
+        ke_data = pickle.load(open(path + '/ke_data_mri.pick', 'rb'))
+        ke_ar = ke_data['ke_ar']
+        # ke_max_ar = ke_data['ke_max_ar']
+        sim_times_ar = ke_data['sim_times_ar']
+        for index in range(start, start+count):
+            ke_avg = file['tasks']['ke'][index][0, 0, 0]
+            ke_ar.append(ke_avg)
+            sim_times_ar.append(file['scales/sim_time'][index])
+        ke_data['ke_ar'] = ke_ar
+        ke_data['sim_times_ar'] = sim_times_ar
+        pickle.dump(ke_data, open(path + '/ke_data_mri.pick', 'wb'))
+        
+
+
+if __name__ == "__main__":
+
+    import pathlib
+    from docopt import docopt
+    from dedalus.tools import logging
+    from dedalus.tools import post
+    from dedalus.tools.parallel import Sync
+
+    global path 
+    path = os.path.dirname(os.path.abspath(__file__))
+    write_data = True
+    plot = True
+    if (write_data):
+        args = docopt(__doc__)
+        output_path = pathlib.Path(args['--output']).absolute()
+        ke_data = {'ke_ar' : [], 'sim_times_ar' : []}
+        pickle.dump(ke_data, open(path + '/ke_data_mri.pick', 'wb'))
+
+        # Create output directory if needed
+        with Sync() as sync:
+            if sync.comm.rank == 0:
[…]

### 2021-08-12 plot_re.py

@@ -1,0 +1,85 @@
+"""
+Plot planes from joint analysis files.
+
+Usage:
+    plot_slices.py <files>... [--output=<dir>] 
+
+Options:
+    --output=<dir>  Output directory [default: ./frames]
+
+"""
+
+import h5py
+import pickle
+import os
+import sys
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+plt.ioff()
+from dedalus.extras import plot_tools
+import publication_settings
+
+matplotlib.rcParams.update(publication_settings.params)
+plt.rcParams.update({'figure.autolayout': True})
+plt.gcf().subplots_adjust(left=0.15)
+
+fig = plt.figure()
+
+def main(filename, start, count, output):
+    """Save plot of specified tasks for given range of analysis writes."""
+
+    # Plot writes
+    with h5py.File(filename, mode='r') as file:
+        ke_data = pickle.load(open(path + '/ke_data_mri.pick', 'rb'))
+        ke_ar = ke_data['ke_ar']
+        # ke_max_ar = ke_data['ke_max_ar']
+        sim_times_ar = ke_data['sim_times_ar']
+        for index in range(start, start+count):
+            ke_avg = file['tasks']['Re'][index][0, 0, 0]
+            ke_ar.append(ke_avg)
+            sim_times_ar.append(file['scales/sim_time'][index])
+        ke_data['ke_ar'] = ke_ar
+        ke_data['sim_times_ar'] = sim_times_ar
+        pickle.dump(ke_data, open(path + '/ke_data_mri.pick', 'wb'))
+        
+
+
+if __name__ == "__main__":
+
+    import pathlib
+    from docopt import docopt
+    from dedalus.tools import logging
+    from dedalus.tools import post
+    from dedalus.tools.parallel import Sync
+
+    global path 
+    path = os.path.dirname(os.path.abspath(__file__))
+    write_data = True
+    plot = True
+    if (write_data):
+        args = docopt(__doc__)
+        output_path = pathlib.Path(args['--output']).absolute()
+        ke_data = {'ke_ar' : [], 'sim_times_ar' : []}
+        pickle.dump(ke_data, open(path + '/ke_data_mri.pick', 'wb'))
+
+        # Create output directory if needed
+        with Sync() as sync:
+            if sync.comm.rank == 0:
[…]

### 2021-08-12 plot_slicepoints.py

@@ -14,6 +14,7 @@ import numpy as np
 import matplotlib
 matplotlib.use('Agg')
 import matplotlib.pyplot as plt
+from matplotlib import ticker
 plt.ioff()
 from dedalus.extras import plot_tools
 
@@ -27,7 +28,7 @@ def midplane(filename, start, count, output):
     """Save plot of specified tasks for given range of analysis writes."""
 
     # Plot settings
-    tasks = ['vx', 'vy', 'vz', 'bx', 'by', 'bz']
+    tasks = ['vx_midz', 'vy_midz', 'vz_midz', 'bx_midz', 'by_midz', 'bz_midz']
     scale = 2.5
     dpi = 100
     title_func = lambda sim_time: 't = {:.3f}'.format(sim_time)
@@ -50,9 +51,27 @@ def midplane(filename, start, count, output):
                 axes = mfig.add_axes(i, j, [0, 0, 1, 1])
                 # Call plotting helper (dset axes: [t, x, y, z])
                 dset = file['tasks'][task]
-                image_axes = (2, 1)
-                data_slices = (index, slice(None), slice(None), 0)
+                image_axes = (3, 1)
+                data_slices = (index, slice(None), 0, slice(None))
+                # if (index % 5 != 0):
+                #     continue
                 plot_tools.plot_bot(dset, image_axes, data_slices, axes=axes, title=task, even_scale=True)
+
+                # x = file['scales/x/1.0'][()]
+                # y = file['scales/y/1.0'][()]
+                # yy, xx = np.meshgrid(y.flatten(), x.flatten())
+                # data = dset[data_slices]
+                # plot = plt.pcolormesh(xx, yy, data)
+                # cbar = plt.colorbar(plot, cax=axes, orientation='horizontal',
+                #     ticks=ticker.MaxNLocator(nbins=5))
+                # cbar.outline.set_visible(False)
+                # axes.xaxis.set_ticks_position('top')                # plt.pcolor(dset[data_slices])
+                # axes.xaxis.set_label_position('top')
+                # plt.xlabel("x")
+                # plt.ylabel("y")
+                # plt.title(task)
+                
+                # plot_tools.plot_bot(dset, image_axes, data_slices, axes=axes, title=task, even_scale=True)
             # Add time title
             title = title_func(file['scales/sim_time'][index])
             title_height = 1 - 0.5 * mfig.margin.top / mfig.fig.y
@@ -60,7 +79,8 @@ def midplane(filename, start, count, output):
[…]

### 2021-08-12 MHD/notes.tex
added txt files

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx,amsmath,amsfonts,amssymb}
\usepackage{enumitem}
\usepackage{xcolor}

\newcommand\Ra{\mathrm{Ra}}
\newcommand\Pran{\mathrm{Pr}}
\newcommand\Rac{\mathrm{Ra}_{\mathrm{c}}}
\newcommand\Ek{\mathrm{Ek}}
\newcommand\Ro{\mathrm{Ro}}
\newcommand\Nu{\mathrm{Nu}}
\newcommand\Sc{\mathrm{Sc}}

\newcommand\eps{\varepsilon}
\renewcommand\L {\mathcal{L}}

\newcommand{\n}{\\ \nonumber \\ }
\newcommand{\nn}{\nonumber}
\newcommand{\nnn}{\\ \nonumber \\ \nonumber}

\newcommand\ie{\textit{i.e.},~}
\newcommand\eg{\textit{e.g.},~}
\newcommand{\omicron}{o}

\newcommand{\pd}[1]{\partial_{#1}}
\renewcommand{\vec}[1]{\boldsymbol{#1}}
\newcommand{\M}[1]{\mathbf{#1}}
\newcommand{\grad}{\vec{\nabla}}
\newcommand{\cross}{\vec{\times}}
\newcommand{\curl}{\grad \vec{\times}}
\newcommand{\laplacian}{\nabla^2}
\newcommand{\veclaplacian}{\grad^2}

\newcommand{\sump}[2]{\sideset{}{'}\sum_{{#1}=0}^{#2}}

\newcommand{\eq}[1]{(\ref{#1})}
\newcommand{\eqs}[2]{(\ref{#1})~\&~(\ref{#2})}
\newcommand{\eqss}[2]{(\ref{#1})--(\ref{#2})}

\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Eqs}[2]{Eqs.~(\ref{#1})~\&~(\ref{#2})}
\newcommand{\Eqss}[2]{Eqs.~(\ref{#1})--(\ref{#2})}

\newcommand{\fig}[1]{Fig.~(\ref{#1})}
\newcommand{\figs}[2]{Figs.~(\ref{#1})~\&~(\ref{#2})}
\newcommand{\T}{{\cal T}}
\newcommand{\Z}{{\cal Z}}

\setlength{\parindent}{0pt}

\begin{document}

\section{MRI Notes}
\subsection{Daniel (8/6/2021)}
\subsubsection{Potential Form}
The induction equation is given by
\begin{align*}
    \partial_t\vec{b} &= \grad \cross (\vec{u} \cross \vec{b}) + \eta \grad^2 \vec{b}
    \intertext{Using the following identity}
    \grad \cross \grad \cross \vec{f} &= \grad \grad \cdot \vec{f} - \grad^2\vec{f}
    \intertext{and assumming $\eta$ to be constant, we use $\grad \cdot \vec{b} = 0$, giving}
    \partial_t\vec{b} &= \grad \cross (\vec{u} \cross \vec{b}) - \eta \grad \cross \grad \cross \vec{b}.
    \intertext{Then we define a vector potential $\grad \cross \vec{A} \equiv \vec{b}$, yielding}
    \partial_t \grad \cross \vec{A} &= \grad \cross (\vec{u} \cross \vec{b}) - \grad \cross (\eta \grad \cross \vec{b}) \\
    \partial_t \vec{A} &= \vec{u} \cross \vec{b} - \eta \grad \cross \vec{b} + \grad \phi
    \intertext{where $\phi$ is a scalar potential arizing from ``uncurling'' the equation. We must then provide an additional constraint to fix $\phi$: the Coulomb gauge $\grad \cdot \vec{A} = 0$. Therefore}
    -\grad \cross \vec{b} &= - \curl \grad \cross \vec{A} = \veclaplacian \vec{A}.
\end{align*}
Next we decompose $\vec{u} = \vec{u_0} + \vec{u'}$ and $\vec{b} = \vec{b_0} + \vec{b'}$.
We assume the mean quantities $\vec{u_0}$ and $\vec{b_0}$ are themselves solutions to the original problem. 
If we consider only the 0th mode of $\vec{b}$, i.e. $\vec{b} \cdot \vec{\hat{e}_i} \sim e^{i0}$ then clearly $\veclaplacian \vec{b} = \vec{0}$ and therefore
\begin{align*}
    \partial_t \vec{b'} &= \curl (\vec{u_0} \cross \vec{b'}) + \curl (\vec{u'} \cross \vec{b_0}) + \curl (\vec{u'} \cross \vec{b'}).
    \intertext{Using another identity}
    \curl (\vec{A} \cross \vec{B}) &= \vec{A} \grad \cdot \vec{B} - \vec{B} \grad \cdot \vec{A} + \vec{B} \cdot \grad \vec{A} - \vec{A} \cdot \grad \vec{B}
\end{align*}
\end{document}

### 2021-08-18 mri_nonideal.sh
linear mri working

@@ -1,7 +1,7 @@
 #PBS -S /bin/bash
 #PBS -N mri_nonideal_diff2en4
-#PBS -l select=10:ncpus=28:mpiprocs=256:model=bro
-#PBS -l walltime=24:00:00
+#PBS -l select=1:ncpus=40:mpiprocs=40:model=sky_ele
+#PBS -l walltime=16:00:00
 #PBS -j oe
 module load mpi-sgi/mpt
 module load comp-intel
@@ -16,10 +16,10 @@ export MPI_UNBUFFERED_STDIO=true
 cd ~/scratch/dedalus/mri_simulations/
 source png2mp4.sh
 
-mpiexec_mpt -np 256 python3 mri_nonideal_bc.py
-mpiexec_mpt -np 256 python3 -m dedalus merge_procs slicepoints_diff2en4
-mpiexec_mpt -np 256 python3 plot_slicepoints.py slicepoints_diff2en4/*.h5 --output=frames_diff2en4
+mpiexec_mpt -np 32 python3 mri_nonideal_bc.py
+mpiexec_mpt -np 32 python3 -m dedalus merge_procs slicepoints_diff2en4
+mpiexec_mpt -np 32 python3 plot_slicepoints.py slicepoints_diff2en4/*.h5 --output=frames_diff2en4
 png2mp4 frames_diff2en4/ mri_diff2en4.mp4 60
 mpiexec_mpt -np 1 python3 plot_ke.py slicepoints_diff2en4/*.h5
 mpiexec_mpt -np 1 python3 plot_be.py slicepoints_diff2en4/*.h5
-mpiexec_mpt -np 256 python3 -m dedalus merge_procs checkpoints_diff2en4
+mpiexec_mpt -np 32 python3 -m dedalus merge_procs checkpoints_diff2en4

### 2021-08-18 mri_nonideal_bc.py

@@ -99,8 +99,8 @@ cutoff =  kx*np.sqrt(R**2 - 1)
 # Create bases and domain
 # Use COMM_SELF so keep calculations independent between processes
 x_basis = de.Chebyshev('x', Nx, interval=(-Lx/2, Lx/2))
-y_basis = de.Fourier('y', Ny, interval=(0, Lx))
-z_basis = de.Fourier('z', Nz, interval=(0, Lx))
+y_basis = de.Fourier('y', Ny, interval=(0, Lx * 8))
+z_basis = de.Fourier('z', Nz, interval=(0, Lx * 8))
 
 ncpu = MPI.COMM_WORLD.size
 log2 = np.log2(ncpu)
@@ -202,7 +202,7 @@ problem.add_bc("right(bx) = 0", condition="(ny != 0) or (nz != 0)")
 problem.add_bc("right(jxx) = 0", condition="(ny != 0) or (nz != 0)")
 
 # setup
-dt = 1e-4
+dt = 1e-3
 solver = problem.build_solver(de.timesteppers.SBDF2)
 
 
@@ -243,10 +243,10 @@ else:
     stop_sim_time = 20
     fh_mode = 'append'
 
-checkpoints = solver.evaluator.add_file_handler('checkpoints_diff2en4', sim_dt=0.1, max_writes=1, mode=fh_mode)
+checkpoints = solver.evaluator.add_file_handler('checkpoints_diff2en4_slice', sim_dt=0.1, max_writes=1, mode=fh_mode)
 checkpoints.add_system(solver.state)
 
-slicepoints = solver.evaluator.add_file_handler('slicepoints_diff2en4', sim_dt=0.005, max_writes=50, mode=fh_mode)
+slicepoints = solver.evaluator.add_file_handler('slicepoints_diff2en4_slice', sim_dt=0.005, max_writes=50, mode=fh_mode)
 
 slicepoints.add_task("interp(vx, y={})".format(Lx / 2), name="vx_midy")
 slicepoints.add_task("interp(vx, z={})".format(Lx / 2), name="vx_midz")
@@ -285,7 +285,7 @@ flow = flow_tools.GlobalFlowProperty(solver, cadence=10)
 flow.add_property("sqrt(vx*vx + vy*vy + vz*vz)", name='Re')
 
 solver.stop_sim_time = 30
-solver.stop_wall_time = 48 * 60 * 60.
+solver.stop_wall_time = 1 * 60 * 60.
 solver.stop_iteration = np.inf
 
 try:

### 2021-08-18 mri_single_yz_mode.py

@@ -16,7 +16,7 @@ import time
 from configparser import ConfigParser
 from pathlib import Path
 import numpy as np
-import os
+import sys
 import h5py
 from eigentools import Eigenproblem
 import dedalus.public as de
@@ -28,8 +28,8 @@ logger = logging.getLogger(__name__)
 args = docopt(__doc__)
 ky = float(args['<ky>'])
 kz = float(args['<kz>'])
-ideal = args['--ideal']
-hardwall = args['--hardwall']
+ideal = False
+hardwall = False
 filename = Path(args['<config_file>'])
 outbase = Path("data")
 
@@ -174,17 +174,21 @@ if not ideal:
 
 
 # GO
-EP = Eigenproblem(problem,sparse=False)
+EP = Eigenproblem(problem)
 t1 = time.time()
-gr, idx, freq = EP.growth_rate({})
+gr, idx, freq = EP.growth_rate(sparse=False)
 t2 = time.time()
 logger.info("Solve time: {}".format(t2-t1))
 
 gamma = EP.evalues_good
-eigenvectors = EP.solver.eigenvectors[:,EP.evalues_good_index]
-index = np.argsort(-gamma.real)
 
-logger.info("saving eigenvector with gamma = {}".format(gamma[index[0]]))
+# eigenvectors = EP.solver.eigenvectors[:,EP.evalues_good_index]
+index = np.argsort(-gamma.real)
+logger.info("eigenvales: " + str(gamma[index]))
+
+
+logger.info("Solve complete. gamma = {}".format(gamma[index[0]]))
+sys.exit()
 nvars = len(problem_variables)
 eigvec = np.zeros((nvars,Nx),dtype=np.complex128)
 EP.solver.set_state(index[0])

### 2021-08-18 mri_single_yz_mode_og.py

@@ -1,0 +1,239 @@
+"""
+The magnetorotational instability prefers three dimensions.
+
+Single mode calculation. Returns spectrum.
+
+Usage:
+    mri_single_yz_mode.py [--ideal] <config_file> <ky> <kz>
+
+Options:
+    --ideal    Use Ideal MHD
+    --hardwall Use experimental boundary conditions
+"""
+
+from docopt import docopt
+import time
+from configparser import ConfigParser
+from pathlib import Path
+import numpy as np
+import os
+import h5py
+from eigentools import Eigenproblem
+import dedalus.public as de
+from mpi4py import MPI
+CW = MPI.COMM_WORLD
+import logging
+logger = logging.getLogger(__name__)
+
+args = docopt(__doc__)
+ky = float(args['<ky>'])
+kz = float(args['<kz>'])
+ideal = args['--ideal']
+hardwall = args['--hardwall']
+filename = Path(args['<config_file>'])
+outbase = Path("data")
+
+logger.info("Solving for mode ky = {}, kz = {}".format(ky, kz))
+# Parse .cfg file to set global parameters for script
+config = ConfigParser()
+config.read(str(filename))
+
+logger.info('Running mri.py with the following parameters:')
+logger.info(config.items('parameters'))
+try:
+    dense = config.getboolean('solver','dense')
+except:
+    dense = False
+if dense:
+    logger.info("Using dense solver.")
+    dense_threshold = config.getfloat('solver','dense_threshold')
+else:
+    logger.info("Using sparse solver.")
+
+Nx = config.getint('parameters','Nx')
+Lx = eval(config.get('parameters','Lx'))
+B = config.getfloat('parameters','B')
+
+Nmodes = config.getint('parameters','Nmodes')
+
+R      =  config.getfloat('parameters','R')
+q      =  config.getfloat('parameters','q')
+
+kymin = config.getfloat('parameters','kymin')
+kymax = config.getfloat('parameters','kymax')
+Nky = config.getint('parameters','Nky')
+
+kzmin = config.getfloat('parameters','kzmin')
[…]

### 2021-08-18 plot_be.py

@@ -13,11 +13,12 @@ import h5py
 import pickle
 import os
 import sys
+import scipy.stats
 import numpy as np
 import matplotlib
 import matplotlib.pyplot as plt
 plt.ioff()
-from dedalus.extras import plot_tools
+# from dedalus.extras import plot_tools
 import publication_settings
 
 matplotlib.rcParams.update(publication_settings.params)
@@ -49,36 +50,56 @@ if __name__ == "__main__":
 
     import pathlib
     from docopt import docopt
-    from dedalus.tools import logging
-    from dedalus.tools import post
-    from dedalus.tools.parallel import Sync
+    # from dedalus.tools import logging
+    # from dedalus.tools import post
+    # from dedalus.tools.parallel import Sync
 
     global path 
     path = os.path.dirname(os.path.abspath(__file__))
-    write_data = True
+    write_data = False
     plot = True
-    if (write_data):
-        args = docopt(__doc__)
-        output_path = pathlib.Path(args['--output']).absolute()
-        be_data = {'be_ar' : [], 'sim_times_ar' : []}
-        pickle.dump(be_data, open(path + '/be_data_mri.pick', 'wb'))
+    regress = True
+    # if (write_data):
+    #     args = docopt(__doc__)
+    #     output_path = pathlib.Path(args['--output']).absolute()
+    #     be_data = {'be_ar' : [], 'sim_times_ar' : []}
+    #     pickle.dump(be_data, open(path + '/be_data_mri.pick', 'wb'))
 
-        # Create output directory if needed
-        with Sync() as sync:
-            if sync.comm.rank == 0:
-                if not output_path.exists():
-                    output_path.mkdir()
-        post.visit_writes(args['<files>'], main, output=output_path)
+    #     # Create output directory if needed
+    #     with Sync() as sync:
+    #         if sync.comm.rank == 0:
+    #             if not output_path.exists():
+    #                 output_path.mkdir()
+        # post.visit_writes(args['<files>'], main, output=output_path)
     if (plot):
         be_data = pickle.load(open(path + '/be_data_mri.pick', 'rb'))
         be_ar = be_data['be_ar']
         sim_times_ar = be_data['sim_times_ar']
         colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
[…]

### 2021-08-18 plot_ke.py

@@ -19,6 +19,7 @@ import matplotlib.pyplot as plt
 plt.ioff()
 from dedalus.extras import plot_tools
 import publication_settings
+import scipy.stats
 
 matplotlib.rcParams.update(publication_settings.params)
 plt.rcParams.update({'figure.autolayout': True})
@@ -55,8 +56,9 @@ if __name__ == "__main__":
 
     global path 
     path = os.path.dirname(os.path.abspath(__file__))
-    write_data = True
+    write_data = False
     plot = True
+    regress = True
     if (write_data):
         args = docopt(__doc__)
         output_path = pathlib.Path(args['--output']).absolute()
@@ -70,16 +72,39 @@ if __name__ == "__main__":
                     output_path.mkdir()
         post.visit_writes(args['<files>'], main, output=output_path)
     if (plot):
+        fig = plt.figure()
         ke_data = pickle.load(open(path + '/ke_data_mri.pick', 'rb'))
         ke_ar = ke_data['ke_ar']
         sim_times_ar = ke_data['sim_times_ar']
         colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
-        plt.plot(sim_times_ar, ke_ar, color=colors[-1])
+        plt.plot(sim_times_ar, ke_ar, color=colors[-1], label='Simulation')
         plt.yscale('log')
+        if (regress):
+            ln_ke_ar = np.log(ke_ar)
+            t_cutoff = 25
+            i_cutoff = -1
+            for i, t in enumerate(sim_times_ar):
+                if (t > t_cutoff):
+                    i_cutoff = i
+                    break
+            
+            t_lin = sim_times_ar[i:]
+            ke_lin = ln_ke_ar[i:]
+            slope, intercept, r_value, p_value, std_err = scipy.stats.linregress(t_lin, ke_lin)
+            print('R2 val: ' + str(r_value**2))
+            print('slope: ' + str(slope))
+            print('intercept: ' + str(intercept))      
+            ke_reg = np.exp(slope * np.array(t_lin) + intercept)
+            plt.plot(t_lin, ke_reg, color='k', linestyle='--', linewidth=2, label='Best fit')
+
+            fig.text(.2, .05, 'Growth rate = ' + str(round(slope, 4)), ha='center')
+            plt.legend()
+
+
         # plt.xlim(0, 400)
[…]

### 2021-08-18 plot_slicepoints.py

@@ -79,7 +79,7 @@ def midplane(filename, start, count, output):
             # Save figure
             savename = savename_func(file['scales/write_number'][index])
             savepath = output.joinpath(savename)
-            if (index % 1 == 0):
+            if (index % 4 == 0):
                 fig.savefig(str(savepath), dpi=dpi)
             fig.clear()
     plt.close(fig)

### 2021-08-18 MHD/notes.tex

@@
 \newcommand{\pd}[1]{\partial_{#1}}
+\newcommand{\vhat}[1]{\hat{\boldsymbol{#1}}}
 \renewcommand{\vec}[1]{\boldsymbol{#1}}
@@
 \end{align*}
+
+\subsection*{Momentum Equation}
+Verbatim from Jeff Oishi's MRI paper:
+\begin{align*}
+    \frac{D\vec{u'}}{Dt} + f\vhat{z} \cross\vec{u'} + Su_x' \vhat{y} + \grad p' + \nu \curl \omega' &= B_0 \partial_z \vec{b'}
+    \intertext{where $f$ is the corriolis parameter, S is the background shearing rate, and $B_0 \vhat{z}$ is a uniform background magnetic field.
+This equation is linearized wrt perturbations.
+Accordingly, the material derivative goes like}
+    \frac{D}{Dt} &\equiv \partial_t + \vec{u} \cdot \grad \\
+    &= \partial_t + Sx\partial_y
+    \intertext{due to the background velocity $\vec{\overline{u}} = Sx\vhat{y}$. 
+    In the nonlinear case we have}
+    &= \partial_t + (Sx\vhat{y} + \vec{u'}) \cdot \grad
+    \intertext{From inspection and stuff, the irrotational momentum equation goes like}
+    \frac{D\vec{u}}{Dt} + \grad p + \nu\cross \omega &= \vec{b} \cdot \grad\vec{b}
+    \intertext{Next we generalize $\vec{u} = \vec{u'} + Sx\vhat{y}$ and $\vec{b} = \vec{b'} + B_0\vhat{z}$, giving}
+    \underline{\partial_t \vec{u'} + \vec{u'} \cdot \grad \vec{u'} + Sx\partial_y \vec{u'} + Su'_x \vhat{y}} + \grad p + \nu \curl \omega &= B_0 \partial_z\vec{b'} + \vec{b'} \cdot \grad \vec{b'}
+    \intertext{where the material derivative $\frac{D\vec{u}}{Dt}$ consists of the underlined terms}
+\end{align*}
+
 \end{document}

### 2021-08-19 MHD/notes.tex
update notes
nonlinear notes and script

@@
 \newcommand{\curl}{\grad \vec{\times}}
+\newcommand{\divergence}{\grad \cdot}
 \newcommand{\laplacian}{\nabla^2}
@@
 
-\section{MRI Notes}
-\subsection{Daniel (8/6/2021)}
-\subsubsection{Potential Form}
+\section*{MRI Notes}
+\subsection*{Induction Equation: Potential Form}
 The induction equation is given by
@@
 
-\subsection*{Momentum Equation}
+\subsection*{Momentum Equation: Nonlinear Terms}
 Verbatim from Jeff Oishi's MRI paper:
 \begin{align*}
-    \frac{D\vec{u'}}{Dt} + f\vhat{z} \cross\vec{u'} + Su_x' \vhat{y} + \grad p' + \nu \curl \omega' &= B_0 \partial_z \vec{b'}
+    \frac{D\vec{u'}}{Dt} + f\vhat{z} \cross\vec{u'} + Su_x' \vhat{y} + \grad p' + \nu \curl \vec{\omega}' &= B_0 \partial_z \vec{b'}
     \intertext{where $f$ is the corriolis parameter, S is the background shearing rate, and $B_0 \vhat{z}$ is a uniform background magnetic field.
@@
     \intertext{From inspection and stuff, the irrotational momentum equation goes like}
-    \frac{D\vec{u}}{Dt} + \grad p + \nu\cross \omega &= \vec{b} \cdot \grad\vec{b}
+    \frac{D\vec{u}}{Dt} + \grad p + \nu\cross \vec{\omega} &= \vec{b} \cdot \grad\vec{b}
     \intertext{Next we generalize $\vec{u} = \vec{u'} + Sx\vhat{y}$ and $\vec{b} = \vec{b'} + B_0\vhat{z}$, giving}
-    \underline{\partial_t \vec{u'} + \vec{u'} \cdot \grad \vec{u'} + Sx\partial_y \vec{u'} + Su'_x \vhat{y}} + \grad p + \nu \curl \omega &= B_0 \partial_z\vec{b'} + \vec{b'} \cdot \grad \vec{b'}
+    \underline{\partial_t \vec{u'} + \vec{u'} \cdot \grad \vec{u'} + Sx\partial_y \vec{u'} + Su'_x \vhat{y}} + \grad p + \nu \curl \vec{\omega} &= B_0 \partial_z\vec{b'} + \vec{b'} \cdot \grad \vec{b'}
     \intertext{where the material derivative $\frac{D\vec{u}}{Dt}$ consists of the underlined terms}
@@
 
+\subsection*{Induction Equation: Nonlinear Terms}
+The MHD induction equation (as given by \textit{Fluid Mechanics of Planets and Stars, 2019}) is given by
+\begin{align*}
+    \partial_t \vec{b} &= \underline{\curl (\vec{u} \cross \vec{b})} + \eta \laplacian \vec{b}.
+    \intertext{We expand the underlined term using the following identity}
+    \curl(\vec{u}\cross\vec{b}) &= \vec{u}\divergence\vec{b} - \vec{b}\divergence\vec{u} + \vec{b}\cdot\grad\vec{u} - \vec{u}\cdot\grad\vec{b}
+    \intertext{where the first two terms on the RHS vanish due to incompressibility and the abscence of magnetic monopoles. 
+    Therefore}
+    \partial_t \vec{b} + \vec{u} \cdot \grad\vec{b} &= \vec{b}\cdot\grad\vec{u} + \eta\laplacian\vec{b}.
+    \intertext{Substituting the decompositions for $\vec{u}$ and $\vec{b}$ as above yields}
+    \partial_t (B_0\vhat{z} + \vec{b'}) + (Sx\vhat{y} + \vec{u'}) \cdot \grad(B_0\vhat{z} + \vec{b'}) &= (B_0\vhat{z} + \vec{b'})\cdot\grad(Sx\vhat{y} + \vec{u'}) + \eta\laplacian(B_0\vhat{z} + \vec{b'}) \\
+    \partial_t \vec{b'} + (Sx\vhat{y} + \vec{u'}) \cdot \grad \vec{b'} &= (B_0\vhat{z} + \vec{b'})\cdot\grad(Sx\vhat{y} + \vec{u'}) + \eta\laplacian\vec{b'} 
+    \intertext{dropping the $'$s and taking the $x,y,$ and $z$ components of the above yields}
+    \partial_t b_x + Sx \partial_y b_x + \vec{u} \cdot \grad b_x &= B_0\partial_z u_x + \vec{b} \cdot \grad u_x + \eta\laplacian b_x \\
+    \partial_t b_y + Sx \partial_y b_y + \vec{u} \cdot \grad b_y &= B_0\partial_z u_y + \vec{b} \cdot \grad u_y + S b_x + \eta\laplacian b_y \\
+    \partial_t b_z + Sx \partial_y b_z + \vec{u} \cdot \grad b_z &= B_0\partial_z u_z + \vec{b} \cdot \grad u_z + \eta\laplacian b_z 
+    \intertext{Note that the linear operator on the LHS is identical for each scalar equation. Accordingly, we redefine $D_t$ to be this operator, i.e.}
+    D_tA &\equiv \partial_t A + Sx \partial_y A + \vec{u} \cdot \grad A.
+    \intertext{Recall that the current density is given by}
+    \vec{j} = j_x\vhat{x} + j_y\vhat{y} + j_z\vhat{z} &= \curl \vec{b} \\
+    j_x &= \partial_y b_z - \partial_z b_y
+    \intertext{We then rewrite the laplacian term as follows}
+    \eta\laplacian \vec{b} &= -\curl\curl \vec{b} = -\curl \vec{j}.
+    \intertext{Therefore the $x$-component of the induction equation is given by}
+    D_t b_x - B_0\partial_z u_x + \eta(\partial_yj_z - \partial_zj_y) &= \vec{b} \cdot \grad u_x \\
+    \intertext{Finally, we combine the $x$ and $y$ components of the induction equation. Taking the material derivative of the $x$-component of the current density gives}
+    D_t j_x &= \partial_y D_t b_z - \partial_z D_t b_y.
+\end{align*}
+\begin{align*}
+    \intertext{We have equations for that stuff!}
+    D_t j_x &= \partial_y (B_0\partial_z u_z + \vec{b} \cdot \grad u_z + \eta\laplacian b_z) - \partial_z (B_0\partial_z u_y + \vec{b} \cdot \grad u_y + S b_x + \eta\laplacian b_y) \\
+    &= B_0 \partial_z \omega_x + \vec{b} \cdot \grad \omega_x + \eta\laplacian j_x - S\partial_z b_x
+\end{align*}
+Rearranging and expanding gives\\
+
+$\partial_t j_x + Sx\partial_y j_x - B_0 \partial_z \omega_x + S\partial_z b_x - \eta\laplacian j_x = \vec{b} \cdot \grad \omega_x - \vec{u} \cdot \grad j_x$
+
 \end{document}

### 2021-08-19 mri_nonlin/mri_nonlin.py

@@ -19,6 +19,8 @@ import logging
 import pathlib
 logger = logging.getLogger(__name__)
 
+
+##### Initial condition functions from Evan H. Anders
 def filter_field(field, frac=0.25):
     """
     Filter a field in coefficient space by cutting off all coefficient above
@@ -40,7 +42,7 @@ def filter_field(field, frac=0.25):
     
 def global_noise(domain, seed=42, **kwargs):
     """
-    Create a field fielled with random noise of order 1.  Modify seed to
+    Create a field filled with random noise of order 1.  Modify seed to
     get varying noise, keep seed the same to directly compare runs.
     """
     # Random perturbations, initialized globally for same results in parallel
@@ -74,19 +76,8 @@ Nz = eval(config.get('parameters','Nz'))
 Lx = eval(config.get('parameters','Lx'))
 
 B = config.getfloat('parameters','B')
-
-Nmodes = config.getint('parameters','Nmodes')
-
 R      =  config.getfloat('parameters','R')
 q      =  config.getfloat('parameters','q')
-
-kymin = config.getfloat('parameters','kymin')
-kymax = config.getfloat('parameters','kymax')
-Nky = config.getint('parameters','Nky')
-
-kzmin = config.getfloat('parameters','kzmin')
-kzmax = config.getfloat('parameters','kzmax')
-Nkz = config.getint('parameters','Nkz')
 
 ν = config.getfloat('parameters','ν')
 η = config.getfloat('parameters','η')
@@ -152,14 +143,13 @@ problem.add_equation("ωz - dx(vy) + dy(vx) = 0")
 
 # MHD equations: bx, by, bz, jxx
 problem.add_equation("dx(bx) + dy(by) + dz(bz) = 0", condition='(ny != 0) or (nz != 0)')
-problem.add_equation("Dt(bx) - B*dz(vx) + η*( dy(jz) - dz(jy) )            = bz*dz(vx) - v_dot_grad(bx)", condition='(ny != 0) or (nz != 0)')
-problem.add_equation("Dt(jx) - B*dz(ωx) + S*dz(bx) - η*( dx(jxx) + L(jx) ) = bz*dz(ωx) - v_dot_grad(ωx)", condition='(ny != 0) or (nz != 0)')
[…]

### 2021-08-19 mri_nonlin/mri_nonlin.py
cleanup, cleanup, everybody everywhere

@@ -84,7 +84,7 @@ q      =  config.getfloat('parameters','q')
 
 kx     =  np.pi/Lx
 S      = -R*B*kx*np.sqrt(q)
-f      =  R*B*kx/np.sqrt(q)
+f      =  0 *        R*B*kx/np.sqrt(q)
 cutoff =  kx*np.sqrt(R**2 - 1)
 
 # Create bases and domain
@@ -134,8 +134,8 @@ problem.substitutions['L(A)'] = "dy(dy(A)) + dz(dz(A))"
 # Variable substitutions
 problem.add_equation("dx(vx) + dy(vy) + dz(vz) = 0")
 
-problem.add_equation("Dt(vx) -     f*vy + dx(p) - B*dz(bx) + ν*(dy(ωz) - dz(ωy)) = b_dot_grad(bx) - v_dot_grad(vx)")
-problem.add_equation("Dt(vy) + (f+S)*vx + dy(p) - B*dz(by) + ν*(dz(ωx) - dx(ωz)) = b_dot_grad(by) - v_dot_grad(vy)")
+problem.add_equation("Dt(vx)            + dx(p) - B*dz(bx) + ν*(dy(ωz) - dz(ωy)) = b_dot_grad(bx) - v_dot_grad(vx)")
+problem.add_equation("Dt(vy) + S*vx     + dy(p) - B*dz(by) + ν*(dz(ωx) - dx(ωz)) = b_dot_grad(by) - v_dot_grad(vy)")
 problem.add_equation("Dt(vz)            + dz(p) - B*dz(bz) + ν*(dx(ωy) - dy(ωx)) = b_dot_grad(bz) - v_dot_grad(vz)")
 
 problem.add_equation("ωy - dz(vx) + dx(vz) = 0")

### 2021-08-19 mri_nonlin/mri_nonlin.py
delete redundant files

@@ -194,7 +194,7 @@ problem.add_bc("right(bx) = 0", condition="(ny != 0) or (nz != 0)")
 problem.add_bc("right(jxx) = 0", condition="(ny != 0) or (nz != 0)")
 
 # setup
-dt = 1e-4
+dt = 1e-3
 solver = problem.build_solver(de.timesteppers.SBDF2)
 
 

### 2021-09-03 mri_nonlin/mri_nonlin.sh
opening night

@@ -1,0 +1,33 @@
+#PBS -S /bin/bash
+#PBS -N mri_nonlin_diff1en3
+#PBS -l select=5:ncpus=28:mpiprocs=28:model=bro
+#PBS -l walltime=24:00:00
+#PBS -j oe
+
+module load mpi-sgi/mpt
+module load comp-intel
+export PATH=$HOME/scripts:$PATH
+deactivate
+unset PYTHONPATH
+source ~/miniconda3/etc/profile.d/conda.sh
+export PYTHONNOUSERSITE=1
+conda activate dedalus
+# support lots of text output to stdio for analysis
+export MPI_UNBUFFERED_STDIO=true
+
+source ~/png2mp4.sh
+cd ~/scratch/dedalus/mri_simulations/mri_nonlin
+
+SUFF="diff1en3_cu"
+DIFF=0.001
+MPIPROC=128
+
+mpiexec_mpt -np $MPIPROC python3 mri_nonlin.py $SUFF $DIFF
+mpiexec_mpt -np $MPIPROC python3 -m dedalus merge_procs scalars_${SUFF}
+mpiexec_mpt -np 1 python3 plot_ke.py scalars_${SUFF}/*.h5 --suffix=$SUFF
+mpiexec_mpt -np 1 python3 plot_be.py scalars_${SUFF}/*.h5 --suffix=$SUFF
+mpiexec_mpt -np $MPIPROC python3 -m dedalus merge_procs slicepoints_${SUFF}
+mpiexec_mpt -np $MPIPROC python3 plot_slicepoints.py slicepoints_${SUFF}/*.h5 --output=frames_${SUFF}
+png2mp4 frames_${SUFF}/ mri_${SUFF}.mp4 60
+mpiexec_mpt -np $MPIPROC python3 -m dedalus merge_procs checkpoints_${SUFF}
+

### 2021-09-03 mri_nonlin/mri_nonlin_hires.py

@@ -9,7 +9,9 @@ from configparser import ConfigParser
 from pathlib import Path
 import numpy as np
 import os
+import sys
 import h5py
+import gc
 import dedalus.public as de
 from dedalus.extras import flow_tools
 from dedalus.extras.plot_tools import plot_bot_2d
@@ -19,6 +21,8 @@ import logging
 import pathlib
 logger = logging.getLogger(__name__)
 
+
+##### Initial condition functions from Evan H. Anders
 def filter_field(field, frac=0.25):
     """
     Filter a field in coefficient space by cutting off all coefficient above
@@ -36,11 +40,13 @@ def filter_field(field, frac=0.25):
     field.set_scales(frac, keep_data=True)
     field['c']
     field['g']
+    # field.require_coeff_space()
+    # field.require_grid_space()
     field.set_scales(orig_scale, keep_data=True)
     
 def global_noise(domain, seed=42, **kwargs):
     """
-    Create a field fielled with random noise of order 1.  Modify seed to
+    Create a field filled with random noise of order 1.  Modify seed to
     get varying noise, keep seed the same to directly compare runs.
     """
     # Random perturbations, initialized globally for same results in parallel
@@ -55,14 +61,25 @@ def global_noise(domain, seed=42, **kwargs):
     filter_field(noise_field, **kwargs)
     return noise_field
 
-
+run_suffix = 'diff1en3_crctd'
+restart_hires = False
+if restart_hires:
+    hires_factor = 1
+else:
+    hires_factor = 1
 hardwall = False
-filename = Path('mri_options.cfg')
-outbase = Path("data")
+filename = Path('mri_.cfg')
 
 # Parse .cfg file to set global parameters for script
 config = ConfigParser()
 config.read(str(filename))
+if len(sys.argv) > 1:
+    run_suffix = sys.argv[1]
+    logger.info("suffix provided for write data: " + run_suffix)
+else:
[…]

### 2021-09-03 mri_nonlin/mri_nonlin_prerf.py

@@ -62,19 +62,27 @@ def global_noise(domain, seed=42, **kwargs):
     return noise_field
 
 run_suffix = 'diff1en3_crctd'
+restart_hires = False
+if restart_hires:
+    hires_factor = 1
+else:
+    hires_factor = 1
 hardwall = False
-filename = Path('mri_options.cfg')
+filename = Path('mri_options_diff1en3.cfg')
+outbase = Path("data")
 
 # Parse .cfg file to set global parameters for script
 config = ConfigParser()
 config.read(str(filename))
-if len(sys.argv) > 1:
+if len(sys.argv) != 1:
     run_suffix = sys.argv[1]
     logger.info("suffix provided for write data: " + run_suffix)
 else:
     logger.warning("write data suffix not found")
     logger.info("cmd arguments: " + str(sys.argv))
     logger.info("Using default suffix \'" + str(run_suffix) + "\'")
+
+
 
 logger.info('Running mri.py with the following parameters:')
 logger.info(config.items('parameters'))
@@ -92,17 +100,10 @@ q      =  config.getfloat('parameters','q')
 ν = config.getfloat('parameters','ν')
 η = config.getfloat('parameters','η')
 
-if len(sys.argv) > 2:
-    diffusivities = float(sys.argv[2])
-    ν = η = diffusivities
-    logger.info("diffusivities provided in cmd args. nu = eta = " + str(diffusivities))
-else:
-    logger.warning("diffusivities not provided in cmd args")
-    logger.info("cmd arguments: " + str(sys.argv))
-    logger.info("Using nu, eta = " + str(ν) + ", " + str(η))
-
-S      = -R*B*np.sqrt(q)
-f      =  R*B/np.sqrt(q)
+kx     =  np.pi/Lx
+S      = -R*B*kx*np.sqrt(q)
+f      =  R*B*kx/np.sqrt(q)
+cutoff =  kx*np.sqrt(R**2 - 1)
 
 # Create bases and domain
 # Use COMM_SELF so keep calculations independent between processes
[…]

### 2021-09-03 mri_nonlin/plot_be.py

@@ -1,0 +1,107 @@
+"""
+Plot planes from joint analysis files.
+
+Usage:
+    plot_slices.py <files>... [--suffix=write_string_suffix] 
+
+"""
+
+import h5py
+import pickle
+import os
+import sys
+import scipy.stats
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+plt.ioff()
+# from dedalus.extras import plot_tools
+import publication_settings
+
+matplotlib.rcParams.update(publication_settings.params)
+plt.rcParams.update({'figure.autolayout': True})
+plt.gcf().subplots_adjust(left=0.15)
+
+def main(filename, start, count, output):
+    """Save plot of specified tasks for given range of analysis writes."""
+
+    # Plot writes
+    with h5py.File(filename, mode='r') as file:
+        be_data = pickle.load(open(path + '/be_data_mri.pick', 'rb'))
+        be_ar = be_data['be_ar']
+        # ke_max_ar = be_data['ke_max_ar']
+        sim_times_ar = be_data['sim_times_ar']
+        for index in range(start, start+count):
+            ke_avg = file['tasks']['be'][index][0, 0, 0]
+            be_ar.append(ke_avg)
+            sim_times_ar.append(file['scales/sim_time'][index])
+        be_data['be_ar'] = be_ar
+        be_data['sim_times_ar'] = sim_times_ar
+        pickle.dump(be_data, open(path + '/be_data_mri.pick', 'wb'))
+        
+
+
+if __name__ == "__main__":
+
+    import pathlib
+    from docopt import docopt
+    from dedalus.tools import logging
+    from dedalus.tools import post
+    from dedalus.tools.parallel import Sync
+
+    global path 
+    path = os.path.dirname(os.path.abspath(__file__))
+    write_data = True
+    plot = True
+    regress = True
+    if (write_data):
+        args = docopt(__doc__)
+        output_path = pathlib.Path('').absolute()
+        be_data = {'be_ar' : [], 'sim_times_ar' : []}
+        pickle.dump(be_data, open(path + '/be_data_mri.pick', 'wb'))
+
+        # Create output directory if needed
+        with Sync() as sync:
+            if sync.comm.rank == 0:
+                if not output_path.exists():
+                    output_path.mkdir()
+        post.visit_writes(args['<files>'], main, output=output_path)
+    if (plot):
[…]

### 2021-09-03 mri_nonlin/plot_ke.py

@@ -1,0 +1,108 @@
+"""
+Plot planes from joint analysis files.
+
+Usage:
+    plot_slices.py <files>... [--suffix=write_string_suffix] 
+
+"""
+
+import h5py
+import pickle
+import os
+import sys
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+plt.ioff()
+from dedalus.extras import plot_tools
+import publication_settings
+import scipy.stats
+
+matplotlib.rcParams.update(publication_settings.params)
+plt.rcParams.update({'figure.autolayout': True})
+plt.gcf().subplots_adjust(left=0.15)
+
+def main(filename, start, count, output):
+    """Save plot of specified tasks for given range of analysis writes."""
+
+    # Plot writes
+    with h5py.File(filename, mode='r') as file:
+        ke_data = pickle.load(open(path + '/ke_data_mri.pick', 'rb'))
+        ke_ar = ke_data['ke_ar']
+        # ke_max_ar = ke_data['ke_max_ar']
+        sim_times_ar = ke_data['sim_times_ar']
+        for index in range(start, start+count):
+            ke_avg = file['tasks']['ke'][index][0, 0, 0]
+            ke_ar.append(ke_avg)
+            sim_times_ar.append(file['scales/sim_time'][index])
+        ke_data['ke_ar'] = ke_ar
+        ke_data['sim_times_ar'] = sim_times_ar
+        pickle.dump(ke_data, open(path + '/ke_data_mri.pick', 'wb'))
+        
+
+
+if __name__ == "__main__":
+
+    import pathlib
+    from docopt import docopt
+    from dedalus.tools import logging
+    from dedalus.tools import post
+    from dedalus.tools.parallel import Sync
+
+    global path 
+    path = os.path.dirname(os.path.abspath(__file__))
+    write_data = True
+    plot = True
+    regress = True
+    if (write_data):
+        args = docopt(__doc__)
+        output_path = pathlib.Path('').absolute()
+        ke_data = {'ke_ar' : [], 'sim_times_ar' : []}
+        pickle.dump(ke_data, open(path + '/ke_data_mri.pick', 'wb'))
+
+        # Create output directory if needed
+        with Sync() as sync:
+            if sync.comm.rank == 0:
+                if not output_path.exists():
+                    output_path.mkdir()
[…]

### 2021-09-03 mri_nonlin/plot_slicepoints.py

@@ -28,7 +28,8 @@ def midplane(filename, start, count, output):
     """Save plot of specified tasks for given range of analysis writes."""
 
     # Plot settings
-    tasks = ['vx_midz', 'vy_midz', 'vz_midz', 'bx_midz', 'by_midz', 'bz_midz']
+    normal_dir = 'y'
+    tasks = ['vx_mid'+normal_dir, 'vy_mid'+normal_dir, 'vz_mid'+normal_dir, 'bx_mid'+normal_dir, 'by_mid'+normal_dir, 'bz_mid'+normal_dir]
     scale = 2.5
     dpi = 100
     title_func = lambda sim_time: 't = {:.3f}'.format(sim_time)
@@ -51,8 +52,8 @@ def midplane(filename, start, count, output):
                 axes = mfig.add_axes(i, j, [0, 0, 1, 1])
                 # Call plotting helper (dset axes: [t, x, y, z])
                 dset = file['tasks'][task]
-                image_axes = (3, 1)
-                data_slices = (index, slice(None), 0, slice(None))
+                image_axes = (3, 2)
+                data_slices = (index, 0, slice(None), slice(None))
                 # if (index % 5 != 0):
                 #     continue
                 plot_tools.plot_bot(dset, image_axes, data_slices, axes=axes, title=task, even_scale=True)
@@ -138,6 +139,7 @@ if __name__ == "__main__":
     args = docopt(__doc__)
 
     output_path = pathlib.Path(args['--output']).absolute()
+
     # Create output directory if needed
     with Sync() as sync:
         if sync.comm.rank == 0:

### 2021-10-10 legacy/plot_points.sh
vp working

@@ -13,7 +13,7 @@ export PYTHONNOUSERSITE=1
 conda activate dedalus
 # support lots of text output to stdio for analysis
 export MPI_UNBUFFERED_STDIO=true
-cd ~/scratch/dedalus/mri_simulations/
+cd ~/scratch/dedalus/mri/
 
 mpiexec_mpt -np 512 python3 -m dedalus merge_procs checkpoints_mri_non
 mpiexec_mpt -np 512 python3 -m dedalus merge_procs slicepoints_mri_non

### 2021-10-10 make_video.py

@@ -12,7 +12,7 @@ directory = path = os.path.dirname(os.path.abspath(__file__)) + "/frames_sliced/
 fileType = "*.png"
 videoName = 'frames_sliced_midz.avi'
 files = natsorted(glob.glob(directory + fileType))
-files = files[:150:]
+files = files[:10:]
 nFrames = len(files)
 
 index = 0

### 2021-10-10 mri_lin/mri_ivp_linear.sh

@@ -13,7 +13,7 @@ export PYTHONNOUSERSITE=1
 conda activate dedalus
 # support lots of text output to stdio for analysis
 export MPI_UNBUFFERED_STDIO=true
-cd ~/scratch/dedalus/mri_simulations/
+cd ~/scratch/dedalus/mri/
 
 mpiexec_mpt -np 128 python3 mri_ivp_linear.py
 mpiexec_mpt -np 128 python3 -m dedalus merge_procs checkpoints_mri_wed

### 2021-10-10 mri_nonlin/diff1en2_R1p01_lr/mri_nonlin_diff1en2_R1p01_lr.sh

@@ -1,10 +1,8 @@
 #PBS -S /bin/bash
-#PBS -l select=10:ncpus=28:mpiprocs=28:model=bro
-#PBS -l walltime=24:00:00
+#PBS -N mri_nonlin_diff1en2_R1p01
+#PBS -l select=1:ncpus=40:mpiprocs=40:model=sky_ele
+#PBS -l walltime=4:00:00
 #PBS -j oe
-file=${0##*/}
-job_name="${file%.*}"
-#PBS -N ${job_name}
 
 module load mpi-sgi/mpt
 module load comp-intel
@@ -20,16 +18,22 @@ export MPI_UNBUFFERED_STDIO=true
 source ~/png2mp4.sh
 cd ~/scratch/dedalus/mri/mri_nonlin
 
-SUFF="diff1en2_N64"
+SUFF="diff1en2_R1p01_lr"
 DIFF=0.01
 MPIPROC=32
 
+mkdir $SUFF
 mpiexec_mpt -np $MPIPROC python3 mri_nonlin.py $SUFF $DIFF
 mpiexec_mpt -np $MPIPROC python3 -m dedalus merge_procs scalars_${SUFF}
 mpiexec_mpt -np 1 python3 plot_ke.py scalars_${SUFF}/*.h5 --suffix=$SUFF
 mpiexec_mpt -np 1 python3 plot_be.py scalars_${SUFF}/*.h5 --suffix=$SUFF
 mpiexec_mpt -np $MPIPROC python3 -m dedalus merge_procs slicepoints_${SUFF}
-mpiexec_mpt -np $MPIPROC python3 plot_slicepoints.py slicepoints_${SUFF}/*.h5 --output=frames_${SUFF}
-png2mp4 frames_${SUFF}/ mri_${SUFF}.mp4 60
+mpiexec_mpt -np $MPIPROC python3 plot_slicepoints_xy.py slicepoints_${SUFF}/*.h5 --output=frames_xy_${SUFF}
+mpiexec_mpt -np $MPIPROC python3 plot_slicepoints_xz.py slicepoints_${SUFF}/*.h5 --output=frames_xz_${SUFF}
+mpiexec_mpt -np $MPIPROC python3 plot_slicepoints_yz.py slicepoints_${SUFF}/*.h5 --output=frames_yz_${SUFF}
+png2mp4 frames_xy_${SUFF}/ mri_${SUFF}_xy.mp4 60
[…]

### 2021-10-10 mri_nonlin/diff1en3_R1p01_N512/mri_nonlin_diff1en3_R1p01_N512.sh

@@ -1,10 +1,8 @@
 #PBS -S /bin/bash
-#PBS -l select=10:ncpus=28:mpiprocs=28:model=bro
-#PBS -l walltime=24:00:00
+#PBS -N mri_nonlin_diff1en3_R1p01
+#PBS -l select=13:ncpus=40:mpiprocs=40:model=sky_ele
+#PBS -l walltime=72:00:00
 #PBS -j oe
-file=${0##*/}
-job_name="${file%.*}"
-#PBS -N ${job_name}
 
 module load mpi-sgi/mpt
 module load comp-intel
@@ -20,16 +18,22 @@ export MPI_UNBUFFERED_STDIO=true
 source ~/png2mp4.sh
 cd ~/scratch/dedalus/mri/mri_nonlin
 
-SUFF="diff1en2_N64"
-DIFF=0.01
-MPIPROC=32
+SUFF="diff1en3_R1p01_N512"
+DIFF=0.001
+MPIPROC=512
 
 mpiexec_mpt -np $MPIPROC python3 mri_nonlin.py $SUFF $DIFF
 mpiexec_mpt -np $MPIPROC python3 -m dedalus merge_procs scalars_${SUFF}
 mpiexec_mpt -np 1 python3 plot_ke.py scalars_${SUFF}/*.h5 --suffix=$SUFF
 mpiexec_mpt -np 1 python3 plot_be.py scalars_${SUFF}/*.h5 --suffix=$SUFF
 mpiexec_mpt -np $MPIPROC python3 -m dedalus merge_procs slicepoints_${SUFF}
-mpiexec_mpt -np $MPIPROC python3 plot_slicepoints.py slicepoints_${SUFF}/*.h5 --output=frames_${SUFF}
-png2mp4 frames_${SUFF}/ mri_${SUFF}.mp4 60
+mpiexec_mpt -np $MPIPROC python3 plot_slicepoints_xy.py slicepoints_${SUFF}/*.h5 --output=frames_xy_${SUFF}
+mpiexec_mpt -np $MPIPROC python3 plot_slicepoints_xz.py slicepoints_${SUFF}/*.h5 --output=frames_xz_${SUFF}
+mpiexec_mpt -np $MPIPROC python3 plot_slicepoints_yz.py slicepoints_${SUFF}/*.h5 --output=frames_yz_${SUFF}
[…]

### 2021-10-10 mri_nonlin/mri_nonlin.sh

@@ -1,8 +1,10 @@
 #PBS -S /bin/bash
-#PBS -N mri_nonlin_diff1en2_L8
 #PBS -l select=10:ncpus=28:mpiprocs=28:model=bro
 #PBS -l walltime=24:00:00
 #PBS -j oe
+file=${0##*/}
+job_name="${file%.*}"
+#PBS -N ${job_name}
 
 module load mpi-sgi/mpt
 module load comp-intel

### 2021-10-10 mri_nonlin/mri_nonlin_diff1en2_R1p01.sh

@@ -1,8 +1,10 @@
 #PBS -S /bin/bash
-#PBS -N mri_nonlin_diff1en2_R1p01
-#PBS -l select=10:ncpus=28:mpiprocs=28:model=bro
-#PBS -l walltime=24:00:00
+#PBS -l select=1:ncpus=40:mpiprocs=40:model=sky_ele
+#PBS -l walltime=4:00:00
 #PBS -j oe
+file=${0##*/}
+job_name="${file%.*}"
+#PBS -N ${job_name}
 
 module load mpi-sgi/mpt
 module load comp-intel
@@ -20,14 +22,20 @@ cd ~/scratch/dedalus/mri/mri_nonlin
 
 SUFF="diff1en2_R1p01_test"
 DIFF=0.01
-MPIPROC=256
+MPIPROC=16
 
+mkdir $SUFF
 mpiexec_mpt -np $MPIPROC python3 mri_nonlin.py $SUFF $DIFF
-# mpiexec_mpt -np $MPIPROC python3 -m dedalus merge_procs scalars_${SUFF}
-# mpiexec_mpt -np 1 python3 plot_ke.py scalars_${SUFF}/*.h5 --suffix=$SUFF
-# mpiexec_mpt -np 1 python3 plot_be.py scalars_${SUFF}/*.h5 --suffix=$SUFF
-# mpiexec_mpt -np $MPIPROC python3 -m dedalus merge_procs slicepoints_${SUFF}
-# mpiexec_mpt -np $MPIPROC python3 plot_slicepoints_xy.py slicepoints_${SUFF}/*.h5 --output=frames_xy_${SUFF}
-# png2mp4 frames_xy_${SUFF}/ mri_${SUFF}_xy.mp4 60
+mpiexec_mpt -np $MPIPROC python3 -m dedalus merge_procs scalars_${SUFF}
+mpiexec_mpt -np 1 python3 plot_ke.py scalars_${SUFF}/*.h5 --suffix=$SUFF
+mpiexec_mpt -np 1 python3 plot_be.py scalars_${SUFF}/*.h5 --suffix=$SUFF
+mpiexec_mpt -np $MPIPROC python3 -m dedalus merge_procs slicepoints_${SUFF}
+mpiexec_mpt -np $MPIPROC python3 plot_slicepoints_xy.py slicepoints_${SUFF}/*.h5 --output=frames_xy_${SUFF}
[…]

### 2021-10-10 mri_nonlin/mri_nonlin_diff4en3_R1p01_N256.sh

@@ -1,6 +1,6 @@
 #PBS -S /bin/bash
 #PBS -l select=10:ncpus=28:mpiprocs=28:model=bro
-#PBS -l walltime=24:00:00
+#PBS -l walltime=72:00:00
 #PBS -j oe
 file=${0##*/}
 job_name="${file%.*}"
@@ -20,16 +20,22 @@ export MPI_UNBUFFERED_STDIO=true
 source ~/png2mp4.sh
 cd ~/scratch/dedalus/mri/mri_nonlin
 
-SUFF="diff1en2_N64"
-DIFF=0.01
-MPIPROC=32
+SUFF="diff2en3_R1p01_N256"
+DIFF=0.004
+MPIPROC=256
 
 mpiexec_mpt -np $MPIPROC python3 mri_nonlin.py $SUFF $DIFF
 mpiexec_mpt -np $MPIPROC python3 -m dedalus merge_procs scalars_${SUFF}
 mpiexec_mpt -np 1 python3 plot_ke.py scalars_${SUFF}/*.h5 --suffix=$SUFF
 mpiexec_mpt -np 1 python3 plot_be.py scalars_${SUFF}/*.h5 --suffix=$SUFF
 mpiexec_mpt -np $MPIPROC python3 -m dedalus merge_procs slicepoints_${SUFF}
-mpiexec_mpt -np $MPIPROC python3 plot_slicepoints.py slicepoints_${SUFF}/*.h5 --output=frames_${SUFF}
-png2mp4 frames_${SUFF}/ mri_${SUFF}.mp4 60
+mpiexec_mpt -np $MPIPROC python3 plot_slicepoints_xy.py slicepoints_${SUFF}/*.h5 --output=frames_xy_${SUFF}
+mpiexec_mpt -np $MPIPROC python3 plot_slicepoints_xz.py slicepoints_${SUFF}/*.h5 --output=frames_xz_${SUFF}
+mpiexec_mpt -np $MPIPROC python3 plot_slicepoints_yz.py slicepoints_${SUFF}/*.h5 --output=frames_yz_${SUFF}
+png2mp4 frames_xy_${SUFF}/ mri_${SUFF}_xy.mp4 60
+png2mp4 frames_xz_${SUFF}/ mri_${SUFF}_xz.mp4 60
+png2mp4 frames_yz_${SUFF}/ mri_${SUFF}_yz.mp4 60
[…]

### 2021-10-10 mri_nonlin/mri_vp.py

@@ -142,7 +142,7 @@ logger.info("running on processor mesh={}".format(mesh))
 domain = de.Domain([y_basis, z_basis, x_basis], grid_dtype=np.float64, mesh=mesh)
 
 # 3D MRI
-problem_variables = ['p','vx','vy','vz','bx','by','bz','ωy','ωz','jxx']
+problem_variables = ['p','vx','vy','vz','Ax','Ay','Az','Axx','Ayx','Azx', 'phi', 'ωy','ωz']
 problem = de.IVP(domain, variables=problem_variables, time='t')
 
 # Local parameters
@@ -161,9 +161,14 @@ problem.substitutions['v_dot_grad(A)'] = "vx * dx(A) + vy * dy(A) + vz * dz(A)"
 
 # non ideal
 problem.substitutions['ωx'] = "dy(vz) - dz(vy)"
+problem.substitutions['bx'] = "dy(Az) - dz(Ay)"
+problem.substitutions['by'] = "dz(Ax) - Azx"
+problem.substitutions['bz'] = "Ayx - dy(Ax)"
+
 problem.substitutions['jx'] = "dy(bz) - dz(by)"
 problem.substitutions['jy'] = "dz(bx) - dx(bz)"
 problem.substitutions['jz'] = "dx(by) - dy(bx)"
+
 problem.substitutions['L(A)'] = "dy(dy(A)) + dz(dz(A))"
 problem.substitutions['A_dot_grad_C(Ax, Ay, Az, C)'] = "Ax*dx(C) + Ay*dy(C) + Az*dz(C)"
 
@@ -179,18 +184,15 @@ problem.add_equation("ωy - dz(vx) + dx(vz) = 0")
 problem.add_equation("ωz - dx(vy) + dy(vx) = 0")
 
 # MHD equations: bx, by, bz, jxx
-problem.add_equation("dx(bx) + dy(by) + dz(bz) = 0", condition='(ny != 0) or (nz != 0)')
-problem.add_equation("Dt(bx) - B*dz(vx) + η*( dy(jz) - dz(jy) )            = b_dot_grad(vx) - v_dot_grad(bx)", condition='(ny != 0) or (nz != 0)')
+problem.add_equation("Axx + dy(Ay) + dz(Az) = 0")
 
-problem.add_equation("Dt(jx) - B*dz(ωx) + S*dz(bx) - η*( dx(jxx) + L(jx) ) = b_dot_grad(ωx) - v_dot_grad(jx)"
[…]

### 2021-10-10 mri_nonlin/plot_be_dirs.py

@@ -16,6 +16,7 @@ import matplotlib
 import matplotlib.pyplot as plt
 plt.ioff()
 # from dedalus.extras import plot_tools
+import dedalus.public as de
 import publication_settings
 
 matplotlib.rcParams.update(publication_settings.params)
@@ -27,17 +28,36 @@ def main(filename, start, count, output):
 
     # Plot writes
     with h5py.File(filename, mode='r') as file:
-        be_data = pickle.load(open(path + '/be_data_' + write_suffix + '.pick', 'rb'))
-        be_ar = be_data['be_ar']
-        # ke_max_ar = be_data['ke_max_ar']
+        be_data = pickle.load(open(path + '/be_dirs.pick', 'rb'))
+        be_ar1 = be_data['be_ar1']
+        be_ar2 = be_data['be_ar2']
+        be_ar3 = be_data['be_ar3']
         sim_times_ar = be_data['sim_times_ar']
+
         for index in range(start, start+count):
-            ke_avg = file['tasks']['be'][index][0, 0, 0]
-            be_ar.append(ke_avg)
+            field['g'] = file['tasks']['bx'][index]
+            be_totx = np.min(de.operators.integrate(field * field, 'x', 'y', 'z').evaluate()['g'])
+        
+            field['g'] = file['tasks']['by'][index]
+            be_toty = np.min(de.operators.integrate(field * field, 'x', 'y', 'z').evaluate()['g'])
+
+            if (just_perts):
+                field['g'] = file['tasks']['bz'][index]
+            else:
+                field['g'] = file['tasks']['bz'][index] + 1
+            be_totz = np.min(de.operators.integrate(field * field, 'x', 'y', 'z').evaluate()['g'])
+
+            be_ar1.append(be_totx)
+            be_ar2.append(be_toty)
+            be_ar3.append(be_totz)
+
             sim_times_ar.append(file['scales/sim_time'][index])
-        be_data['be_ar'] = be_ar
+        be_data['be_ar1'] = be_ar1
+        be_data['be_ar2'] = be_ar2
+        be_data['be_ar3'] = be_ar3
+            
         be_data['sim_times_ar'] = sim_times_ar
-        pickle.dump(be_data, open(path + '/be_data_' + write_suffix + '.pick', 'wb'))
[…]

### 2021-10-10 mri_nonlin/plot_job.sh

@@ -1,0 +1,28 @@
+#PBS -S /bin/bash
+#PBS -l select=1:ncpus=28:mpiprocs=28:model=bro
+#PBS -l walltime=1:00:00
+#PBS -j oe
+file=${0##*/}
+job_name="${file%.*}"
+#PBS -N ${job_name}
+
+module load mpi-sgi/mpt
+module load comp-intel
+export PATH=$HOME/scripts:$PATH
+deactivate
+unset PYTHONPATH
+source ~/miniconda3/etc/profile.d/conda.sh
+export PYTHONNOUSERSITE=1
+conda activate dedalus
+# support lots of text output to stdio for analysis
+export MPI_UNBUFFERED_STDIO=true
+
+cd ~/scratch/dedalus/mri/mri_nonlin
+
+SUFF="diff2en3_R1p01_N512"
+
+mpiexec_mpt -np 1 python3 plot_be.py scalars_${SUFF}/*.h5 --suffix=$SUFF
+mpiexec_mpt -np 1 python3 plot_ke.py scalars_${SUFF}/*.h5 --suffix=$SUFF
+
+# mpiexec_mpt -np 1 python3 plot_be.py ${SUFF}/scalars_${SUFF}/*.h5 --suffix=$SUFF
+# mpiexec_mpt -np 1 python3 plot_ke.py ${SUFF}/scalars_${SUFF}/*.h5 --suffix=$SUFF

### 2021-10-10 mri_nonlin/plot_job_ke.sh

@@ -1,6 +1,6 @@
 #PBS -S /bin/bash
-#PBS -l select=10:ncpus=28:mpiprocs=28:model=bro
-#PBS -l walltime=24:00:00
+#PBS -l select=1:ncpus=28:mpiprocs=28:model=bro
+#PBS -l walltime=8:00:00
 #PBS -j oe
 file=${0##*/}
 job_name="${file%.*}"
@@ -17,11 +17,8 @@ conda activate dedalus
 # support lots of text output to stdio for analysis
 export MPI_UNBUFFERED_STDIO=true
 
-source ~/png2mp4.sh
 cd ~/scratch/dedalus/mri/mri_nonlin
 
 SUFF="diff1en2_R1p01_lr"
 
-# mpiexec_mpt -np $MPIPROC python3 -m dedalus merge_procs scalars_${SUFF}
-mpiexec_mpt -np 1 python3 plot_ke.py ${SUFF}/scalars_${SUFF}/*.h5 --suffix=$SUFF
-mpiexec_mpt -np 1 python3 plot_be.py ${SUFF}/scalars_${SUFF}/*.h5 --suffix=$SUFF
+mpiexec_mpt -np 1 python3 plot_ke_dirs.py ${SUFF}/checkpoints_${SUFF}/*.h5 --suffix=$SUFF

### 2021-10-10 mri_nonlin/plot_ke.py

@@ -14,7 +14,7 @@ import numpy as np
 import matplotlib
 import matplotlib.pyplot as plt
 plt.ioff()
-from dedalus.extras import plot_tools
+# from dedalus.extras import plot_tools
 import publication_settings
 import scipy.stats
 
@@ -27,7 +27,7 @@ def main(filename, start, count, output):
 
     # Plot writes
     with h5py.File(filename, mode='r') as file:
-        ke_data = pickle.load(open(path + '/ke_data_mri.pick', 'rb'))
+        ke_data = pickle.load(open(path + '/ke_data_' + write_suffix + '.pick', 'rb'))
         ke_ar = ke_data['ke_ar']
         # ke_max_ar = ke_data['ke_max_ar']
         sim_times_ar = ke_data['sim_times_ar']
@@ -37,7 +37,7 @@ def main(filename, start, count, output):
             sim_times_ar.append(file['scales/sim_time'][index])
         ke_data['ke_ar'] = ke_ar
         ke_data['sim_times_ar'] = sim_times_ar
-        pickle.dump(ke_data, open(path + '/ke_data_mri.pick', 'wb'))
+        pickle.dump(ke_data, open(path + '/ke_data_' + write_suffix + '.pick', 'wb'))
         
 
 
@@ -50,15 +50,17 @@ if __name__ == "__main__":
     from dedalus.tools.parallel import Sync
 
     global path 
+    global write_suffix
+    args = docopt(__doc__)
+    write_suffix = args['--suffix']
     path = os.path.dirname(os.path.abspath(__file__))
     write_data = True
     plot = True
     regress = True
     if (write_data):
-        args = docopt(__doc__)
         output_path = pathlib.Path('').absolute()
         ke_data = {'ke_ar' : [], 'sim_times_ar' : []}
-        pickle.dump(ke_data, open(path + '/ke_data_mri.pick', 'wb'))
+        pickle.dump(ke_data, open(path + '/ke_data_' + write_suffix + '.pick', 'wb'))
 
         # Create output directory if needed
         with Sync() as sync:
@@ -68,22 +70,24 @@ if __name__ == "__main__":
         post.visit_writes(args['<files>'], main, output=output_path)
     if (plot):
         fig = plt.figure()
-        write_suffix = args['--suffix']
[…]

### 2021-10-10 mri_nonlin/test.sh

@@ -1,6 +1,7 @@
+#!/bin/bash
+#PBS -l select=1:ncpus=28:mpiprocs=28:model=bro
 #PBS -S /bin/bash
-#PBS -l select=10:ncpus=28:mpiprocs=28:model=bro
-#PBS -l walltime=24:00:00
+#PBS -l walltime=8:00:00
 #PBS -j oe
 file=${0##*/}
 job_name="${file%.*}"
@@ -17,11 +18,4 @@ conda activate dedalus
 # support lots of text output to stdio for analysis
 export MPI_UNBUFFERED_STDIO=true
 
-source ~/png2mp4.sh
 cd ~/scratch/dedalus/mri/mri_nonlin
-
-SUFF="diff1en2_R1p01_lr"
-
-# mpiexec_mpt -np $MPIPROC python3 -m dedalus merge_procs scalars_${SUFF}
-mpiexec_mpt -np 1 python3 plot_ke.py ${SUFF}/scalars_${SUFF}/*.h5 --suffix=$SUFF
-mpiexec_mpt -np 1 python3 plot_be.py ${SUFF}/scalars_${SUFF}/*.h5 --suffix=$SUFF

### 2021-10-10 mri_nonlin/mri_vp.py
vp working

@@ -1,6 +1,6 @@
 """
 Modified from: The magnetorotational instability prefers three dimensions.
-3D MHD initial value problem (simulation)
+3D MHD initial value problem
 """
 
 from docopt import docopt

### 2022-02-04 adjop/vp_bvp.py
vector potential bvp, d3 dir, adjoint optim

@@ -1,0 +1,70 @@
+
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+
+from dedalus import public as de
+from dedalus.extras import plot_tools
+
+
+# Create bases and domain
+y_basis = de.Fourier('y', 32, interval=(0, 8*np.pi))
+z_basis = de.Fourier('z', 32, interval=(0, 8*np.pi))
+x_basis = de.Chebyshev('x', 8, interval=(-np.pi/2.0, np.pi/2.0))
+domain = de.Domain([y_basis, z_basis, x_basis], grid_dtype=np.float64)
+
+y = domain.grid(0)
+z = domain.grid(1)
+x = domain.grid(2)
+
+by = domain.new_field()
+bz = domain.new_field()
+bx = domain.new_field()
+
+by['g'] = 0.
+bz['g'] = np.sin(y) * np.cos(x)
+bx['g'] = 0.
+
+problem = de.LBVP(domain, variables=['Ax','Ay', 'Az', 'phi'])
+
+problem.parameters['by'] = by
+problem.parameters['bz'] = bz
+problem.parameters['bx'] = bx
+
+problem.add_equation("phi = 0", condition="(ny==0) and (nz==0)")
+problem.add_equation("Ax = 0", condition="(ny==0) and (nz==0)")
+problem.add_equation("Ay = 0", condition="(ny==0) and (nz==0)")
+problem.add_equation("Az = 0", condition="(ny==0) and (nz==0)")
+
+problem.add_equation("dx(Ax) + dy(Ay) + dz(Az) = 0", condition="(ny!=0) or (nz!=0)")
+problem.add_equation("dy(Az) - dz(Ay) + dx(phi) = bx", condition="(ny!=0) or (nz!=0)")
+problem.add_equation("dz(Ax) - dx(Az) + dy(phi) = by", condition="(ny!=0) or (nz!=0)")
+problem.add_equation("dx(Ay) - dy(Ax) + dz(phi) = bz", condition="(ny!=0) or (nz!=0)")
+
+problem.add_bc("left(Ay) = 0", condition="(ny!=0) or (nz!=0)")
+problem.add_bc("left(Az) = 0", condition="(ny!=0) or (nz!=0)")
+problem.add_bc("right(Ay) = 0", condition="(ny!=0) or (nz!=0)")
+problem.add_bc("right(Az) = 0", condition="(ny!=0) or (nz!=0)")
+
+# Build solver
[…]

### 2022-02-04 mri_d3/mri.py

@@ -1,0 +1,183 @@
+"""
+Dedalus script simulating 2D horizontally-periodic Rayleigh-Benard convection.
+This script demonstrates solving a 2D cartesian initial value problem. It can
+be ran serially or in parallel, and uses the built-in analysis framework to save
+data snapshots to HDF5 files. The `plot_snapshots.py` script can be used to
+produce plots from the saved data. The simulation should take roughly 10
+cpu-minutes to run.
+The problem is non-dimensionalized using the box height and freefall time, so
+the resulting thermal diffusivity and viscosity are related to the Prandtl
+and Rayleigh numbers as:
+    kappa = (Rayleigh * Prandtl)**(-1/2)
+    nu = (Rayleigh / Prandtl)**(-1/2)
+For incompressible hydro with two boundaries, we need two tau terms for each the
+velocity and buoyancy. Here we choose to use a first-order formulation, putting
+one tau term each on auxiliary first-order gradient variables and the others in
+the PDE, and lifting them all to the first derivative basis. This formulation puts
+a tau term in the divergence constraint, as required for this geometry.
+To run and plot using e.g. 4 processes:
+    $ mpiexec -n 4 python3 rayleigh_benard.py
+    $ mpiexec -n 4 python3 plot_snapshots.py snapshots/*.h5
+"""
+
+import numpy as np
+import dedalus.public as d3
+import logging
+from mpi4py import MPI
+logger = logging.getLogger(__name__)
+
+# TODO: maybe fix plotting to directly handle vectors
+# TODO: get unit vectors from coords?
+# TODO: cleanup integ shortcuts
+
+
+# Parameters
+Lx = np.pi
+ar = 8
+Ly = Lz = ar * Lx
+Nx, Ny, Nz = 32, 128, 128
+Rayleigh = 1e6
+Prandtl = 1
+dealias = 3/2
+stop_sim_time = 25
+timestepper = d3.RK222
+max_timestep = 0.01
+dtype = np.float64
+
+
+ncpu = MPI.COMM_WORLD.size
+log2 = np.log2(ncpu)
+if log2 == int(log2):
[…]

### 2022-02-04 mri_nonlin/viff1en2_R0p6_Bsin0p5y_N128_U04e0_tau1e0/mri_vp.py

@@ -291,11 +291,11 @@ if not pathlib.Path(restart_state_dir).exists():
     slices = domain.dist.grid_layout.slices(scales=1)
 
     # Linear background + perturbations damped at walls
-    vx['g'] += noise / 1e1
+    vx['g'] += noise * U0
     # bz['g'] = 1e2*(np.sin((x))*np.cos(y) - 2.0/np.pi)
     # Ay['g'] += -(np.cos(2*x) + 1) / 2.0
     # Ay.differentiate('x', out = Ayx)
-    Ax['g'] += U0 * np.cos(x) * (np.cos(z) + 1) * 2.0
+    Ax['g'] += U0 * np.cos(x) * (np.cos(0.5*y) + 1) * 2.0
     # Ax.differentiate('y', out = Axy)
     
     filter_field(vx)
@@ -350,6 +350,10 @@ slicepoints.add_task("interp(bz, y={})".format(ar * Lx / 2.0), name="bz_midy")
 slicepoints.add_task("interp(bz, z={})".format(ar * Lx / 2.0), name="bz_midz")
 slicepoints.add_task("interp(bz, x={})".format(0.0), name="bz_midx")
 
+slicepoints.add_task("integ(integ(integ(vx*vx + vy*vy + vz*vz, 'x'), 'y'), 'z') / (Lx**3 * ar**2)", name="ke")
+slicepoints.add_task("integ(integ(integ(bx*bx + by*by + bz*bz, 'x'), 'y'), 'z') / (Lx**3 * ar**2)", name="be")
+
+slicepoints.add_task("integ(integ(integ(sqrt(vx*vx + vy*vy + vz*vz), 'x'), 'y'), 'z') / (Lx**3 * ar**2)", name="Re")
 
 slicepoints.add_task("integ(integ(vx, 'y'), 'z') / (Lx**2 * ar**2)", name="vx_profs")
 slicepoints.add_task("integ(integ(bx, 'y'), 'z') / (Lx**2 * ar**2)", name="bx_profs")
@@ -366,19 +370,19 @@ scalars = solver.evaluator.add_file_handler('scalars_' + run_suffix, sim_dt=0.1,
 scalars.add_task("integ(integ(integ(vx*vx + vy*vy + vz*vz, 'x'), 'y'), 'z')", name="ke")
[…]

### 2022-02-04 mri_nonlin/viff1en2_R0p6_Bsin0p5y_N128_U04e0_tau1e0/plot_ke.py

@@ -1,0 +1,120 @@
+"""
+Plot planes from joint analysis files.
+
+Usage:
+    plot_slices.py <files>... [--suffix=write_string_suffix] 
+
+"""
+
+import h5py
+import pickle
+import os
+import sys
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+plt.ioff()
+# from dedalus.extras import plot_tools
+import publication_settings
+import scipy.stats
+
+matplotlib.rcParams.update(publication_settings.params)
+plt.rcParams.update({'figure.autolayout': True})
+plt.gcf().subplots_adjust(left=0.15)
+
+def main(filename, start, count, output):
+    """Save plot of specified tasks for given range of analysis writes."""
+
+    # Plot writes
+    with h5py.File(filename, mode='r') as file:
+        ke_data = pickle.load(open(path + '/ke_data_' + write_suffix + '.pick', 'rb'))
+        ke_ar = ke_data['ke_ar']
+        ke_tot_ar = ke_data['ke_tot_ar']
+        # ke_max_ar = ke_data['ke_max_ar']
+        sim_times_ar = ke_data['sim_times_ar']
+        for index in range(start, start+count):
+            ke_avg = file['tasks']['ke'][index][0, 0, 0]
+            ke_tot_avg = file['tasks']['ke_tot'][index][0, 0, 0]
+            ke_ar.append(ke_avg)
+            ke_tot_ar.append(ke_tot_avg)
+            sim_times_ar.append(file['scales/sim_time'][index])
+        ke_data['ke_ar'] = ke_ar
+        ke_data['ke_tot_ar'] = ke_tot_ar
+        ke_data['sim_times_ar'] = sim_times_ar
+        pickle.dump(ke_data, open(path + '/ke_data_' + write_suffix + '.pick', 'wb'))
+        
+
+
+if __name__ == "__main__":
+
+    import pathlib
+    from docopt import docopt
+    from dedalus.tools import logging
+    from dedalus.tools import post
+    from dedalus.tools.parallel import Sync
+
+    global path 
+    global write_suffix
+    args = docopt(__doc__)
+    write_suffix = args['--suffix']
+    path = os.path.dirname(os.path.abspath(__file__))
+    write_data = True
+    plot = True
+    regress = True
+    if (write_data):
[…]

### 2022-02-04 mri_nonlin/viff1en2_R0p6_Bysinz_N128_U04e0_noslip1/mri_vp.py

@@ -291,7 +291,7 @@ if not pathlib.Path(restart_state_dir).exists():
     slices = domain.dist.grid_layout.slices(scales=1)
 
     # Linear background + perturbations damped at walls
-    vx['g'] += noise / 1e1
+    vx['g'] += noise * U0
     # bz['g'] = 1e2*(np.sin((x))*np.cos(y) - 2.0/np.pi)
     # Ay['g'] += -(np.cos(2*x) + 1) / 2.0
     # Ay.differentiate('x', out = Ayx)
@@ -350,6 +350,10 @@ slicepoints.add_task("interp(bz, y={})".format(ar * Lx / 2.0), name="bz_midy")
 slicepoints.add_task("interp(bz, z={})".format(ar * Lx / 2.0), name="bz_midz")
 slicepoints.add_task("interp(bz, x={})".format(0.0), name="bz_midx")
 
+slicepoints.add_task("integ(integ(integ(vx*vx + vy*vy + vz*vz, 'x'), 'y'), 'z') / (Lx**3 * ar**2)", name="ke")
+slicepoints.add_task("integ(integ(integ(bx*bx + by*by + bz*bz, 'x'), 'y'), 'z') / (Lx**3 * ar**2)", name="be")
+
+slicepoints.add_task("integ(integ(integ(sqrt(vx*vx + vy*vy + vz*vz), 'x'), 'y'), 'z') / (Lx**3 * ar**2)", name="Re")
 
 slicepoints.add_task("integ(integ(vx, 'y'), 'z') / (Lx**2 * ar**2)", name="vx_profs")
 slicepoints.add_task("integ(integ(bx, 'y'), 'z') / (Lx**2 * ar**2)", name="bx_profs")
@@ -366,19 +370,19 @@ scalars = solver.evaluator.add_file_handler('scalars_' + run_suffix, sim_dt=0.1,
 scalars.add_task("integ(integ(integ(vx*vx + vy*vy + vz*vz, 'x'), 'y'), 'z')", name="ke")
 scalars.add_task("integ(integ(integ(bx*bx + by*by + bz*bz, 'x'), 'y'), 'z')", name="be")
 
-scalars.add_task("integ(integ(integ(vx*vx, 'x'), 'y'), 'z') / (Lx**3 * ar**2)", name="ke_x")
[…]

### 2022-02-04 mri_nonlin/viff1en2_R0p6_Bysinz_N128_U04e0_noslip1/mvp_viff1en2.sh

@@ -1,6 +1,6 @@
 #PBS -S /bin/bash
 #PBS -l select=10:ncpus=28:mpiprocs=28:model=bro
-#PBS -l walltime=2:00:00
+#PBS -l walltime=4:00:00
 #PBS -j oe
 #PBS -W group_list=s2276
 file=${0##*/}
@@ -21,7 +21,7 @@ export MPI_UNBUFFERED_STDIO=true
 source ~/png2mp4.sh
 cd ~/scratch/dedalus/mri/mri_nonlin
 
-SUFF="viff1en2_R0p6_Bysinz_N128_U01e0_tau1e0"
+SUFF="viff1en2_R0p6_Bysinz_N128_U04e0_noslip1"
 MPIPROC=256
 
 mkdir $SUFF

### 2022-06-16 adjop/kdv/kdv_twomodes.py
kdv obj heatmaps

@@ -1,0 +1,154 @@
+from distutils.command.bdist import show_formats
+import os
+path = os.path.dirname(os.path.abspath(__file__))
+from ast import For
+from contextlib import nullcontext
+from turtle import backward
+import numpy as np
+import pickle
+import sys
+sys.path.append(path + "/..")
+import h5py
+import gc
+import dedalus.public as d3
+from dedalus.core.domain import Domain
+from mpi4py import MPI
+CW = MPI.COMM_WORLD
+import logging
+import pathlib
+logger = logging.getLogger(__name__)
+import ForwardKDV
+import matplotlib.pyplot as plt
+from docopt import docopt
+from pathlib import Path
+from configparser import ConfigParser
+from scipy import optimize
+from datetime import datetime
+
+filename = path + '/kdv_options.cfg'
+config = ConfigParser()
+config.read(str(filename))
+
+logger.info('Running diffusion_burgers.py with the following parameters:')
+logger.info(config.items('parameters'))
+
+# Parameters
+Lx = config.getfloat('parameters', 'Lx')
+N = config.getint('parameters', 'Nx')
+
+a = config.getfloat('parameters', 'a')
+b = config.getfloat('parameters', 'b')
+T = config.getfloat('parameters', 'T')
+dt = config.getfloat('parameters', 'dt')
+num_cp = config.getint('parameters', 'num_cp')
+
+# Simulation Parameters
+dealias = 3/2
+dtype = np.float64
+
+periodic = config.getboolean('parameters', 'periodic')
+epsilon_safety = default_gamma = 0.6
+
+# Bases
+xcoord = d3.Coordinate('x')
+dist = d3.Distributor(xcoord, dtype=np.float64)
+
+if periodic:
+    xbasis = d3.RealFourier(xcoord, size=N, bounds=(0, Lx), dealias=3/2)
+else:
+    xbasis = d3.ChebyshevT(xcoord, size=N, bounds=(0, Lx), dealias=3/2)
+
+domain = Domain(dist, [xbasis])
+dist = domain.dist
+
+x = dist.local_grid(xbasis)
+forward_problem = ForwardKDV.build_problem(domain, xcoord, a, b)
+forward_solver = forward_problem.build_solver(d3.RK443)
+both = True
+write_objectives = False
[…]

### 2022-06-23 adjop/shear_cheb/main_toy.sh
toy shear and d3 bugs

@@ -35,8 +35,8 @@ SUFFIX="test"
 MPIPROC=4
 # OLDSUFFIX=$SUFFIX
 
-# if [ false ] || [ ! -d "$SUFFIX" ]; then
-if [ ! -d "$SUFFIX" ]; then
+if [ true ]; then
+# if [ ! -d "$SUFFIX" ]; then
 
     mkdir $SUFFIX
     mkdir $SUFFIX/checkpoints

### 2022-06-23 adjop/shear_cheb/shear_cg.py

@@ -34,14 +34,17 @@ from OptimizationContext import OptimizationContext
 from ShearOptimization import ShearOptimization
 import ForwardShear
 import BackwardShear
-import PressureBVP
 import matplotlib.pyplot as plt
 from scipy.optimize import minimize, basinhopping, OptimizeResult
 from natsort import natsorted
 
-args = docopt(__doc__)
-filename = Path(args['<config_file>'])
-write_suffix = args['<run_suffix>']
+try:
+    args = docopt(__doc__)
+    filename = Path(args['<config_file>'])
+    write_suffix = args['<run_suffix>']
+except:
+    filename = path + '/shear_options_toy.cfg'
+    write_suffix = 'test'
 
 config = ConfigParser()
 config.read(str(filename))
@@ -217,8 +220,8 @@ opt.objectiveT += psi_weight   * (psi - Psi)**2
 
 opt.backward_ic['s_t']  =  2*s_weight*(s - S)
 opt.backward_ic['u_t'] += -2.0*omega_weight*d3.skew(d3.grad((w - W)))
-opt.backward_ic['u_t'] +=  2.0*psi_weight* ( (psi - Psi).evaluate().antidifferentiate(zbasis, ('left', 0)) )
-# opt.backward_ic['u_t'] +=  2.0*psi_weight* ( (psi - Psi) * (Z*ex - X*ez) - integ((psi - Psi) * (Z*ex - X*ez)) )
+# opt.backward_ic['u_t'] +=  2.0*psi_weight* ( (psi - Psi).evaluate().antidifferentiate(zbasis, ('left', 0)) )
+opt.backward_ic['u_t'] +=  2.0*psi_weight* ( psi - Psi )
 opt.backward_ic['u_t'] +=  2.0*cc_weight*curl(lap(w - W))
 
 # opt.backward_ic['u_t'] += 2.0*psi_weight*d3.skew(d3.grad((psi - Psi)))

### 2022-06-23 d3_bugs/antidifferentiate_bug.py

@@ -1,0 +1,8 @@
+import numpy as np
+import dedalus.public as d3
+from dedalus.core.domain import Domain
+
+xcoord = d3.Coordinate('x')
+dist = d3.Distributor(xcoord, dtype=np.float64)
+
+xbasis = d3.RealFourier(xcoord, size=8, bounds=(0, 2*np.pi), dealias=3/2)

### 2022-06-23 d3_bugs/test_antidifferentiate_fourier.py

@@ -1,0 +1,37 @@
+"""Test antidifferentiate method with fourier and chebyshev bases."""
+
+import pytest
+import numpy as np
+from dedalus.core import coords, distributor, basis, field, operators, problems, solvers
+
+
+N_range = [8, 16, 32]
+L_range = [1, 2, 4]
+k_range = [0, 1]
+dtype_range = [np.float64, np.complex128]
+
+
+@pytest.mark.parametrize('N', N_range)
+@pytest.mark.parametrize('a', ab_range)
+@pytest.mark.parametrize('b', ab_range)
+@pytest.mark.parametrize('k', k_range)
+@pytest.mark.parametrize('dtype', dtype_range)
+@pytest.mark.parametrize('layout', ['g', 'c'])
+def test_jacobi_convert_constant(N, a, b, k, dtype, layout):
+    c = coords.Coordinate('x')
+    d = distributor.Distributor((c,))
+    b = basis.Jacobi(c, size=N, a0=a, b0=b, a=a+k, b=b+k, bounds=(0, 1))
+    fc = field.Field(dist=d, dtype=dtype)
+    fc['g'] = 1
+    fc[layout]
+    f = operators.convert(fc, (b,)).evaluate()
+    assert np.allclose(fc['g'], f['g'])
+
+
+
+xcoord = d3.Coordinate('x')
+dist = d3.Distributor(xcoord, dtype=np.float64)
+
+xbasis = d3.RealFourier(xcoord, size=8, bounds=(0, 2*np.pi), dealias=3/2)
+f = dist.Field(name='f', bases=(xbasis))
+f['g'] = 

### 2022-06-23 adjop/kdv/phasespace/combine_images.py
kdv phasespace parallel

@@ -1,0 +1,41 @@
+import sys
+from PIL import Image
+
+abvec = []
+abvec.append((0.2, 0.05))
+abvec.append((0.2, 1.0))
+abvec.append((0.2, 3.0))
+abvec.append((0.05, 0.05))
+abvec.append((0.05, 1.0))
+abvec.append((0.05, 3.0))
+
+paths = []
+# paths.append('objtest_a0p0b1p0.png')
+# paths.append('objtest_a0p0b8p0.png')
+# paths.append('objtest_a0p25b1p0.png')
+# paths.append('objtest_a0p5b1p0.png')
+
+for (a, b) in abvec:
+  a_str = str(a).replace('.', 'p')
+  b_str = str(b).replace('.', 'p')
+  paths.append('objtest_a{}b{}.png'.format(a_str, b_str))
+
+images = [Image.open(x) for x in paths]
+widths, heights = zip(*(i.size for i in images))
+
+total_width = round(sum(widths) / 2.0)
+max_height = round(sum(heights) / 3.0)
+
+new_im = Image.new('RGB', (total_width, max_height))
+
+x_offset = 0
+for im in images[:3]:
+  new_im.paste(im, (x_offset,0))
+  x_offset += im.size[0]
+
+x_offset = 0
+for im in images[3:]:
+  new_im.paste(im, (x_offset,im.size[1]))
+  x_offset += im.size[0]
+
+new_im.save('kdv_objectives_T5.png')

### 2022-06-23 adjop/kdv/phasespace/comm_min_ex.py

@@ -1,0 +1,38 @@
+import os
+path = os.path.dirname(os.path.abspath(__file__))
+import numpy as np
+import dedalus.public as d3
+from mpi4py import MPI
+CW = MPI.COMM_WORLD
+import logging
+logger = logging.getLogger(__name__)
+
+# Simulation Parameters
+dealias = 3/2
+dtype = np.float64
+
+# Bases
+xcoord = d3.Coordinate('x')
+dist = d3.Distributor(xcoord, dtype=np.float64, comm=MPI.COMM_SELF)
+xbasis = d3.RealFourier(xcoord, size=16, bounds=(0, 1), dealias=3/2)
+x = dist.local_grid(xbasis)
+
+u = dist.Field(name='u', bases=xbasis)
+dx = lambda A: d3.Differentiate(A, xcoord)
+
+# Problem
+problem = d3.IVP([u], namespace=locals())
+problem.add_equation("dt(u) = 0")
+
+CW.Barrier()
+forward_solver = problem.build_solver(d3.RK443)
+CW.Barrier()
+
+forward_solver.state[0]['g'] = mode1 = np.sin(2*np.pi * x)
+for i in range(100):
+    forward_solver.step(0.001)
+
+end_state = forward_solver.state[0]['g'].copy()
+soln = np.sum(end_state * end_state)
+CW.Barrier()
+print('soln L2 norm = {}; rank = {}'.format(soln, CW.rank))

### 2022-06-23 adjop/kdv/phasespace/kdv_twomodes.py

@@ -36,8 +36,12 @@ logger.info(config.items('parameters'))
 Lx = config.getfloat('parameters', 'Lx')
 N = config.getint('parameters', 'Nx')
 
-a = config.getfloat('parameters', 'a')
-b = config.getfloat('parameters', 'b')
+if (len(sys.argv) == 3):
+    a = float(sys.argv[1])
+    b = float(sys.argv[2])
+else:
+    a = config.getfloat('parameters', 'a')
+    b = config.getfloat('parameters', 'b')
 T = config.getfloat('parameters', 'T')
 dt = config.getfloat('parameters', 'dt')
 num_cp = config.getint('parameters', 'num_cp')
@@ -62,13 +66,17 @@ domain = Domain(dist, [xbasis])
 dist = domain.dist
 
 x = dist.local_grid(xbasis)
+CW.Barrier()
 forward_problem = ForwardKDV.build_problem(domain, xcoord, a, b)
+CW.Barrier()
 forward_solver = forward_problem.build_solver(d3.RK443)
+CW.Barrier()
+
 both = True
 write_objectives = False
 k1 = 1
-k2 = 6
-Nmodes = 20
+k2 = 2
+Nmodes = 100
 
 mode1 = np.sin(k1 * 2*np.pi * x / Lx)
 mode2 = np.sin(k2 * 2*np.pi * x / Lx)
@@ -85,27 +93,47 @@ def compute_coeffs(u, mode1_f, mode2_f):
     c2 = 2.0 / Lx * d3.Integrate(u*mode2_f).evaluate()['g'].flat[0]
     return (c1, c2)
 
-def evolve(ic, forward_solver, T, dt):
-    forward_solver.sim_time = 0.0
-    forward_solver.stop_sim_time = T
-    forward_solver.state[0].change_scales(1)
-    forward_solver.state[0]['g'] = ic.copy()
-    while forward_solver.proceed:
-        forward_solver.step(dt)
-    forward_solver.state[0].change_scales(1)
-    return forward_solver.state[0]
+def evolve(ic, fsolve, T, dt):
+    fsolve.sim_time = 0.0
+    fsolve.stop_sim_time = T
+    fsolve.state[0].change_scales(1)
+    fsolve.state[0]['g'] = ic.copy()
 
+    while fsolve.proceed:
+        fsolve.step(dt)
+    fsolve.state[0].change_scales(1)
+    # CW.barrier()
[…]

### 2022-06-23 adjop/kdv/phasespace/main.sh

@@ -1,0 +1,12 @@
+#!/bin/bash
+
+source ~/miniconda3/etc/profile.d/conda.sh
+conda activate dedalus3
+
+MPIPROCS=1000
+mpiexec_mpt -np $MPIPROCS python3 kdv_twomodes.py 0.2 0.05
+mpiexec_mpt -np $MPIPROCS python3 kdv_twomodes.py 0.2 1.0
+mpiexec_mpt -np $MPIPROCS python3 kdv_twomodes.py 0.2 3.0
+mpiexec_mpt -np $MPIPROCS python3 kdv_twomodes.py 0.05 0.05
+mpiexec_mpt -np $MPIPROCS python3 kdv_twomodes.py 0.05 1.0
+mpiexec_mpt -np $MPIPROCS python3 kdv_twomodes.py 0.05 3.0

### 2022-06-23 adjop/shear_cheby/BackwardShear.py

@@ -1,47 +1,0 @@
-import numpy as np
-import dedalus.public as d3
-from mpi4py import MPI
-CW = MPI.COMM_WORLD
-import logging
-logger = logging.getLogger(__name__)
-import sys
-import matplotlib.pyplot as plt
-
-def build_problem(domain, coords, Reynolds):
-
-    # unpack domain
-    dist = domain.dist
-    bases = domain.bases
-    xbasis, zbasis = bases[0], bases[1]
-
-    # Fields
-    p_t = dist.Field(name='p_t', bases=bases)
-    u_t = dist.VectorField(coords, name='u_t', bases=bases)
-    u = dist.VectorField(coords, name='u', bases=bases)
-    tau_p_t = dist.Field(name='tau_p_t')
-    tau1u_t = dist.VectorField(coords, name='tau1u_t', bases=(xbasis))
-    tau2u_t = dist.VectorField(coords, name='tau2u_t', bases=(xbasis))
-
-    # Substitutions
-    nu = 1 / Reynolds
-    S = 1
-    x, z = dist.local_grids(xbasis, zbasis)
-    ex, ez = coords.unit_vector_fields(dist)
-    integ = lambda A: d3.Integrate(d3.Integrate(A, 'z'), 'x')
-
-    # nccs
-    U0 = dist.VectorField(coords, name='U0', bases=zbasis)
-    U0['g'][0] = S * z
-
-    lift_basis = zbasis.clone_with(a=1/2, b=1/2) # First derivative basis
-    lift = lambda A, n: d3.Lift(A, lift_basis, n)
-    grad_u_t = d3.grad(u_t) + ez*lift(tau1u_t,-1) # First-order reduction
-
-    # Problem
-    problem = d3.IVP([u_t, p_t, tau_p_t, tau1u_t, tau2u_t], namespace=locals())
-    problem.add_equation("dt(u_t) + grad(p_t) + nu*div(grad_u_t) + lift(tau2u_t, -1) = u_t @ transpose(grad(u + U0)) - (u + U0) @ grad(u_t)")
-    problem.add_equation("trace(grad_u_t) + tau_p_t = 0")
-    problem.add_equation("integ(p_t) = 0") # Pressure gauge
-    problem.add_equation("u_t(z='left') = 0")
-    problem.add_equation("u_t(z='right') = 0")
-    return problem

### 2022-06-23 adjop/shear_cheby/ForwardShear.py

@@ -1,59 +1,0 @@
-import numpy as np
-import dedalus.public as d3
-from mpi4py import MPI
-CW = MPI.COMM_WORLD
-import logging
-logger = logging.getLogger(__name__)
-import sys
-import matplotlib.pyplot as plt
-
-def build_problem(domain, coords, Reynolds):
-
-    # unpack domain
-    dist = domain.dist
-    bases = domain.bases
-    xbasis, zbasis = bases[0], bases[1]
-
-    # Fields
-    p = dist.Field(name='p', bases=(xbasis, zbasis))
-    u = dist.VectorField(coords, name='u', bases=(xbasis, zbasis))
-    s = dist.Field(name='s', bases=(xbasis, zbasis))
-
-    tau_p = dist.Field(name='tau_p')
-    tau1u = dist.VectorField(coords, name='tau1u', bases=(xbasis))
-    tau2u = dist.VectorField(coords, name='tau2u', bases=(xbasis))
-    tau1s = dist.Field(name='tau1s', bases=(xbasis))
-    tau2s = dist.Field(name='tau2s', bases=(xbasis))
-
-    # Substitutions
-    nu = 1 / Reynolds
-    S = 1
-    x, z = dist.local_grids(xbasis, zbasis)
-    ex, ez = coords.unit_vector_fields(dist)
-    integ = lambda A: d3.Integrate(d3.Integrate(A, 'z'), 'x')
-
-    # nccs
-    U0 = dist.VectorField(coords, name='U0', bases=zbasis)
-    U0['g'][0] = S * z
-
-    lift_basis = zbasis.derivative_basis(1) # First derivative basis
-    lift = lambda A, n: d3.Lift(A, lift_basis, n)
-    grad_u = d3.grad(u) + ez*lift(tau1u,-1) # First-order reduction
-    grad_s = d3.grad(s) + ez*lift(tau1s,-1) # First-order reduction
-    u_tot = u + U0
-    # Problem
-    problem = d3.IVP([u, s, p, tau_p, tau1u, tau2u, tau1s, tau2s], namespace=locals())
-
-
-    # problem.add_equation("u = u_tot - U0")
-    problem.add_equation("trace(grad_u) + tau_p = 0")
-    problem.add_equation("dt(u) + dot(u,grad(U0)) + dot(U0,grad(u)) + grad(p) - nu*div(grad_u) + lift(tau2u,-1) = - u @ grad(u)")
[…]

### 2022-06-23 adjop/shear_cheby/main_cheby.py

@@ -1,142 +1,0 @@
-from distutils.command.bdist import show_formats
-import os
-from ast import For
-from contextlib import nullcontext
-from turtle import backward
-import numpy as np
-import sys
-sys.path.append("..")
-import h5py
-import gc
-import dedalus.public as d3
-from dedalus.core import domain
-from mpi4py import MPI
-CW = MPI.COMM_WORLD
-import logging
-import pathlib
-logger = logging.getLogger(__name__)
-from OptimizationContext import OptimizationContext
-import ForwardShear
-import BackwardShear
-import matplotlib.pyplot as plt
-path = os.path.dirname(os.path.abspath(__file__))
-
-# keys are forward variables
-# items are (backward variables, adjoint initial condition function: i.e. ux(T) = func(ux_t(T)))
-
-T = 1.0
-num_cp = 1
-dt = 2.5e-3
-
-Reynolds = 1e3
-gamma = 5e-1
-gamma_factor = 1.1
-use_euler_gradient_descend = True
-show_forward = False
-cadence = 1
-opt_iters = 1000
-
-# Bases
-Lx, Lz = 4, 1
-Nx, Nz = 256, 128
-
-dealias = 3/2
-coords = d3.CartesianCoordinates('x', 'z')
-coords.name = coords.names
-dist = d3.Distributor(coords, dtype=np.float64)
-
-xbasis = d3.RealFourier(coords['x'], size=Nx, bounds=(0, Lx), dealias=dealias)
-zbasis = d3.Chebyshev(coords['z'], size=Nz, bounds=(-Lz/2, Lz/2), dealias=dealias)
-bases = [xbasis, zbasis]
-x, z = dist.local_grids(bases[0], bases[1])
-
-domain = domain.Domain(dist, bases)
-dist = domain.dist
-
-forward_problem = ForwardShear.build_problem(domain, coords, Reynolds)
-backward_problem = BackwardShear.build_problem(domain, coords, Reynolds)
-
-# Names of the forward, and corresponding adjoint variables
-lagrangian_dict = {'u' : 'u_t'}
-
-timestepper = d3.RK443
-forward_solver = forward_problem.build_solver(timestepper)
-backward_solver = backward_problem.build_solver(timestepper)
-
-write_suffix = 'kdv0'
-
-# HT is maximized at t = T
-u = next(field for field in forward_solver.state if field.name == 'u')
-U = dist.VectorField(coords, name='U', bases=bases)
[…]

### 2022-06-23 adjop/shear_cheby/shear_cheby.py

@@ -1,131 +1,0 @@
-"""
-Dedalus script simulating a 2D periodic incompressible shear flow with a passive
-tracer field for visualization. This script demonstrates solving a 2D periodic
-initial value problem. It can be ran serially or in parallel, and uses the
-built-in analysis framework to save data snapshots to HDF5 files. The
-`plot_snapshots.py` script can be used to produce plots from the saved data.
-The simulation should take a few cpu-minutes to run.
-
-The initial flow is in the x-direction and depends only on z. The problem is
-non-dimensionalized usign the shear-layer spacing and velocity jump, so the
-resulting viscosity and tracer diffusivity are related to the Reynolds and
-Schmidt numbers as:
-
-    nu = 1 / Reynolds
-    D = nu / Schmidt
-
-To run and plot using e.g. 4 processes:
-    $ mpiexec -n 4 python3 shear_flow.py
-    $ mpiexec -n 4 python3 plot_snapshots.py snapshots/*.h5
-"""
-
-import numpy as np
-import dedalus.public as d3
-import logging
-logger = logging.getLogger(__name__)
-import os
-path = os.path.dirname(os.path.abspath(__file__))
-import sys
-
-# Parameters
-Lx, Lz = 4, 1
-Nx, Nz = 32, 16
-Reynolds = 1e3
-S = 1
-dealias = 3/2
-stop_sim_time = 1.0
-timestepper = d3.RK222
-max_timestep = 1e-3
-dtype = np.float64
-
-# Bases
-coords = d3.CartesianCoordinates('x', 'z')
-dist = d3.Distributor(coords, dtype=dtype)
-xbasis = d3.RealFourier(coords['x'], size=Nx, bounds=(0, Lx), dealias=dealias)
-zbasis = d3.ChebyshevT(coords['z'], size=Nz, bounds=(-Lz/2, Lz/2), dealias=dealias)
-
-
-# Fields
-p = dist.Field(name='p', bases=(xbasis, zbasis))
-s = dist.Field(name='s', bases=(xbasis, zbasis))
-u = dist.VectorField(coords, name='u', bases=(xbasis, zbasis))
-
-tau_p = dist.Field(name='tau_p')
-
-tau1u = dist.VectorField(coords, name='tau1u', bases=(xbasis))
-tau2u = dist.VectorField(coords, name='tau2u', bases=(xbasis))
-
[…]

### 2022-06-24 adjop/kdv/phasespace/combine_images.py
phasespace burgers nl

@@ -2,29 +2,27 @@ import sys
 from PIL import Image
 
 abvec = []
-abvec.append((0.2, 0.05))
-abvec.append((0.2, 1.0))
-abvec.append((0.2, 3.0))
-abvec.append((0.05, 0.05))
-abvec.append((0.05, 1.0))
-abvec.append((0.05, 3.0))
+abvec.append((5.0,    0.03))
+abvec.append((1.5,    0.1 ))
+abvec.append((1.0,    0.15 ))
+abvec.append((0.5,    0.3))
+abvec.append((0.15,   1.0))
+abvec.append((0.05,   3.0))
 
 paths = []
-# paths.append('objtest_a0p0b1p0.png')
-# paths.append('objtest_a0p0b8p0.png')
-# paths.append('objtest_a0p25b1p0.png')
-# paths.append('objtest_a0p5b1p0.png')
 
-for (a, b) in abvec:
+for (a, T) in abvec:
   a_str = str(a).replace('.', 'p')
-  b_str = str(b).replace('.', 'p')
-  paths.append('objtest_a{}b{}.png'.format(a_str, b_str))
+  T_str = str(T).replace('.', 'p')
+  paths.append('objtest_a{}T{}.png'.format(a_str, T_str))
 
 images = [Image.open(x) for x in paths]
 widths, heights = zip(*(i.size for i in images))
 
-total_width = round(sum(widths) / 2.0)
-max_height = round(sum(heights) / 3.0)
+rows, cols = 2, 3
+
+total_width = round(sum(widths) / rows)
+max_height = round(sum(heights) / cols)
 
 new_im = Image.new('RGB', (total_width, max_height))
 
@@ -38,4 +36,4 @@ for im in images[3:]:
   new_im.paste(im, (x_offset,im.size[1]))
   x_offset += im.size[0]
 
-new_im.save('kdv_objectives_T5.png')
+new_im.save('kdv_objectives_nuT0p15.png')

### 2022-06-24 adjop/kdv/phasespace/kdv_twomodes.py

@@ -38,11 +38,12 @@ N = config.getint('parameters', 'Nx')
 
 if (len(sys.argv) == 3):
     a = float(sys.argv[1])
-    b = float(sys.argv[2])
+    T = float(sys.argv[2])
 else:
     a = config.getfloat('parameters', 'a')
-    b = config.getfloat('parameters', 'b')
-T = config.getfloat('parameters', 'T')
+    T = config.getfloat('parameters', 'T')
+
+b = config.getfloat('parameters', 'b')
 dt = config.getfloat('parameters', 'dt')
 num_cp = config.getint('parameters', 'num_cp')
 
@@ -76,7 +77,7 @@ both = True
 write_objectives = False
 k1 = 1
 k2 = 2
-Nmodes = 100
+Nmodes = 500
 
 mode1 = np.sin(k1 * 2*np.pi * x / Lx)
 mode2 = np.sin(k2 * 2*np.pi * x / Lx)
@@ -85,8 +86,8 @@ mode2_f = dist.Field(name='mode2_f', bases=xbasis)
 mode1_f['g'] = mode1.copy()
 mode2_f['g'] = mode2.copy()
 
-k1_coeffs = np.linspace(0, 8, Nmodes)
-k2_coeffs = np.linspace(0, 8, Nmodes)
+k1_coeffs = np.linspace(-4, 12, Nmodes)
+k2_coeffs = np.linspace(-4, 12, Nmodes)
 
 def compute_coeffs(u, mode1_f, mode2_f):
     c1 = 2.0 / Lx * d3.Integrate(u*mode1_f).evaluate()['g'].flat[0]
@@ -163,13 +164,8 @@ if (write_objectives or both):
         ic = k1_coeffs[jrow] * mode1 + k2_coeffs[jcol] * mode2
         u2.change_scales(1)
         UT.change_scales(1)
-        CW.barrier()
         u2['g'] = evolve(ic, forward_solver, T, dt)
-        CW.barrier()
-        
         objectives[jrow, jcol] = evaluate_objective(u2, UT)
-        CW.Barrier()
-        CW.barrier()
         # sys.exit()
         # print(objectives[jrow, jcol])
 
@@ -212,10 +208,10 @@ if (not write_objectives or both):
                 plt.plot(c1s, c2s, color='lime', linewidth=3)
         plt.xlabel(r'$(2/L_x) \; \langle u\sin${}$x \rangle$'.format(k1))
[…]

### 2022-06-24 adjop/kdv/phasespace/main.sh

@@ -3,10 +3,15 @@
 source ~/miniconda3/etc/profile.d/conda.sh
 conda activate dedalus3
 
-MPIPROCS=1000
-mpiexec_mpt -np $MPIPROCS python3 kdv_twomodes.py 0.2 0.05
-mpiexec_mpt -np $MPIPROCS python3 kdv_twomodes.py 0.2 1.0
-mpiexec_mpt -np $MPIPROCS python3 kdv_twomodes.py 0.2 3.0
-mpiexec_mpt -np $MPIPROCS python3 kdv_twomodes.py 0.05 0.05
-mpiexec_mpt -np $MPIPROCS python3 kdv_twomodes.py 0.05 1.0
-mpiexec_mpt -np $MPIPROCS python3 kdv_twomodes.py 0.05 3.0
+MPIPROCS=1372
+# mpiexec_mpt -np $MPIPROCS python3 kdv_twomodes.py 5.0    0.03 
+# mpiexec_mpt -np $MPIPROCS python3 kdv_twomodes.py 0.5    0.3
+# mpiexec_mpt -np $MPIPROCS python3 kdv_twomodes.py 0.05   3.0
+
+# mpiexec_mpt -np $MPIPROCS python3 kdv_twomodes.py 3.0    0.05 
+mpiexec_mpt -np $MPIPROCS python3 kdv_twomodes.py 1.0    0.15 
+# mpiexec_mpt -np $MPIPROCS python3 kdv_twomodes.py 0.15   1.0
+
+# mpiexec_mpt -np $MPIPROCS python3 kdv_twomodes.py 0.05 0.05
+# mpiexec_mpt -np $MPIPROCS python3 kdv_twomodes.py 0.05 1.0
+# mpiexec_mpt -np $MPIPROCS python3 kdv_twomodes.py 0.05 3.0

### 2022-06-24 adjop/diffusion_analysis.tex

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx,amsmath,amsfonts,amssymb,mathtools}
\usepackage{enumitem}
\usepackage{empheq}
\usepackage{xcolor}

\newcommand\Ra{\mathrm{Ra}}
\newcommand\Pran{\mathrm{Pr}}
\newcommand\Rac{\mathrm{Ra}_{\mathrm{c}}}
\newcommand\Ek{\mathrm{Ek}}
\newcommand\Ro{\mathrm{Ro}}
\newcommand\Nu{\mathrm{Nu}}
\newcommand\Sc{\mathrm{Sc}}

\newcommand\eps{\varepsilon}
\renewcommand\L {\mathcal{L}}

\newcommand{\n}{\\ \nonumber \\ }
\newcommand{\nn}{\nonumber}
\newcommand{\nnn}{\\ \nonumber \\ \nonumber}

\newcommand\ie{\textit{i.e.},~}
\newcommand\eg{\textit{e.g.},~}
\newcommand{\omicron}{o}
\newcommand*\widefbox[1]{\fbox{\hspace{2em}#1\hspace{2em}}}

\newcommand{\pd}[1]{\partial_{#1}}
\newcommand{\vhat}[1]{\hat{\boldsymbol{#1}}}
\renewcommand{\vec}[1]{\boldsymbol{#1}}
\newcommand{\M}[1]{\mathbf{#1}}
\newcommand{\grad}{\vec{\nabla}}
\renewcommand{\skew}{\vec{\nabla}^{\perp}}
\newcommand{\cross}{\vec{\times}}
\newcommand{\curl}{\grad \vec{\times}}
\newcommand{\divergence}{\grad \cdot}
\newcommand{\laplacian}{\nabla^2}
\newcommand{\veclaplacian}{\grad^2}

\newcommand{\sump}[2]{\sideset{}{'}\sum_{{#1}=0}^{#2}}

\newcommand{\eq}[1]{(\ref{#1})}
\newcommand{\eqs}[2]{(\ref{#1})~\&~(\ref{#2})}
\newcommand{\eqss}[2]{(\ref{#1})--(\ref{#2})}

\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Eqs}[2]{Eqs.~(\ref{#1})~\&~(\ref{#2})}
\newcommand{\Eqss}[2]{Eqs.~(\ref{#1})--(\ref{#2})}

\newcommand{\fig}[1]{Fig.~(\ref{#1})}
\newcommand{\figs}[2]{Figs.~(\ref{#1})~\&~(\ref{#2})}
\newcommand{\T}{{\cal T}}
\newcommand{\Z}{{\cal Z}}
\newcommand{\px}{\partial_x}
\newcommand{\py}{\partial_y}
\newcommand{\pz}{\partial_z}
\setlength{\parindent}{0pt}

\begin{document}

\section*{1D: Diffusion Inversion}
\subsection*{Preliminaries}
\begin{align*}
    \intertext{We begin by defining the Lagrangian}
    \L &= \int_0^T \langle \mu \cdot [\partial_t u - \nu \partial^2_x u] \rangle dt + \frac{1}{2}\langle (u(T) - U(T))^2 \rangle
    \intertext{where $U(T)$ is the target end state, whose initial condition $U(0)$ we wish to recover. We continue by deriving the forward and adjoint problems}
    0 = \frac{\delta\L}{\delta\mu} &= \int_0^T \langle \partial_t u - \nu \partial^2_x u \rangle dt \quad \longrightarrow \quad \partial_t u = \nu \partial^2_x u 
    \intertext{Before deriving the adjoint system, we note that}
    \L &= \int_0^T \langle \mu \partial_t u - \nu \mu\partial^2_x u \rangle dt + \frac{1}{2}\langle (u(T) - U(T))^2 \rangle\\
    &= \int_0^T \langle  \partial_t (u\mu) - u\partial_t \mu - \nu \mu\partial^2_x u \rangle dt + \frac{1}{2}\langle (u(T) - U(T))^2 \rangle\\
    &= \Big\langle\int_0^T   \partial_t (u\mu) dt\Big\rangle + \int_0^T \langle- u\partial_t \mu - \nu \mu\partial^2_x u \rangle dt + \frac{1}{2}\langle (u(T) - U(T))^2 \rangle\\
    &= \langle   u(T)\mu(T) - u(0)\mu(0) \rangle + \int_0^T \langle- u\partial_t \mu - \nu \mu\partial^2_x u \rangle dt + \frac{1}{2}\langle (u(T) - U(T))^2 \rangle\\
    \intertext{We continue by deriving the adjoint system. For periodic (among other) boundary conditions}
    0 = \frac{\delta\L}{\delta u} &= -\int_0^T \langle \partial_t\mu + \nu\partial^2_x \mu \rangle dt \quad \longrightarrow \quad \partial_t \mu = -\nu\partial^2_x \mu
    \intertext{To initialize the adjoint system at $t=T$, we require}
    0 = \frac{\delta\L}{\delta u(T)} &= \langle \mu(T) + u(T) - U(T) \rangle \quad \longrightarrow \quad \mu(T) = U(T) - u(T)
    \intertext{To determine a trajectory which is coincident with $U(0)$, we take}
    \frac{\delta\L}{\delta u(0)} &= -\mu(0) 
\end{align*}

\newpage
\subsection*{Linearity Cancellation}
\begin{align*}
    \intertext{Here we demonstrate that the target state $U(0)$ manifests itself as a constant shift in the objective function wrt $u(0)$. Let $\tilde{f}$ represent a function $f$ which has undergone diffusion with diffusivity $\nu$ over a time period $T$, i.e.}
    u(0) &= f \quad \longrightarrow \quad u(T) = \tilde{f}
    \intertext{Suppose our guess's deviation from the target initial state is given by $-w$, i.e.}
    u(0) &= U(0) - w
    \intertext{Then}
    u(T) &= U(T) - \tilde{w} \quad \longrightarrow \quad \frac{1}{2}\langle (u(T) - U(T))^2\rangle = \frac{1}{2}\langle w^2 \rangle
    \intertext{which is independent of the target state. So the objective only depends on our deviation from the target state, rather than the target state itself.}
\end{align*}

\subsection*{Objective Level Sets}
\begin{align*}
    \intertext{Given that the target state doesn't matter for the following analysis, we take $U(0) = U(T) = 0$ for all $x$}
    \intertext{Suppose}
    u(0) &= a_1 e^{i k_1 x} + a_2 e^{i k_2 x} + ...
    \intertext{We wish to characterize the objective's dependence on the initial guess in this $n$-dimensional phase space $\vec{a} = (a_1, a_2, ..., a_n) \in \mathbb{C}^n$. We can solve for $u(T)$ analytically}
    u(T) &= a_1 e^{i k_1 x - \nu k_1^2 T} + a_2 e^{i k_2 x - \nu k_2^2 T} + ...\\
    &= a_1 e^{- \nu k_1^2 T}e^{i k_1 x } + a_2 e^{- \nu k_2^2 T}e^{i k_2 x } + ...
    \intertext{The objective }
    \frac{1}{2}\langle |u(T)|^2 \rangle &= \frac{1}{2}\langle \big( a_1 e^{- \nu k_1^2 T}e^{i k_1 x } + a_2 e^{- \nu k_2^2 T}e^{i k_2 x } + ...\big)^2 \rangle \\
    &= \frac{1}{2}\Big[a_1^2 e^{- 2\nu k_1^2 T} \langle \cos^2(k_1 x) + \sin^2(k_1 x) \rangle + a_2^2 e^{- 2\nu k_2^2 T} \langle \cos^2(k_2 x) + \sin^2(k_2 x)  \rangle  + ... \Big]\\
    &= \big(\pi e^{- 2\nu k_1^2 T}\big)\; a_1^2 + \big(\pi e^{- 2\nu k_2^2 T}\big) \; a_2^2 + ... + \big(\pi e^{- 2\nu k_n^2 T}\big) \; a_n^2  \\
    \intertext{Therefore the objective's level sets are given by infinite-dimensional hyperellipses in $\vec{a}$ space.}
\end{align*}

\subsection*{Preconditioner}
\begin{align*}
    \intertext{The appropriate preconditioner should be obvious: we need to undiffuse the adjoint initial condition twice! Doing so remaps the levelsets of hyperellipses into hyperspheres, thereby ensuring that the gradient is always in the direction of $u(0) - U(0)$}
    \intertext{Suppose the fourier coefficients of the adjoint initial condition are given by $\vec{b} = (b_1, b_2, b_3, ..., b_n) \in \mathbb{C}^n$. Then it would be advantageous to initialize the adjoint system with fourier coefficients $\tilde{\vec{b}}$ where}
    \begin{bmatrix}
        \tilde{b_1}\\
        \tilde{b_2}\\
        \tilde{b_3}\\
        \vdots \\
        \tilde{b_n}\\
    \end{bmatrix}
    &= 
    \begin{bmatrix}
        e^{2\nu k_1^2T} & 0 & 0 & \hdots & 0\\
        0 & e^{2\nu k_2^2T} & 0 & \hdots & 0\\
        0 & 0 & e^{2\nu k_3^2T} & \hdots & 0\\
        \vdots & \vdots & \vdots & \ddots & \vdots \\
        0 & 0 & 0 & \hdots & e^{2\nu k_n^2T}\\
    \end{bmatrix}
    \begin{bmatrix}
        {b_1}\\
        {b_2}\\
        {b_3}\\
        \vdots \\
        {b_n}\\
    \end{bmatrix}
\end{align*}

\end{document}

### 2022-07-03 adjop/lsrtm/lsrtm.tex
lsrtm

--- 2022-06-24 adjop/diffusion_analysis.tex
@@
 \begin{document}
-
-\section*{1D: Diffusion Inversion}
-\subsection*{Preliminaries}
-\begin{align*}
-    \intertext{We begin by defining the Lagrangian}
-    \L &= \int_0^T \langle \mu \cdot [\partial_t u - \nu \partial^2_x u] \rangle dt + \frac{1}{2}\langle (u(T) - U(T))^2 \rangle
-    \intertext{where $U(T)$ is the target end state, whose initial condition $U(0)$ we wish to recover. We continue by deriving the forward and adjoint problems}
-    0 = \frac{\delta\L}{\delta\mu} &= \int_0^T \langle \partial_t u - \nu \partial^2_x u \rangle dt \quad \longrightarrow \quad \partial_t u = \nu \partial^2_x u 
-    \intertext{Before deriving the adjoint system, we note that}
-    \L &= \int_0^T \langle \mu \partial_t u - \nu \mu\partial^2_x u \rangle dt + \frac{1}{2}\langle (u(T) - U(T))^2 \rangle\\
-    &= \int_0^T \langle  \partial_t (u\mu) - u\partial_t \mu - \nu \mu\partial^2_x u \rangle dt + \frac{1}{2}\langle (u(T) - U(T))^2 \rangle\\
-    &= \Big\langle\int_0^T   \partial_t (u\mu) dt\Big\rangle + \int_0^T \langle- u\partial_t \mu - \nu \mu\partial^2_x u \rangle dt + \frac{1}{2}\langle (u(T) - U(T))^2 \rangle\\
-    &= \langle   u(T)\mu(T) - u(0)\mu(0) \rangle + \int_0^T \langle- u\partial_t \mu - \nu \mu\partial^2_x u \rangle dt + \frac{1}{2}\langle (u(T) - U(T))^2 \rangle\\
-    \intertext{We continue by deriving the adjoint system. For periodic (among other) boundary conditions}
-    0 = \frac{\delta\L}{\delta u} &= -\int_0^T \langle \partial_t\mu + \nu\partial^2_x \mu \rangle dt \quad \longrightarrow \quad \partial_t \mu = -\nu\partial^2_x \mu
-    \intertext{To initialize the adjoint system at $t=T$, we require}
-    0 = \frac{\delta\L}{\delta u(T)} &= \langle \mu(T) + u(T) - U(T) \rangle \quad \longrightarrow \quad \mu(T) = U(T) - u(T)
-    \intertext{To determine a trajectory which is coincident with $U(0)$, we take}
-    \frac{\delta\L}{\delta u(0)} &= -\mu(0) 
-\end{align*}
-
-\newpage
-\subsection*{Linearity Cancellation}
-\begin{align*}
-    \intertext{Here we demonstrate that the target state $U(0)$ manifests itself as a constant shift in the objective function wrt $u(0)$.
-Let $\tilde{f}$ represent a function $f$ which has undergone diffusion with diffusivity $\nu$ over a time period $T$, i.e.}
-    u(0) &= f \quad \longrightarrow \quad u(T) = \tilde{f}
-    \intertext{Suppose our guess's deviation from the target initial state is given by $-w$, i.e.}
-    u(0) &= U(0) - w
-    \intertext{Then}
-    u(T) &= U(T) - \tilde{w} \quad \longrightarrow \quad \frac{1}{2}\langle (u(T) - U(T))^2\rangle = \frac{1}{2}\langle w^2 \rangle
-    \intertext{which is independent of the target state. So the objective only depends on our deviation from the target state, rather than the target state itself.}
-\end{align*}
-
-\subsection*{Objective Level Sets}
-\begin{align*}
-    \intertext{Given that the target state doesn't matter for the following analysis, we take $U(0) = U(T) = 0$ for all $x$}
-    \intertext{Suppose}
-    u(0) &= a_1 e^{i k_1 x} + a_2 e^{i k_2 x} + ...
-    \intertext{We wish to characterize the objective's dependence on the initial guess in this $n$-dimensional phase space $\vec{a} = (a_1, a_2, ..., a_n) \in \mathbb{C}^n$.
-We can solve for $u(T)$ analytically}
-    u(T) &= a_1 e^{i k_1 x - \nu k_1^2 T} + a_2 e^{i k_2 x - \nu k_2^2 T} + ...\\
-    &= a_1 e^{- \nu k_1^2 T}e^{i k_1 x } + a_2 e^{- \nu k_2^2 T}e^{i k_2 x } + ...
-    \intertext{The objective }
-    \frac{1}{2}\langle |u(T)|^2 \rangle &= \frac{1}{2}\langle \big( a_1 e^{- \nu k_1^2 T}e^{i k_1 x } + a_2 e^{- \nu k_2^2 T}e^{i k_2 x } + ...\big)^2 \rangle \\
-    &= \frac{1}{2}\Big[a_1^2 e^{- 2\nu k_1^2 T} \langle \cos^2(k_1 x) + \sin^2(k_1 x) \rangle + a_2^2 e^{- 2\nu k_2^2 T} \langle \cos^2(k_2 x) + \sin^2(k_2 x)  \rangle  + ... \Big]\\
-    &= \big(\pi e^{- 2\nu k_1^2 T}\big)\; a_1^2 + \big(\pi e^{- 2\nu k_2^2 T}\big) \; a_2^2 + ... + \big(\pi e^{- 2\nu k_n^2 T}\big) \; a_n^2  \\
-    \intertext{Therefore the objective's level sets are given by infinite-dimensional hyperellipses in $\vec{a}$ space.}
-\end{align*}
-
-\subsection*{Preconditioner}
-\begin{align*}
-    \intertext{The appropriate preconditioner should be obvious: we need to undiffuse the adjoint initial condition twice!
-Doing so remaps the levelsets of hyperellipses into hyperspheres, thereby ensuring that the gradient is always in the direction of $u(0) - U(0)$}
-    \intertext{Suppose the fourier coefficients of the adjoint initial condition are given by $\vec{b} = (b_1, b_2, b_3, ..., b_n) \in \mathbb{C}^n$.
-Then it would be advantageous to initialize the adjoint system with fourier coefficients $\tilde{\vec{b}}$ where}
-    \begin{bmatrix}
-        \tilde{b_1}\\
-        \tilde{b_2}\\
-        \tilde{b_3}\\
-        \vdots \\
-        \tilde{b_n}\\
-    \end{bmatrix}
-    &= 
-    \begin{bmatrix}
-        e^{2\nu k_1^2T} & 0 & 0 & \hdots & 0\\
-        0 & e^{2\nu k_2^2T} & 0 & \hdots & 0\\
-        0 & 0 & e^{2\nu k_3^2T} & \hdots & 0\\
-        \vdots & \vdots & \vdots & \ddots & \vdots \\
-        0 & 0 & 0 & \hdots & e^{2\nu k_n^2T}\\
-    \end{bmatrix}
-    \begin{bmatrix}
-        {b_1}\\
-        {b_2}\\
-        {b_3}\\
-        \vdots \\
-        {b_n}\\
-    \end{bmatrix}
-\end{align*}
+\title{Reverse Time Migration (RTM) Least-Squares Objectives}
 

### 2022-07-03 adjop/lsrtm/publication_materials/ms_rbc.tex
lsrtm

\documentclass[longbibliography,twocolumn,amsmath,amssymb,aps,nofootinbib]{revtex4-2}

\usepackage{graphicx}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{longtable}
\usepackage{gensymb}
\usepackage{enumerate}
\usepackage{varwidth}
\usepackage{float}
\usepackage{nonfloat}
\usepackage{soul}
\usepackage{ulem}
% \usepackage{multicol}
% \usepackage{afterpage}
\usepackage[usenames, dvipsnames]{color}
\newcommand{\note}[1]{\noindent \textbf{\textit{\textcolor{Red}{#1}}}}

\newcommand\Ra{\mathrm{Ra}}
\newcommand\Pran{\mathrm{Pr}}
\newcommand\Rac{\mathrm{Ra}_{\mathrm{c}}}
\newcommand\Ek{\mathrm{Ek}}
\newcommand\Ro{\mathrm{Ro}}
\newcommand\Nu{\mathrm{Nu}}
\newcommand\Sc{\mathrm{Sc}}

\newcommand\eps{\varepsilon}
\renewcommand\L {\mathcal{L}}
\renewcommand{\citet}[1]{ref.~\cite{#1}}
\renewcommand{\Citet}[1]{Ref.~\cite{#1}}
\newcommand{\citets}[1]{refs.~\cite{#1}}
\newcommand{\Citets}[1]{Refs.~\cite{#1}}

\newcommand{\n}{\\ \nonumber \\ }
\newcommand{\nn}{\nonumber}
\newcommand{\nnn}{\\ \nonumber \\ \nonumber}

\newcommand\ie{\textit{i.e.},~}
\newcommand\eg{\textit{e.g.},~}
\newcommand{\omicron}{o}

\newcommand{\pd}[1]{\partial_{#1}}
\renewcommand{\vec}[1]{\boldsymbol{#1}}
\newcommand{\M}[1]{\mathbf{#1}}
\newcommand{\grad}{\vec{\nabla}}
\newcommand{\cross}{\vec{\times}}
\newcommand{\laplacian}{\nabla^2}

\newcommand{\sump}[2]{\sideset{}{'}\sum_{{#1}=0}^{#2}}

\newcommand{\eq}[1]{(\ref{#1})}
\newcommand{\eqs}[2]{(\ref{#1})~\&~(\ref{#2})}
\newcommand{\eqss}[2]{(\ref{#1})--(\ref{#2})}

\newcommand{\Eq}[1]{Eq.~(\ref{#1})}
\newcommand{\Eqs}[2]{Eqs.~(\ref{#1})~\&~(\ref{#2})}
\newcommand{\Eqss}[2]{Eqs.~(\ref{#1})--(\ref{#2})}

\newcommand{\fig}[1]{Fig.~(\ref{#1})}
\newcommand{\figs}[2]{Figs.~(\ref{#1})~\&~(\ref{#2})}
\newcommand{\T}{{\cal T}}
\newcommand{\Z}{{\cal Z}}

\makeatletter
\let\Hy@backout\@gobble
\makeatother

\newcommand*{\GtrSim}{\smallrel\gtrsim}

\makeatletter
\newcommand*{\smallrel}[2][.8]{%
  \mathrel{\mathpalette{\smallrel@{#1}}{#2}}%
}
\newcommand*{\smallrel@}[3]{%
  % #1: scale factor
  % #2: math style
  % #3: symbol
  \sbox0{$#2\vcenter{}$}%
  \dimen@=\ht0 %
  \raise\dimen@\hbox{%
    \scalebox{#1}{%
      \raise-\dimen@\hbox{$#2#3\m@th$}%
    }%
  }%
}
\makeatother

\begin{document}

\title{Marginally-Stable Thermal Equilibria of Rayleigh-B\'enard Convection}

\author{Liam O'Connor$^{1,2}$}
\affiliation{%
$^1$Department of Engineering Sciences and Applied Mathematics, Northwestern University, Evanston, IL 60208 USA}
\affiliation{%
$^2$Center for Interdisciplinary Exploration and Research in Astrophysics, Northwestern University, Evanston, IL, 60201 USA}

\begin{abstract}
    Natural convection exhibits turbulent flows which are difficult or impossible to resolve in direct numerical simulations. In this work, we investigate a quasilinear form of the Rayleigh-B\'enard problem which describes the bulk one-dimensional properties of convection without resolving the turbulent dynamics. We represent perturbations away from the mean using a sum of marginally-stable eigenmodes.  By constraining the perturbation amplitudes, the marginal stability criterion allows us to evolve the background temperature profile under the influence of multiple eigenmodes representing flows at different length scales.
    We find the quasilinear system evolves to an equilibrium state where advective and diffusive fluxes sum to a constant.
    These marginally-stable thermal equilibria (MSTE) are exact solutions of the quasilinear equations.
    The mean MSTE temperature profiles have thinner boundary layers and larger Nusselt numbers than thermally-equilibrated 2D and 3D simulations of the full nonlinear equations.  
    MSTE solutions exhibit a classic boundary-layer scaling of the Nusselt number $\Nu$ with the Rayleigh number $\Ra$ of $\Nu \sim \Ra^{1/3}$.
    When an MSTE is used as initial conditions for a 2D simulation, we find that Nu quickly equilibrates without the burst of turbulence often induced by purely conductive initial conditions, but we also find that the kinetic energy is too large and viscously attenuates on a long viscous time scale.
\end{abstract}

\maketitle

\section{Introduction}
Rayleigh-B\'enard convection plays a foundational role in astrophysical and geophysical settings.
The resulting buoyancy-driven flows regulate heat transfer and generate large-scale vortices \cite{Couston}.
Turbulent convection, which is associated with large Rayleigh numbers $\Ra$, is difficult to simulate. 
State of the art simulations performed by \cite{Zhu_2018} have reached $\Ra \sim 10^{14}$ but estimates for the sun's convective zone and earth's interior are $\Ra \sim 10^{16}-10^{20}$ and $\Ra \sim 10^{20}-10^{30}$ respectively \cite{Ossendrijver,Gubbins_2001}.

\bibliography{ms_rbc.bib}

\end{document}

### 2022-07-03 adjop/diffusion/ForwardDiffusion.py

@@ -43,5 +43,6 @@ def build_problem(domain, xcoord, a):
 
     # Problem
     problem = d3.IVP([u], namespace=locals())
-    problem.add_equation("dt(u) - a*dx(dx(u)) = 0")
+    # problem.add_equation("dt(u) = u*dx(u)")
+    problem.add_equation("dt(u) - 2000*a*dx(dx(u)) - a*dx(dx(dx(u))) = 0")
     return problem

### 2022-07-03 adjop/diffusion/diffusion_twomodes.py

@@ -66,8 +66,8 @@ forward_solver = forward_problem.build_solver(d3.RK443)
 both = True
 write_objectives = False
 k1 = 1
-k2 = 6
-Nmodes = 11
+k2 = 2
+Nmodes = 31
 
 mode1 = np.sin(k1 * 2*np.pi * x / Lx)
 mode2 = np.sin(k2 * 2*np.pi * x / Lx)
@@ -97,10 +97,10 @@ def diffuse(ic, forward_solver, T, dt):
     return forward_solver.state[0]
 
 def evaluate_objective(u):
-    ux = d3.Differentiate(u, xcoord)
-    uxx = d3.Differentiate(ux, xcoord)
-    fac1 = d3.Integrate(np.exp(a*np.abs(ux))).evaluate()['g'].flat[0]
-    fac2 = d3.Integrate(ux * ux).evaluate()['g'].flat[0]
+    # ux = d3.Differentiate(u, xcoord)
+    # uxx = d3.Differentiate(ux, xcoord)
+    # fac1 = d3.Integrate(np.exp(a*np.abs(ux))).evaluate()['g'].flat[0]
+    fac2 = d3.Integrate(u * u).evaluate()['g'].flat[0]
     return fac2
     # return fac1 + fac2
 
@@ -145,6 +145,6 @@ if (not write_objectives or both):
     plt.xlabel(r'$(2/L_x) \; \langle u\sin${}$x \rangle$'.format(k1))
     plt.ylabel(r'$(2/L_x) \; \langle u\sin${}$x \rangle$'.format(k2))
     plt.title(r'$\langle u^2 \rangle$')
-    plt.savefig(path + '/objtest_k' + str(k1) + 'k' + str(k2) + '.png')
+    plt.savefig(path + '/burgobjtest_k' + str(k1) + 'k' + str(k2) + '.png')
     plt.show()
 

### 2022-07-06 adjop/lsrtm/lsrtm.tex
cleanup lsrtm

@@
-\documentclass{article}
-\usepackage[utf8]{inputenc}
-\usepackage{graphicx,amsmath,amsfonts,amssymb,mathtools}
-\usepackage{enumitem}
-\usepackage{empheq}
-\usepackage{xcolor}
+\documentclass[longbibliography,twocolumn,amsmath,amssymb,aps,nofootinbib]{revtex4-2}
+
+\usepackage{graphicx}
+\usepackage{natbib}
+\usepackage{amsmath}
+\usepackage{longtable}
+\usepackage{gensymb}
+\usepackage{enumerate}
+\usepackage{varwidth}
+\usepackage{float}
+\usepackage{nonfloat}
+\usepackage{soul}
+\usepackage{ulem}
+% \usepackage{multicol}
+% \usepackage{afterpage}
+\usepackage[usenames, dvipsnames]{color}
+\newcommand{\note}[1]{\noindent \textbf{\textit{\textcolor{Red}{#1}}}}
 
@@
 \renewcommand\L {\mathcal{L}}
-
+\renewcommand{\citet}[1]{ref.~\cite{#1}}
+\renewcommand{\Citet}[1]{Ref.~\cite{#1}}
+\newcommand{\citets}[1]{refs.~\cite{#1}}
+\newcommand{\Citets}[1]{Refs.~\cite{#1}}
+\newcommand{\davg}[1]{\langle {#1} \rangle}
 \newcommand{\n}{\\ \nonumber \\ }
@@
 \newcommand{\omicron}{o}
-\newcommand*\widefbox[1]{\fbox{\hspace{2em}#1\hspace{2em}}}
 
 \newcommand{\pd}[1]{\partial_{#1}}
-\newcommand{\vhat}[1]{\hat{\boldsymbol{#1}}}
 \renewcommand{\vec}[1]{\boldsymbol{#1}}
@@
 \newcommand{\grad}{\vec{\nabla}}
-\renewcommand{\skew}{\vec{\nabla}^{\perp}}
 \newcommand{\cross}{\vec{\times}}
-\newcommand{\curl}{\grad \vec{\times}}
-\newcommand{\divergence}{\grad \cdot}
 \newcommand{\laplacian}{\nabla^2}
-\newcommand{\veclaplacian}{\grad^2}
 
@@
 \newcommand{\Z}{{\cal Z}}
-\newcommand{\px}{\partial_x}
-\newcommand{\py}{\partial_y}
-\newcommand{\pz}{\partial_z}
-\setlength{\parindent}{0pt}
+
+\makeatletter
+\let\Hy@backout\@gobble
+\makeatother
+
+\newcommand*{\GtrSim}{\smallrel\gtrsim}
+
+\makeatletter
+\newcommand*{\smallrel}[2][.8]{%
+  \mathrel{\mathpalette{\smallrel@{#1}}{#2}}%
+}
+\newcommand*{\smallrel@}[3]{%
+  % #1: scale factor
+  % #2: math style
+  % #3: symbol
+  \sbox0{$#2\vcenter{}$}%
+  \dimen@=\ht0 %
+  \raise\dimen@\hbox{%
+    \scalebox{#1}{%
+      \raise-\dimen@\hbox{$#2#3\m@th$}%
+    }%
+  }%
+}
+\makeatother
 
 \begin{document}
-\title{Reverse Time Migration (RTM) Least-Squares Objectives}
+
+\title{Least-Squares Time Reversal of PDEs}
+
+\author{Liam O'Connor$^{1,2}$}
+\affiliation{%
+$^1$Department of Engineering Sciences and Applied Mathematics, Northwestern University, Evanston, IL 60208 USA}
+\affiliation{%
+$^2$Center for Interdisciplinary Exploration and Research in Astrophysics, Northwestern University, Evanston, IL, 60201 USA}
+
+\begin{abstract}
+    abstract
+\end{abstract}
+
+\maketitle
+
+\section{Introduction}
+Rayleigh-B\'enard convection plays a foundational role in astrophysical and geophysical settings.
+The resulting buoyancy-driven flows regulate heat transfer and generate large-scale vortices \cite{Couston}.
+Turbulent convection, which is associated with large Rayleigh numbers $\Ra$, is difficult to simulate. 
+State of the art simulations performed by \cite{Zhu_2018} have reached $\Ra \sim 10^{14}$ but estimates for the sun's convective zone and earth's interior are $\Ra \sim 10^{16}-10^{20}$ and $\Ra \sim 10^{20}-10^{30}$ respectively \cite{Ossendrijver,Gubbins_2001}.
+
+\clearpage
+\section{Generalized Problem Setup}
+Here I'll present a generalized PDE which we want to invert along with a bunch of notation.
+We'll define our target simulation, least-squares objective, and cite examples where the adjoint looping method has been successfully applied to inverse problems.
+We'll conclude this discussion by stressing the important of the objective function's shape in ``initial condition space'', because ultimately this is what curtails our gradient-based inversion technique.
+
+\section{Analytic Analysis}
+\subsection{Linearity and Reversiblity}
+Reversible linear problems are trivial cases of the problems we intend to study. 
+Linear systems' objective functions' shapes do not depend on the target--the target manifests itself as a constant shift in ic space. 
+Thus we might as well consider problems with $U(0) = U(T) = 0$ for all $x$.
+Reversible wave equations, which conserve energy yield objective functions whose levelsets in ic space are spherical, so the gradient at any given guess always provides the shortest path to the target.
+In contrast, irreversible diffusive terms distort the spherical levelsets into hyperellipses, whose semi-minimum (maximum) axis is oriented in the direction of the lowest (highest) wavenumber fourier mode's coefficient.
+Thus, when traversing an arbitrary path of steepest descent toward the target IC, low wavenumber modes are matched with the target first, followed by increasingly higher wavenumber modes.
+
+\subsection{Nonlinear Reversible Problems}
+Here the target matters. We consider burger's equation
+\begin{align*}
+  \partial_t u + u\partial_x u &= 0
+  \intertext{where energy is conserved}
+  0 &= \partial_t \langle \frac{1}{2}u^2 \rangle + \langle u^2\partial_x u\rangle \\
+  &= \partial_t \langle \frac{1}{2}u^2 \rangle + \langle  \frac{1}{3}\partial_x u^3\rangle \\
+  &= \partial_t \langle \frac{1}{2}u^2 \rangle \\
+\end{align*}
+Next we consider a guess with a known deviation from the target $u(x, 0) = U(x, 0) + u'(x, 0)$.
+Our objective function is given by $\langle (u(T) - U(T))^2 \rangle = \langle u'(T)^2 \rangle$.
+The forward evolution equation for the perturbation is given by
+
+\begin{align*}
+  0 &= \partial_t u' + u'\partial_x U + U\partial_x u' + u'\partial_x u' \\
+  \intertext{then we have that}
+  \intertext{the evolution of the target is given by}
+  0 &= \frac{1}{2}\partial_t \davg{u'^2} + \davg{u' \partial_x (u'U)} + \davg{\frac{1}{3}\partial_x u'^3} \\
+  &= \frac{1}{2}\partial_t \davg{u'^2} + \davg{u' \partial_x (u'U)} \\
+  &= \frac{1}{2}\partial_t \davg{u'^2} + \davg{U \partial_x (u'^2)} \\
+  \intertext{Suppose $u' = \alpha U$.
+In this case $u = U + u' = (1+\alpha)U$ evolves identically to $U$ but at a rate which is $(1 + \alpha)^2$ faster.
+Thus the objective is given by $\davg{(U((1+\alpha)^2 T) - U(T))^2}$}.
+  % \intertext{Suppose the perturbation is small enough that its nonlinear term can be neglected. In this case we find that}
+  u'(T) &= u'(0) + \int_0^T \partial_tu'dt \\
+  &= u'(0)-\int_0^T \partial_x (u'U) + \frac{1}{2}\partial_x (u'^2) dt
+  \intertext{Therefore $\davg{u'(T)} = \davg{u'(0)}$. The objective function}
+  \davg{u'(T)^2}&= \davg{u'(0)^2}- \davg{u'(0)\int_0^T \partial_x (u'U) + \frac{1}{2}\partial_x (u'^2) dt} \\
+  &\;\; + \davg{\Big[\int_0^T \partial_x (u'U) + \frac{1}{2}\partial_x (u'^2) dt\Big]^2} \\
+\end{align*}
+
+\clearpage
+\bibliography{lsrtm.bib}
 

### 2022-07-08 adjop/lsrtm/lsrtm.tex
section 1 and 2

@@
 \section{Generalized Problem Setup}
-Here I'll present a generalized PDE which we want to invert along with a bunch of notation.
+Given a generalized initial value problem
+\begin{align}
+  \mathcal{F}[U] &\equiv \partial_t U + L[U] + N[U, U] = 0 \label{genpde}
+  \intertext{and a target end state $U(x, T)$, our aim is to recover the associated initial condition $U(x, 0)$.  
+  This is realized by performing gradient-based minimization of the functional}
+  H[u, U] &= \frac{1}{2}\davg{(u(x, T) - U(x, T))^2} = \frac{1}{2}\davg{u'(x, T)^2} \label{Hdef}, 
+\end{align}
+where the operator $\davg{\cdot}$ denotes the domain-integrated average and the quantity $u'(x, t)$ refers to $u(x, t)$'s the deviation from the target solution.
+For the remainder of this paper, capital symbols (such as $U(x, t)$) will be used to denote the target solution, while their corresponding lowercase symbols (such as $u(x, t)$) play the role of optimization parameters.
+
+The Lagrangian associated with our initial value problem is given by
+\begin{align*}
+  \mathcal{L} &= \int_0^T \davg{\mu \cdot \mathcal{F}[u]} dt + H[u, U].
+  \intertext{Where $\mu(x, t)$ is the Lagrangian Multiplier corresponding to $u(x, t)$. Equating} \frac{\delta\L}{\delta u} &= 0 \quad \text{and} \quad \frac{\delta\L}{\delta \mu} = 0 
+  \intertext{yields the original and adjoint systems} 
+  \mathcal{F}[u] &= 0 \quad \text{and} \quad \mathcal{F}^{\dagger}[\mu, u] = 0 \intertext{respectively. At some particular guess $u^n(x, 0)$, we compute the functional derivative}
+  \frac{\delta\L}{\delta u^n(x, 0)} &= -\mu^n(x, 0)
+  \intertext{to determine the direction of steepest descent of our objective functional $H[u, U]$ with respect to the optimization parameter $u(x, 0)$.
+Naive gradient-descent can then be performed by computing an updated guess}
+  u^{n+1}(x, 0) &= u^n(x, 0) + \varepsilon \mu^n(x, 0)
+  \intertext{where $\varepsilon$ is the step size.}
+\end{align*}
+In addition to this algorithm, more advanced gradient-based optimization techniques such as conjugate gradient and BFGS have been employed in conjunction with adjoint-looping to improve performance and accelerate convergence.
+In the context of our time-inversion optimization problem, such techniques traverse the space of possible initial conditions $u(x, 0)$ by relying on local information pertaining to the objective functional $H$ and its functional derivative $\frac{\delta H}{\delta u(x, 0)}$.
+For this reason, the focus of our investigation is largely centered around the objective functional's dependence on the initial condition.
+More specifically, we study the levelsets of the objective functional because these surfaces characterize the directions of steepest descent at any arbitrary point in the space of initial conditions.
+In doing so, we highlight the challenges and deficits associated with using the least squares functional to recover the initial condition $U(x, 0)$ belonging to a target end state $U(x, T)$.
+
+% Here I'll present a generalized PDE which we want to invert along with a bunch of notation.
 We'll define our target simulation, least-squares objective, and cite examples where the adjoint looping method has been successfully applied to inverse problems.
@@
 \subsection{Linearity and Reversiblity}
-Reversible linear problems are trivial cases of the problems we intend to study. 
+In this section, we study linear initial value problems which obey (\ref{genpde}) along with the property $\mathcal{F}[u'] = \mathcal{F}[u] - \mathcal{F}[U]$. 
+In these special cases, the objective functional (\ref{Hdef}) depends only on the deviation's initial condition $u'(x, 0)$ about any arbitrary target $U$.
+In the following section, we demonstrate that the objective functionals of nonlinear systems with nonzero targets are directory influenced by one's choice of $U$.
+
+time-reversible initial value problems are trivial cases of this optimization problem. 
 Linear systems' objective functions' shapes do not depend on the target--the target manifests itself as a constant shift in ic space. 

### 2022-07-08 adjop/OptimizationContext.py
kdv4

@@ -403,7 +403,7 @@ class OptimizationContext:
    # This works really well for periodic kdv
     def compute_gamma(self, epsilon_safety):
         if (self.loop_index == 1):
-            return 1e-3
+            return self.gamma_init
         else:
             # https://en.wikipedia.org/wiki/Gradient_descent
             grad_diff = self.new_grad - self.old_grad
@@ -420,7 +420,9 @@ class OptimizationContext:
                 old_grad_sqrd = 0.0
             gamma = CW.bcast(gamma, root=0)
             self.old_grad_sqrd = CW.bcast(old_grad_sqrd, root=0)
-            return gamma
+        if (epsilon_safety == 0.0):
+            return self.gamma_init
+        return gamma
 
 #     def descend(self, gamma, **kwargs):
 

### 2022-07-08 adjop/lsrtm/lsrtm.tex
sum brgr

@@
   This is realized by performing gradient-based minimization of the functional}
-  H[u, U] &= \frac{1}{2}\davg{(u(x, T) - U(x, T))^2} = \frac{1}{2}\davg{u'(x, T)^2} \label{Hdef}, 
+  H[u, U] &= \frac{1}{2}\davg{(u(x, T) - U(x, T))^2} = \frac{1}{2}\davg{u'^2(x, T)} \label{Hdef}, 
 \end{align}
@@
 
-\section{Analytic Analysis}
+\section{Preliminary Analytic Analysis}
 \subsection{Linearity and Reversiblity}
@@
 In these special cases, the objective functional (\ref{Hdef}) depends only on the deviation's initial condition $u'(x, 0)$ about any arbitrary target $U$.
+For convenience when dealing with linear problems, we take $U(0) = U(T) = 0$ for all $x$.
 In the following section, we demonstrate that the objective functionals of nonlinear systems with nonzero targets are directory influenced by one's choice of $U$.
 
-time-reversible initial value problems are trivial cases of this optimization problem. 
-Linear systems' objective functions' shapes do not depend on the target--the target manifests itself as a constant shift in ic space. 
-Thus we might as well consider problems with $U(0) = U(T) = 0$ for all $x$.
+Time-reversible wave equations of the form
+\begin{align}
+  \partial_t u &= \sum_{n \geq 1} c_n \partial^{2n - 1}_x u \label{waveeq}
+  \intertext{are trivial cases of this optimization problem, because the quantity $\davg{u^2(x, t)}$ is conserved. 
+  This implies that the objective functional's levelsets form sphere's in $u'$-space, because its value is given by $\davg{u'^2(T)} = \davg{u'^2(0)}$.
+  The spatial derivatives in (\ref{waveeq}) are anti-self-adjoint (i.e. $L[u] = -L^{\dagger}[u]$), so the path of steepest descent is always in the radial derivation }
+  \frac{\delta H}{\delta u(x, 0)} &= -u'(x, 0) \nonumber
+  \intertext{In contrast, diffusive equations of the form}
+  \partial_t u &= \sum_{n \geq 1} b_n \partial^{2n}_x u \label{waveeq}
+  \intertext{do not conserve $\davg{u^2(x, t)}$, and consequently their objective functionals depend on $u'(x, 0)$'s \textit{direction} along with its magnitude.
+Suppose initial deviation can be written as}
+  u'(x, 0) &= \sum_{k \geq 0} a_k e^{ikx}.
+  \intertext{then}
+  \davg{u'(x, T)} &= \pi \sum_{k \geq 0} \Big[\exp\Big({-2T\sum_{n \geq 1} k^{2n}  }\Big)a_k^2\Big].
+\end{align}
+When the initial deviation $u'(x, 0)$ is described using a basis of fourier coefficients, its levelsets form hyperellipses, whose semi-major (semi-minor) axes are aligned with $u'(x, 0)$'s largest (smallest) fourier coefficient.
+Accordingly, the path of steepest descent is only aligned with $u'(x, 0)$ when $u'(x, 0)$ is composed of a single fourier mode. Such systems are self-adjoint, meaning $L[u] = L^{\dagger}[u]$
+
+\textbf{summary: }Linear systems' objective functions' shapes do not depend on the target--the target manifests itself as a constant shift in ic space. 
 Reversible wave equations, which conserve energy yield objective functions whose levelsets in ic space are spherical, so the gradient at any given guess always provides the shortest path to the target.
@@
 Next we consider a guess with a known deviation from the target $u(x, 0) = U(x, 0) + u'(x, 0)$.
-Our objective function is given by $\langle (u(T) - U(T))^2 \rangle = \langle u'(T)^2 \rangle$.
+Our objective function is given by $\langle (u(T) - U(T))^2 \rangle = \langle u'^2(T) \rangle$.
 The forward evolution equation for the perturbation is given by

### 2022-07-08 adjop/kdv/kdv_burgers.py
spheres

@@ -87,8 +87,8 @@ else:
 
 # Initial conditions
 x = dist.local_grid(xbasis)
-# u['g'] = ic_scale*np.exp(-np.power(x - mu, 2.) / (2 * np.power(sig, 2.)))
-u['g'] = 2*np.sin(2*np.pi*x / Lx) + 2*np.sin(4*np.pi*x / Lx)
+u['g'] = ic_scale*np.exp(-np.power(x - mu, 2.) / (2 * np.power(sig, 2.)))
+# u['g'] = 2*np.sin(2*np.pi*x / Lx) + 2*np.sin(4*np.pi*x / Lx)
 
 # Solver
 solver = problem.build_solver(timestepper)

### 2022-07-08 adjop/kdv/kdv_parallel.py

@@ -59,6 +59,10 @@ show_forward = config.getboolean('parameters', 'show')
 show_iter_cadence = config.getint('parameters', 'show_iter_cadence')
 show_loop_cadence = config.getint('parameters', 'show_loop_cadence')
 
+sig = config.getfloat('parameters', 'sig_ic')
+mu = config.getfloat('parameters', 'mu_ic')
+ic_scale = config.getfloat('parameters', 'ic_scale')
+
 # Bases
 xcoord = d3.Coordinate('x')
 dist = d3.Distributor(xcoord, dtype=np.float64, comm=MPI.COMM_SELF)
@@ -132,20 +136,23 @@ Nics = 20
 if (CW.size != Nics):
     raise
 
-theta = np.linspace(0, 2*np.pi, CW.size)[CW.rank]
+opt.ic['u']['g'] = ic_scale*np.exp(-np.power(x - mu, 2.) / (2 * np.power(sig, 2.)))
 
-if restart:
-    mode1_coeff = 2 + np.cos(theta)
-    mode2_coeff = 2 + np.sin(theta)
-    opt.ic['u']['g'] = mode1_coeff*mode1['g'] + mode2_coeff*mode2['g']
+R = 1.0
+modes_dim = 5
+np.random.seed(CW.rank)
+coeffs = 2*np.random.rand(5) - 1.0
+coeffs *= R / np.sqrt(np.sum(coeffs**2))
+for kx, coeff in enumerate(coeffs):
+    opt.ic['u']['g'] += coeff*np.cos((kx + 1)*2*np.pi*x / Lx)
 
 # opt.ic['u']['g'] = 2.987*mode1['g'] + 7.3*mode2['g']
 
-opt.metrics0['A1'] = opt.ic['u']*mode1 / Lx * 2.0
-opt.metrics0['A2'] = opt.ic['u']*mode2 / Lx * 2.0
-opt.metrics0['Arem'] = opt.ic['u']*opt.ic['u'] - (opt.metrics0['A1']**2 + opt.metrics0['A2']**2) * Lx * 2.0
+# # opt.metrics0['A1'] = opt.ic['u']*mode1 / Lx * 2.0
+# # opt.metrics0['A2'] = opt.ic['u']*mode2 / Lx * 2.0
+# # opt.metrics0['Arem'] = opt.ic['u']*opt.ic['u'] - (opt.metrics0['A1']**2 + opt.metrics0['A2']**2) * Lx * 2.0
 
[…]

### 2022-07-08 adjop/kdv/main.sh

@@ -1,6 +1,6 @@
 #PBS -S /bin/bash
 #PBS -l select=1:ncpus=28:mpiprocs=28:model=bro
-#PBS -l walltime=10:00:00
+#PBS -l walltime=8:00:00
 #PBS -j oe
 #PBS -W group_list=s2276
 file=${0##*/}
@@ -26,7 +26,7 @@ CONFIG="kdv_options.cfg"
 # SUFFIX="T5_N256_reverse"
 # OLDSUFFIX=$SUFFIX
 
-# python3 kdv_burgers.py
+python3 kdv_burgers.py
 mpiexec_mpt -np 20 python3 kdv_parallel.py
 mpiexec_mpt -np 1 python3 paths_pod.py
 # python3 kdv_cg.py

### 2022-07-11 adjop/lsrtm/lsrtm.tex
postmalone

@@
 
+\begin{figure}
+  \includegraphics[width=9cm]{../diffusion/obj_k1k5.png}
+  \caption{Objective function and an example gradient descent path for the diffusion equation.
+The target simulation $U(t) = 0$ and the deviation is composed of two fourier modes, i.e.
+$u'(t) = a_1\sin x + a_5\sin 5x$.
+Because the system is linear self-adjoint, each subsequent guess in the gradient descent trajectory is constrained to lie in this 2D phase space.}
+\end{figure}
+
 \textbf{summary: }Linear systems' objective functions' shapes do not depend on the target--the target manifests itself as a constant shift in ic space. 
@@
 Thus, when traversing an arbitrary path of steepest descent toward the target IC, low wavenumber modes are matched with the target first, followed by increasingly higher wavenumber modes.
-
+\clearpage
 \subsection{Nonlinear Reversible Problems}
@@
 
-\begin{align*}
-  0 &= \partial_t u' + u'\partial_x U + U\partial_x u' + u'\partial_x u' \\
+\begin{align}
+  0 &= \partial_t u' + u'\partial_x U + U\partial_x u' + u'\partial_x u' \label{pertbrg}\\
   \intertext{then we have that}
   \intertext{the evolution of the target is given by}
-  0 &= \frac{1}{2}\partial_t \davg{u'^2} + \davg{u' \partial_x (u'U)} + \davg{\frac{1}{3}\partial_x u'^3} \\
-  &= \frac{1}{2}\partial_t \davg{u'^2} + \davg{u' \partial_x (u'U)} \\
-  &= \frac{1}{2}\partial_t \davg{u'^2} + \davg{U \partial_x (u'^2)} \\
-  \intertext{Suppose $u' = \alpha U$.
+  0 &= \frac{1}{2}\partial_t \davg{u'^2} + \davg{u' \partial_x (u'U)} + \davg{\frac{1}{3}\partial_x u'^3} \nonumber \\
+  &= \frac{1}{2}\partial_t \davg{u'^2} + \davg{u' \partial_x (u'U)} \nonumber \\
+  &= \frac{1}{2}\partial_t \davg{u'^2} - \frac{1}{2}\davg{U \partial_x (u'^2)} \nonumber \\
+  &\therefore\nonumber\\
+  \davg{u'^2(T)} &= \davg{u'^2(0)} + \int_0^T \davg{U \partial_x (u'^2)} dt \label{objburgeq}
+  % \intertext{Suppose $u' = \alpha U$.
 In this case $u = U + u' = (1+\alpha)U$ evolves identically to $U$ but at a rate which is $(1 + \alpha)^2$ faster.
 Thus the objective is given by $\davg{(U((1+\alpha)^2 T) - U(T))^2}$}.
-  % \intertext{Suppose the perturbation is small enough that its nonlinear term can be neglected. In this case we find that}
-  u'(T) &= u'(0) + \int_0^T \partial_tu'dt \\
-  &= u'(0)-\int_0^T \partial_x (u'U) + \frac{1}{2}\partial_x (u'^2) dt
-  \intertext{Therefore $\davg{u'(T)} = \davg{u'(0)}$. The objective function}
-  \davg{u'(T)^2}&= \davg{u'(0)^2}- \davg{u'(0)\int_0^T \partial_x (u'U) + \frac{1}{2}\partial_x (u'^2) dt} \\
-  &\;\; + \davg{\Big[\int_0^T \partial_x (u'U) + \frac{1}{2}\partial_x (u'^2) dt\Big]^2} \\
-\end{align*}
+  % % \intertext{Suppose the perturbation is small enough that its nonlinear term can be neglected. In this case we find that}
+  % u'(T) &= u'(0) + \int_0^T \partial_tu'dt \\
+  % &= u'(0)-\int_0^T \partial_x (u'U) + \frac{1}{2}\partial_x (u'^2) dt
+  % \intertext{Therefore $\davg{u'(T)} = \davg{u'(0)}$. The objective function}
+  % \davg{u'(T)^2}&= \davg{u'(0)^2}- \davg{u'(0)\int_0^T \partial_x (u'U) + \frac{1}{2}\partial_x (u'^2) dt} \\
+  % &\;\; + \davg{\Big[\int_0^T \partial_x (u'U) + \frac{1}{2}\partial_x (u'^2) dt\Big]^2} \\
+\end{align}
+
+\subsection*{Numerical Examples}
+Suppose the target solution is $U(t) = 0$.
+Then the perturbation equation (\ref{pertbrg}) is linear in $u'$, and we revert back to the hyperspherical levelsets described for wave equations in the previous section, as shown in Fig \ref{brgU0}.
+
+\begin{figure}[H]
+  \includegraphics[width=9cm]{../diffusion/burgobjtest_k1k2_U0.png}
+  \caption{Objective functional the inviscid burger's equation.
+The target simulation $U(x, 0) = 0$ and the deviation is composed of two fourier modes, i.e.
+$u'(t) = a_1\sin x + a_5\sin 2x$.
+Gradient descent paths will not deviate from this 2D phase space because the path of steepest descent is always in the radial direction.}
+  \label{brgU0}
+\end{figure}
+
+Suppose instead that our target initial condition is given by $U(x, 0) = 5\sin(x)$ and we are interested in the 2D phase space of initial deviations $u'(x, 0) = a_1\sin x + a_2 \sin 2x$, as shown in Fig \ref{brgU5m1}..
+
+\begin{figure}[H]
+  \includegraphics[width=9cm]{../diffusion/burgobjtest_k1k2_U5m1.png}
+  \caption{Objective functional the inviscid burger's equation.
+The target simulation $U(x, 0) = 5\sin x$ and the deviation is composed of two fourier modes, i.e.
+$u'(t) = a_1\sin x + a_5\sin 2x$.
+Gradient descent paths will deviate from this 2D phase space because external fourier modes are populated by the nonlinearities in the forward and adjoint equations.}
+  \label{brgU5m1}
+\end{figure}
 

### 2022-07-11 adjop/diffusion/diffusion_twomodes.py

@@ -81,7 +81,7 @@ mode2_f['g'] = mode2.copy()
 k1_coeffs = np.linspace(-0.01, 0.01, Nmodes)
 k2_coeffs = np.linspace(-0.01, 0.01, Nmodes)
 
-targetic = -5.0*mode2.copy()
+targetic = 0.0*mode1.copy()
 
 def compute_coeffs(u, mode1_f, mode2_f):
     c1 = 2.0 / Lx * d3.Integrate(u*mode1_f).evaluate()['g'].flat[0]
@@ -152,7 +152,8 @@ if (not write_objectives or both):
             plt.plot(c1s, c2s, color='lime', linewidth=3)
     plt.xlabel(r'$(2/L_x) \; \langle u\sin${}$x \rangle$'.format(k1))
     plt.ylabel(r'$(2/L_x) \; \langle u\sin${}$x \rangle$'.format(k2))
-    plt.title(r'$\langle u^2 \rangle$')
-    plt.savefig(path + '/burgobjtest_k' + str(k1) + 'k' + str(k2) + '_Un5m2.png')
+    plt.title(r'$\langle {u\prime}^2(T) \rangle$')
+    plt.savefig(path + '/burgobjtest_k' + str(k1) + 'k' + str(k2) + '_U0.png')
+    # plt.savefig(path + '/burgobjtest_k' + str(k1) + 'k' + str(k2) + '_U5m1.png')
     plt.show()
 

### 2022-07-11 adjop/kdv/kdv_burgers.py

@@ -90,6 +90,14 @@ x = dist.local_grid(xbasis)
 u['g'] = ic_scale*np.exp(-np.power(x - mu, 2.) / (2 * np.power(sig, 2.)))
 # u['g'] = 2*np.sin(2*np.pi*x / Lx) + 2*np.sin(4*np.pi*x / Lx)
 
+R = 1.0
+modes_dim = 5
+np.random.seed(0)
+coeffs = 2*np.random.rand(5) - 1.0
+coeffs *= R / np.sqrt(np.sum(coeffs**2))
+for kx, coeff in enumerate(coeffs):
+    u['g'] += coeff*np.cos((kx + 1)*2*np.pi*x / Lx)
+
 # Solver
 solver = problem.build_solver(timestepper)
 solver.stop_sim_time = stop_sim_time

### 2022-07-12 adjop/lsrtm/twomodes.py
figure annotations

@@ -1,23 +1,16 @@
-from distutils.command.bdist import show_formats
 import os
 path = os.path.dirname(os.path.abspath(__file__))
-from ast import For
-from contextlib import nullcontext
-from turtle import backward
 import numpy as np
 import pickle
 import sys
 sys.path.append(path + "/..")
 sys.path.append(path + "/../diffusion")
 sys.path.append(path + "/../kdv")
-import h5py
-import gc
 import dedalus.public as d3
 from dedalus.core.domain import Domain
 from mpi4py import MPI
 CW = MPI.COMM_WORLD
 import logging
-import pathlib
 logger = logging.getLogger(__name__)
 import ForwardKDV
 import ForwardDiffusion
@@ -67,6 +60,16 @@ T = config.getfloat('parameters', 'T')
 
 dt = config.getfloat('parameters', 'dt')
 num_cp = config.getint('parameters', 'num_cp')
+
+a_str = str(a).replace('.', 'p')
+b_str = str(b).replace('.', 'p')
+T_str = str(T).replace('.', 'p')
+kt1_str = str(target_coeffs[0]).replace('.', 'p')
+kt2_str = str(target_coeffs[1]).replace('.', 'p')
+R_str = str(R).replace('.', 'p')
+
+objectives_str = path + '/objectives_a' + a_str + 'b' + b_str + 'T' + T_str + 'R' + R_str + 'kt1' + kt1_str + 'kt2' + kt2_str + '.txt'
+save_dir = path + '/objtest_a' + a_str + 'b' + b_str + 'T' + T_str + 'R' + R_str + 'kt1' + kt1_str + 'kt2' + kt2_str + '.png'
 
 # Simulation Parameters
 dealias = 3/2
@@ -211,17 +214,24 @@ if (write_objectives or both):
         CW.Reduce(objectives, objectives, op=MPI.SUM, root=0)
 
     if CW.rank == 0:
-        np.savetxt(path + '/objectives_' + problem + '.txt', objectives)
+        np.savetxt(objectives_str, objectives)
         logger.info('saved final state')
 
 
 if (not write_objectives or both):
     if CW.rank == 0:
-        objectives = np.loadtxt(path + '/objectives_' + problem + '.txt')
+        def fmt(x):
+            print(str(x))
+            return '  ' + str(x) + '  '
+
+        objectives = np.loadtxt(objectives_str)
         # print(objectives)
[…]

### 2022-07-13 adjop/lsrtm/twomodes.py
generalized kdvburg

@@ -56,6 +56,7 @@ N = config.getint('parameters', 'Nx')
 
 a = config.getfloat('parameters', 'a')
 b = config.getfloat('parameters', 'b')
+c = config.getfloat('parameters', 'c')
 T = config.getfloat('parameters', 'T')
 
 dt = config.getfloat('parameters', 'dt')
@@ -63,13 +64,14 @@ num_cp = config.getint('parameters', 'num_cp')
 
 a_str = str(a).replace('.', 'p')
 b_str = str(b).replace('.', 'p')
+c_str = str(c).replace('.', 'p')
 T_str = str(T).replace('.', 'p')
 kt1_str = str(target_coeffs[0]).replace('.', 'p')
 kt2_str = str(target_coeffs[1]).replace('.', 'p')
 R_str = str(R).replace('.', 'p')
 
-objectives_str = path + '/objectives_a' + a_str + 'b' + b_str + 'T' + T_str + 'R' + R_str + 'kt1' + kt1_str + 'kt2' + kt2_str + '.txt'
-save_dir = path + '/objtest_a' + a_str + 'b' + b_str + 'T' + T_str + 'R' + R_str + 'kt1' + kt1_str + 'kt2' + kt2_str + '.png'
+objectives_str = path + '/objectives_a' + a_str + 'b' + b_str + 'c' + c_str + 'T' + T_str + 'R' + R_str + 'kt1' + kt1_str + 'kt2' + kt2_str + '.txt'
+save_dir = path + '/objtest_a' + a_str + 'b' + b_str + 'c' + c_str + 'T' + T_str + 'R' + R_str + 'kt1' + kt1_str + 'kt2' + kt2_str + '.png'
 
 # Simulation Parameters
 dealias = 3/2

### 2022-07-13 adjop/lsrtm/ForwardKDV.py
graddesc path

@@ -1,0 +1,48 @@
+import numpy as np
+import dedalus.public as d3
+from mpi4py import MPI
+CW = MPI.COMM_WORLD
+import logging
+logger = logging.getLogger(__name__)
+import sys
+import matplotlib.pyplot as plt
+
+def build_problem(domain, xcoord, a, b, c):
+
+    # unpack domain
+    dist = domain.dist
+    xbasis = domain.bases[0]
+    
+    u = dist.Field(name='u', bases=xbasis)
+    dx = lambda A: d3.Differentiate(A, xcoord)
+
+    if (not isinstance(xbasis, d3.RealFourier)):
+        tau_1 = dist.Field(name='tau_1')
+        tau_2 = dist.Field(name='tau_2')     
+        tau_3 = dist.Field(name='tau_3')
+
+        lift_basis = xbasis.clone_with(a=1/2, b=1/2) # First derivative basis
+        lift = lambda A: d3.Lift(A, lift_basis, -1)
+        
+        # Substitutions
+        ux = dx(u) + lift(tau_1) # First-order reduction
+        uxx = dx(ux) + lift(tau_2) # First-order reduction
+
+        # Problem
+        problem = d3.IVP([u, tau_1, tau_2, tau_3], namespace=locals())
+        problem.add_equation("dt(u) - a*dx(ux) - b*dx(uxx) + lift(tau_3) = -c*u*ux")
+
+        problem.add_equation("u(x='left') = 0")
+        problem.add_equation("u(x='right') = 0")
+        problem.add_equation("ux(x='left') = 0")
+
+        # problem.add_equation("u(x='left') - u(x='right') = 0")
+        # problem.add_equation("ux(x='left') - ux(x='right') = 0")
+        # problem.add_equation("uxx(x='left') - uxx(x='right') = 0")
+        return problem
+        
+
+    # Problem
+    problem = d3.IVP([u], namespace=locals())
+    problem.add_equation("dt(u) - a*dx(dx(u)) - b*dx(dx(dx(u))) = -u*dx(u)")
+    return problem

### 2022-07-13 adjop/lsrtm/twomodes.py

@@ -5,7 +5,7 @@ import pickle
 import sys
 sys.path.append(path + "/..")
 sys.path.append(path + "/../diffusion")
-sys.path.append(path + "/../kdv")
+# sys.path.append(path + "/../kdv")
 import dedalus.public as d3
 from dedalus.core.domain import Domain
 from mpi4py import MPI
@@ -39,7 +39,7 @@ logger.info(config.items('parameters'))
 problem = str(config.get('parameters', 'problem'))
 both = config.getboolean('parameters', 'both')
 write_objectives = config.getboolean('parameters', 'write_objectives')
-
+plot_paths = config.getboolean('parameters', 'plot_paths')
 
 ks = ast.literal_eval(config.get('parameters', 'ks'))
 k1, k2 = ks[0], ks[1]
@@ -96,7 +96,7 @@ x = dist.local_grid(xbasis)
 CW.Barrier()
 
 if problem == 'kdv':
-    forward_problem = ForwardKDV.build_problem(domain, xcoord, a, b)
+    forward_problem = ForwardKDV.build_problem(domain, xcoord, a, b, c)
 
 elif problem == 'diffusion':
     forward_problem = ForwardDiffusion.build_problem(domain, xcoord, a, b)
@@ -235,6 +235,27 @@ if (not write_objectives or both):
         plt.clabel(cont, cont.levels, inline=True, fmt=fmt, fontsize=7)
         plt.legend(loc='lower right')
         epsilon = 1.0
+        if plot_paths:
+            grad_ar = np.gradient(objectives)
+            from scipy import interpolate
+            k1k1, k2k2 = np.meshgrid(k1_coeffs, k2_coeffs)
+            gradinterp1 = interpolate.interp2d(k1k1, k2k2, grad_ar[0], kind='cubic')
+            gradinterp2 = interpolate.interp2d(k1k1, k2k2, grad_ar[1], kind='cubic')
+            
+            sdpath = [(k1_coeffs[0], k2_coeffs[0])]
+            eps = 0.001
+            for stp in range(1000):
+                print(stp)
+                lastpt = sdpath[-1]
+                grad1 = gradinterp1(lastpt[0], lastpt[1])
+                grad2 = gradinterp2(lastpt[0], lastpt[1])
+                sdpath.append((lastpt[0] + eps*grad1, lastpt[1] + eps*grad2))
+            print(sdpath)
+            uzpath = list(zip(*sdpath))
+            k1path = uzpath[0]
[…]

### 2022-07-13 adjop/lsrtm/lsrtm.tex
misc ppr

@@
 \section{Generalized Problem Setup}
-Given a generalized initial value problem
+We define a generalized solution state vector $\vec{U}$ and associated initial value problem
 \begin{align}
-  \mathcal{F}[U] &\equiv \partial_t U + L[U] + N[U, U] = 0 \label{genpde}
-  \intertext{and a target end state $U(x, T)$, our aim is to recover the associated initial condition $U(x, 0)$.  
+  \mathcal{F}[\vec{U}] &\equiv \partial_t \vec{U} + L[\vec{U}] + N[\vec{U}, \vec{U}] = 0 \label{genpde}
+  \intertext{Given a target end state $\vec{U}(\vec{x}, T)$, our aim is to recover the associated initial condition $\vec{U}(x, 0)$.  
   This is realized by performing gradient-based minimization of the functional}
-  H[u, U] &= \frac{1}{2}\davg{(u(x, T) - U(x, T))^2} = \frac{1}{2}\davg{u'^2(x, T)} \label{Hdef}, 
+  H[\vec{u}, \vec{U}] &= \frac{1}{2}\davg{|\vec{u}(\vec{x}, T) - \vec{U}(\vec{x}, T)|^2} = \frac{1}{2}\davg{\vec{u}'^2(\vec{x}, T)} \label{Hdef}, 
 \end{align}
-where the operator $\davg{\cdot}$ denotes the domain-integrated average and the quantity $u'(x, t)$ refers to $u(x, t)$'s the deviation from the target solution.
-For the remainder of this paper, capital symbols (such as $U(x, t)$) will be used to denote the target solution, while their corresponding lowercase symbols (such as $u(x, t)$) play the role of optimization parameters.
+where the operator $\davg{\cdot}$ denotes the domain-integrated average and the quantity $\vec{u}'(\vec{x}, T)$ refers to $\vec{u}(\vec{x}, T)$'s the deviation from the target solution.
+For the remainder of this paper, capital symbols (such as $\vec{u}(\vec{x}, T)$) will be used to denote the target solution, while their corresponding lowercase symbols (such as $\vec{u}(\vec{x}, T)$) play the role of optimization parameters.
 
@@
 \begin{align*}
-  \mathcal{L} &= \int_0^T \davg{\mu \cdot \mathcal{F}[u]} dt + H[u, U].
-  \intertext{Where $\mu(x, t)$ is the Lagrangian Multiplier corresponding to $u(x, t)$. Equating} \frac{\delta\L}{\delta u} &= 0 \quad \text{and} \quad \frac{\delta\L}{\delta \mu} = 0 
+  \mathcal{L} &= \int_0^T \davg{\vec{\mu} \cdot \mathcal{F}[\vec{u}]} dt + H[\vec{u}, \vec{U}].
+  \intertext{Where $\vec{\mu}(\vec{x}, T)$ is the Lagrangian Multiplier corresponding to $\vec{u}(\vec{x}, T)$.
+Equating} \frac{\delta\L}{\delta \vec{u}} &= 0 \quad \text{and} \quad \frac{\delta\L}{\delta \vec{\mu}} = 0
   \intertext{yields the original and adjoint systems} 
-  \mathcal{F}[u] &= 0 \quad \text{and} \quad \mathcal{F}^{\dagger}[\mu, u] = 0 \intertext{respectively. At some particular guess $u^n(x, 0)$, we compute the functional derivative}
-  \frac{\delta\L}{\delta u^n(x, 0)} &= -\mu^n(x, 0)
-  \intertext{to determine the direction of steepest descent of our objective functional $H[u, U]$ with respect to the optimization parameter $u(x, 0)$.
+  \mathcal{F}[\vec{u}] &= 0 \quad \text{and} \quad \mathcal{F}^{\dagger}[\vec{\mu}, u] = 0 \intertext{respectively.
+At some particular guess $\vec{u}^n(\vec{x}, 0)$, we compute the functional derivative}
+  \frac{\delta\L}{\delta \vec{u}^n(\vec{x}, 0)} &= -\vec{\mu}^n(\vec{x}, 0)
+  \intertext{to determine the direction of steepest descent of our objective functional $H[\vec{u}, \vec{U}]$ with respect to the optimization parameter $\vec{u}(\vec{x}, 0)$.
 Naive gradient-descent can then be performed by computing an updated guess}
-  u^{n+1}(x, 0) &= u^n(x, 0) + \varepsilon \mu^n(x, 0)
+  \vec{u}^{n+1}(\vec{x}, 0) &= \vec{u}^n(\vec{x}, 0) + \varepsilon \vec{\mu}^n(\vec{x}, 0)
   \intertext{where $\varepsilon$ is the step size.}
@@
 In addition to this algorithm, more advanced gradient-based optimization techniques such as conjugate gradient and BFGS have been employed in conjunction with adjoint-looping to improve performance and accelerate convergence.
-In the context of our time-inversion optimization problem, such techniques traverse the space of possible initial conditions $u(x, 0)$ by relying on local information pertaining to the objective functional $H$ and its functional derivative $\frac{\delta H}{\delta u(x, 0)}$.
+In the context of our time-inversion optimization problem, such techniques traverse the space of possible initial conditions $\vec{u}(\vec{x}, 0)$ by relying on local information pertaining to the objective functional $H$ and its functional derivative $\frac{\delta H}{\delta \vec{u}(\vec{x}, 0)}$.
 For this reason, the focus of our investigation is largely centered around the objective functional's dependence on the initial condition.
 More specifically, we study the levelsets of the objective functional because these surfaces characterize the directions of steepest descent at any arbitrary point in the space of initial conditions.
-In doing so, we highlight the challenges and deficits associated with using the least squares functional to recover the initial condition $U(x, 0)$ belonging to a target end state $U(x, T)$.
+In doing so, we highlight the challenges and deficits associated with using the least squares functional to recover the initial condition $\vec{U}(\vec{x}, 0)$ belonging to a target end state $\vec{U}(\vec{x}, T)$.
 
@@
 \begin{align*}
-  \partial_t u + u\partial_x u &= 0
+  \partial_t u + u\partial_x u &= b \, \partial_x^3 u
   \intertext{where energy is conserved}
-  0 &= \partial_t \langle \frac{1}{2}u^2 \rangle + \langle u^2\partial_x u\rangle \\
+  0 &= \partial_t \langle \frac{1}{2}u^2 \rangle + \langle u^2\partial_x u\rangle - b\davg{u\partial_x^3 u}\\
   &= \partial_t \langle \frac{1}{2}u^2 \rangle + \langle  \frac{1}{3}\partial_x u^3\rangle  \\
   &= \partial_t \langle \frac{1}{2}u^2 \rangle \\
+  \\
+  \partial_x^3 (u*u) &= 2\partial_x^2[ u\partial_x u ] \\
+  &= 2\partial_x[ (\partial_x u)^2 + u\partial_x^2 u ]\\
+  &= 2\partial_x[ (\partial_x u)^2] + \partial_x u\partial_x^2 u + u\partial_x^3 u \\
+  &= \frac{5}{2} \partial_x[ (\partial_x u)^2] + u\partial_x^3 u \\
 \end{align*}
@@
 \begin{align}
-  0 &= \partial_t u' + u'\partial_x U + U\partial_x u' + u'\partial_x u' \label{pertbrg}\\
+  0 &= \partial_t u' + u'\partial_x U + U\partial_x u' + u'\partial_x u' - b\partial_x^3 u'\label{pertbrg}\\
   \intertext{then we have that}
@@
 \begin{figure}[H]
-  \includegraphics[width=9cm]{../diffusion/burgobjtest_k1k2_U0.png}
+  \includegraphics[width=9cm]{objtest_a0p0b0p0T3p0R0p01kt10p0kt20.png}
+  % \includegraphics[width=9cm]{../diffusion/burgobjtest_k1k2_U0.png}
   \caption{Objective functional the inviscid burger's equation.
@@
 \begin{figure}[H]
-  \includegraphics[width=9cm]{../diffusion/burgobjtest_k1k2_U5m1.png}
+  \includegraphics[width=9cm]{objtest_a0p0b0p0T3p0R0p01kt10p15kt20.png}
   \caption{Objective functional the inviscid burger's equation.

### 2022-07-13 adjop/lsrtm/twomodes.py
diffusion path

@@ -233,29 +233,38 @@ if (not write_objectives or both):
         plt.scatter([target_coeffs[0]], [target_coeffs[1]], s=60, marker='*', color='k', label = 'target')
         cont = plt.contour(k1_coeffs.ravel(), k2_coeffs.ravel(), objectives.T, levels=[0.00001, 0.00002, 0.00004], colors=['k', 'k'])
         plt.clabel(cont, cont.levels, inline=True, fmt=fmt, fontsize=7)
-        plt.legend(loc='lower right')
         epsilon = 1.0
         if plot_paths:
             grad_ar = np.gradient(objectives)
             from scipy import interpolate
             k1k1, k2k2 = np.meshgrid(k1_coeffs, k2_coeffs)
-            gradinterp1 = interpolate.interp2d(k1k1, k2k2, grad_ar[0], kind='cubic')
-            gradinterp2 = interpolate.interp2d(k1k1, k2k2, grad_ar[1], kind='cubic')
+            stride = 10
+            gradinterp1 = interpolate.interp2d(k1k1[::stride, ::stride], k2k2[::stride, ::stride], grad_ar[0][::stride, ::stride].T, kind='linear')
+            gradinterp2 = interpolate.interp2d(k1k1[::stride, ::stride], k2k2[::stride, ::stride], grad_ar[1][::stride, ::stride].T, kind='linear')
             
-            sdpath = [(k1_coeffs[0], k2_coeffs[0])]
-            eps = 0.001
-            for stp in range(1000):
-                print(stp)
+            sdpath = [(-0.01, -0.01)]
+            plt.scatter([sdpath[-1][0]], [sdpath[-1][1]], s=60, marker='o', color='k', label = 'guess')
+
+            eps = 0.000005
+            stp = 0
+            # for stp in range(100):
+            while (stp < 10000 and sdpath[-1][0]**2 + sdpath[-1][0]**2 > 1e-8):
                 lastpt = sdpath[-1]
                 grad1 = gradinterp1(lastpt[0], lastpt[1])
                 grad2 = gradinterp2(lastpt[0], lastpt[1])
-                sdpath.append((lastpt[0] + eps*grad1, lastpt[1] + eps*grad2))
-            print(sdpath)
+                grad_mag = (grad1**2 + grad2**2)**0.5
+                grad1 /= grad_mag
+                grad2 /= grad_mag
+                print('step {}, pt {}, grad1 {}, grad2 {}'.format(stp, lastpt, grad1, grad2))
[…]

### 2022-07-13 adjop/lsrtm/lsrtm.tex
minor ppr changes

@@
 \begin{align}
-  \mathcal{F}[\vec{U}] &\equiv \partial_t \vec{U} + L[\vec{U}] + N[\vec{U}, \vec{U}] = 0 \label{genpde}
+  \mathcal{F}[\vec{U}] &\equiv \partial_t \vec{U} + L[\vec{U}] + N[\vec{U}, \vec{U}] = 0. \label{genpde}
   \intertext{Given a target end state $\vec{U}(\vec{x}, T)$, our aim is to recover the associated initial condition $\vec{U}(x, 0)$.  
@@
 where the operator $\davg{\cdot}$ denotes the domain-integrated average and the quantity $\vec{u}'(\vec{x}, T)$ refers to $\vec{u}(\vec{x}, T)$'s the deviation from the target solution.
-For the remainder of this paper, capital symbols (such as $\vec{u}(\vec{x}, T)$) will be used to denote the target solution, while their corresponding lowercase symbols (such as $\vec{u}(\vec{x}, T)$) play the role of optimization parameters.
+For the remainder of this paper, capital symbols (such as $\vec{U}(\vec{x}, T)$) will be used to denote the target solution, while their corresponding lowercase symbols (such as $\vec{u}(\vec{x}, T)$) play the role of optimization parameters.
 
@@
 For convenience when dealing with linear problems, we take $U(0) = U(T) = 0$ for all $x$.
-In the following section, we demonstrate that the objective functionals of nonlinear systems with nonzero targets are directory influenced by one's choice of $U$.
+In the following section, we demonstrate that the objective functionals of nonlinear systems with nonzero targets are directly influenced by one's choice of $U$.
 
@@
 \begin{figure}
-  \includegraphics[width=9cm]{../diffusion/obj_k1k5.png}
+  \includegraphics[width=9cm]{objtest_a0p03b0p0c0p0T3p0R0p015kt10p0kt20.png}
   \caption{Objective function and an example gradient descent path for the diffusion equation.

### 2022-07-14 adjop/kdv/kdv_parallel.py
spheres again

@@ -179,7 +179,7 @@ def euler_descent(fun, x0, args, **kwargs):
             with open(tracker_name, 'wb') as file:
                 pickle.dump(opt.descent_tracker, file)
 
-        x0 -= 1e4 * gamma * gradf / np.sum(gradf**2)
+        # x0 -= 1e4 * gamma * gradf / np.sum(gradf**2)
     logger.info('success')
     logger.info('maxiter = {}'.format(maxiter))
     return optimize.OptimizeResult(x=x0, success=True, message='beep boop')

### 2022-07-14 adjop/kdv/paths_pod.py

@@ -139,3 +139,4 @@ for comp_ind in range(comps):
     plt.ylabel('Iteration')
     plt.title('Descent Paths POD: SV = {:.2f}'.format(pca.singular_values_[comp_ind]))
     plt.savefig(path + '/' + write_suffix + '/pod_comp{}.png'.format(comp_ind))
+    logger.info('figure saved to ' + path + '/' + write_suffix + '/pod_comp{}.png'.format(comp_ind))

### 2022-07-14 adjop/kdv/kdv_burgers.py
fixed bugs kdv parallel

@@ -90,13 +90,13 @@ x = dist.local_grid(xbasis)
 u['g'] = ic_scale*np.exp(-np.power(x - mu, 2.) / (2 * np.power(sig, 2.)))
 # u['g'] = 2*np.sin(2*np.pi*x / Lx) + 2*np.sin(4*np.pi*x / Lx)
 
-R = 1.0
-modes_dim = 5
-np.random.seed(0)
-coeffs = 2*np.random.rand(5) - 1.0
-coeffs *= R / np.sqrt(np.sum(coeffs**2))
-for kx, coeff in enumerate(coeffs):
-    u['g'] += coeff*np.cos((kx + 1)*2*np.pi*x / Lx)
+# R = 1.0
+# modes_dim = 5
+# np.random.seed(0)
+# coeffs = 2*np.random.rand(5) - 1.0
+# coeffs *= R / np.sqrt(np.sum(coeffs**2))
+# for kx, coeff in enumerate(coeffs):
+#     u['g'] += coeff*np.cos((kx + 1)*2*np.pi*x / Lx)
 
 # Solver
 solver = problem.build_solver(timestepper)

### 2022-07-16 adjop/lsrtm/lsrtm.tex
misc lsrtm reach

@@
 \section{Generalized Problem Setup}
-We define a generalized solution state vector $\vec{U}$ and associated initial value problem
+We define a generalized solution state vector $\bar{X}$ and associated initial value problem
 \begin{align}
-  \mathcal{F}[\vec{U}] &\equiv \partial_t \vec{U} + L[\vec{U}] + N[\vec{U}, \vec{U}] = 0. \label{genpde}
-  \intertext{Given a target end state $\vec{U}(\vec{x}, T)$, our aim is to recover the associated initial condition $\vec{U}(x, 0)$.  
+  \mathcal{F}[\bar{X}] &\equiv \partial_t \bar{X} + L . \bar{X} + N[\bar{X}, \bar{X}] = 0. \label{genpde}
+  \intertext{Given a target end state $\bar{X}(\vec{x}, T)$, our aim is to recover the associated initial condition $\bar{X}(\vec{x}, 0)$.  
   This is realized by performing gradient-based minimization of the functional}
-  H[\vec{u}, \vec{U}] &= \frac{1}{2}\davg{|\vec{u}(\vec{x}, T) - \vec{U}(\vec{x}, T)|^2} = \frac{1}{2}\davg{\vec{u}'^2(\vec{x}, T)} \label{Hdef}, 
+  H[X, \bar{X}] &= \frac{1}{2}\davg{|X(\vec{x}, T) - \bar{X}(\vec{x}, T)|^2} = \frac{1}{2}\davg{X'^2(\vec{x}, T)} \label{Hdef}, 
 \end{align}
-where the operator $\davg{\cdot}$ denotes the domain-integrated average and the quantity $\vec{u}'(\vec{x}, T)$ refers to $\vec{u}(\vec{x}, T)$'s the deviation from the target solution.
-For the remainder of this paper, capital symbols (such as $\vec{U}(\vec{x}, T)$) will be used to denote the target solution, while their corresponding lowercase symbols (such as $\vec{u}(\vec{x}, T)$) play the role of optimization parameters.
+with respect to $X(\vec{x}, 0)$, where $X(\vec{x}, t)$ obeys (\ref{genpde}).
+The operator $\davg{\cdot}$ denotes the domain-integrated average and the quantity $X'(\vec{x}, t) \equiv X(\vec{x}, t) - \bar{X}(\vec{x}, t)$.
+For the remainder of this paper, symbols with overbars (such as $\bar{X}(\vec{x}, T)$) will be used to denote the target solution, while their corresponding unannotated symbols (such as $X(\vec{x}, T)$) play the role of optimization parameters.
 
@@
 \begin{align*}
-  \mathcal{L} &= \int_0^T \davg{\vec{\mu} \cdot \mathcal{F}[\vec{u}]} dt + H[\vec{u}, \vec{U}].
-  \intertext{Where $\vec{\mu}(\vec{x}, T)$ is the Lagrangian Multiplier corresponding to $\vec{u}(\vec{x}, T)$.
-Equating} \frac{\delta\L}{\delta \vec{u}} &= 0 \quad \text{and} \quad \frac{\delta\L}{\delta \vec{\mu}} = 0
+  \mathcal{L} &= \int_0^T \davg{\chi \cdot \mathcal{F}[X]} dt + H[X, \bar{X}].
+  \intertext{Where $\chi(\vec{x}, T)$ is the Lagrangian Multiplier corresponding to $X(\vec{x}, T)$. Equating} \frac{\delta\L}{\delta X} &= 0 \quad \text{and} \quad \frac{\delta\L}{\delta \chi} = 0 
   \intertext{yields the original and adjoint systems} 
-  \mathcal{F}[\vec{u}] &= 0 \quad \text{and} \quad \mathcal{F}^{\dagger}[\vec{\mu}, u] = 0 \intertext{respectively.
-At some particular guess $\vec{u}^n(\vec{x}, 0)$, we compute the functional derivative}
-  \frac{\delta\L}{\delta \vec{u}^n(\vec{x}, 0)} &= -\vec{\mu}^n(\vec{x}, 0)
-  \intertext{to determine the direction of steepest descent of our objective functional $H[\vec{u}, \vec{U}]$ with respect to the optimization parameter $\vec{u}(\vec{x}, 0)$.
-Naive gradient-descent can then be performed by computing an updated guess}
-  \vec{u}^{n+1}(\vec{x}, 0) &= \vec{u}^n(\vec{x}, 0) + \varepsilon \vec{\mu}^n(\vec{x}, 0)
+  \mathcal{F}[X] &= 0 \quad \text{and} \quad \mathcal{F}^{\dagger}[\chi, X] = 0 \intertext{respectively. At some particular guess $X^n(\vec{x}, 0)$, we compute the functional derivative}
+  \frac{\delta\L}{\delta X^n(\vec{x}, 0)} &= -\chi^n(\vec{x}, 0)
+  \intertext{to determine the direction of steepest descent. Naive gradient-descent can then be performed by computing an updated guess}
+  X^{n+1}(\vec{x}, 0) &= X^n(\vec{x}, 0) + \varepsilon \chi^n(\vec{x}, 0)
   \intertext{where $\varepsilon$ is the step size.}
@@
 In addition to this algorithm, more advanced gradient-based optimization techniques such as conjugate gradient and BFGS have been employed in conjunction with adjoint-looping to improve performance and accelerate convergence.
-In the context of our time-inversion optimization problem, such techniques traverse the space of possible initial conditions $\vec{u}(\vec{x}, 0)$ by relying on local information pertaining to the objective functional $H$ and its functional derivative $\frac{\delta H}{\delta \vec{u}(\vec{x}, 0)}$.
+In the context of our time-inversion optimization problem, such techniques traverse the space of possible initial conditions $X(\vec{x}, 0)$ by relying on local information pertaining to the objective functional $H$ and its functional derivative $\frac{\delta H}{\delta X(\vec{x}, 0)}$.
 For this reason, the focus of our investigation is largely centered around the objective functional's dependence on the initial condition.
 More specifically, we study the levelsets of the objective functional because these surfaces characterize the directions of steepest descent at any arbitrary point in the space of initial conditions.
-In doing so, we highlight the challenges and deficits associated with using the least squares functional to recover the initial condition $\vec{U}(\vec{x}, 0)$ belonging to a target end state $\vec{U}(\vec{x}, T)$.
+In doing so, we highlight the challenges and deficits associated with using the least squares functional (\ref{Hdef}) to recover the initial condition $\bar{X}(\vec{x}, 0)$ belonging to a target end state $\bar{X}(\vec{x}, T)$.
 
@@
 
-\section{Preliminary Analytic Analysis}
+\section{Preliminary Analysis}
 \subsection{Linearity and Reversiblity}
-In this section, we study linear initial value problems which obey (\ref{genpde}) along with the property $\mathcal{F}[u'] = \mathcal{F}[u] - \mathcal{F}[U]$. 
-In these special cases, the objective functional (\ref{Hdef}) depends only on the deviation's initial condition $u'(x, 0)$ about any arbitrary target $U$.
+In this section, we study linear initial value problems which do not include the nonlinear term in (\ref{genpde}), i.e.
+$N[X, X] = N[\bar{X}, \bar{X}] = 0$.
+Therefore $\mathcal{F}[X'] = \mathcal{F}[X] - \mathcal{F}[\bar{X}]$.
+In these special cases, the objective functional (\ref{Hdef}) depends only on the deviation's initial condition $X'(x, 0)$ about any arbitrary target $\bar{X}(x, 0)$.
 For convenience when dealing with linear problems, we take $U(0) = U(T) = 0$ for all $x$.

### 2022-07-16 adjop/kdv/kdv_burgers.py
sines gausss

@@ -158,7 +158,7 @@ for iter in range(int(solver.stop_sim_time // timestep) + 1):
 logger.info('solve complete, sim time = {}'.format(solver.sim_time))
 u.change_scales(1)
 u_T = u['g'].copy()
-np.savetxt(path + '/kdv_U.txt', u_T)
+np.savetxt(path + '/' + write_suffix + '/kdv_U.txt', u_T)
 logger.info('saved final state')
 
 

### 2022-07-16 adjop/kdv/kdv_burgers_sphere.py

@@ -110,9 +110,9 @@ coeffs = 2*np.random.rand(modes_dim) - 1.0
 coeffs *= R / np.sqrt(np.sum(coeffs**2))
 for kx, coeff in enumerate(coeffs):
     if (target == 'gauss'):
-        opt.ic['u']['g'] += coeff*np.cos((kx + 1)*2*np.pi*x / Lx)
+        u['g'] += coeff*np.cos((kx + 1)*2*np.pi*x / Lx)
     else:
-        opt.ic['u']['g'] += coeff*np.sin((kx + 1)*2*np.pi*x / Lx)
+        u['g'] += coeff*np.sin((kx + 1)*2*np.pi*x / Lx)
         
 # Solver
 solver = problem.build_solver(timestepper)

### 2022-07-16 adjop/kdv/kdv_parallel.py

@@ -63,6 +63,8 @@ opt_iters = config.getint('parameters', 'opt_iters')
 method = str(config.get('parameters', 'scipy_method'))
 euler_safety = config.getfloat('parameters', 'euler_safety')
 gamma_init = config.getfloat('parameters', 'gamma_init')
+ri_init = config.getint('parameters', 'ri_init')
+
 R = config.getfloat('parameters', 'R')
 modes_dim = config.getint('parameters', 'modes_dim')
 
@@ -130,7 +132,7 @@ else:
     opt.loop_index = len(opt.descent_tracker['x'])
     load_ind = opt.loop_index
 
-U_data = np.loadtxt(path + '/kdv_U.txt')
+U_data = np.loadtxt(path + '/' + write_suffix + '/kdv_U.txt')
 u = next(field for field in forward_solver.state if field.name == 'u')
 U = dist.Field(name='U', bases=xbasis)
 U['g'] = U_data
@@ -140,15 +142,19 @@ objectiveT = 0.5*(U - u)**2
 opt.set_objectiveT(objectiveT)
 
 mode1 = dist.Field(name='mode1', bases=xbasis)
-mode1['g'] = np.sin(2*np.pi*x / Lx)
-
 mode2 = dist.Field(name='mode2', bases=xbasis)
-mode2['g'] = np.sin(4*np.pi*x / Lx)
+mode3 = dist.Field(name='mode3', bases=xbasis)
 
 
 if (target == 'gauss'):
+    mode1['g'] = np.cos(2*np.pi*x / Lx)
+    mode2['g'] = np.cos(4*np.pi*x / Lx)
+    mode3['g'] = np.cos(6*np.pi*x / Lx)
     opt.ic['u']['g'] = ic_scale*np.exp(-np.power(x - mu, 2.) / (2 * np.power(sig, 2.)))
 elif (target == 'sinecos'):
+    mode1['g'] = np.sin(2*np.pi*x / Lx)
+    mode2['g'] = np.sin(4*np.pi*x / Lx)
+    mode3['g'] = np.sin(6*np.pi*x / Lx)
     opt.ic['u']['g'] = 2*np.sin(2*np.pi*x / Lx) + 2*np.sin(4*np.pi*x / Lx)
 else:
     logger.error('unrecognized target paramter. terminating script')
@@ -167,6 +173,7 @@ for kx, coeff in enumerate(coeffs):
 
[…]

### 2022-07-16 adjop/kdv/main.sh

@@ -1,6 +1,6 @@
 #PBS -S /bin/bash
 #PBS -l select=5:ncpus=40:mpiprocs=40:model=sky_ele
-#PBS -l walltime=8:00:00
+#PBS -l walltime=20:00:00
 #PBS -j oe
 #PBS -W group_list=s2276
 file=${0##*/}
@@ -21,8 +21,9 @@ cd ~/scratch/dedalus/mri/adjop/kdv
 
 FILE="$(readlink -f "$0")"
 DIR="$(dirname "$(readlink -f "$0")")/"
-CONFIG="kdv_sines.cfg"
-PROCS=20
+CONFIG="kdv_options.cfg"
+# CONFIG="kdv_sines.cfg"
+PROCS=200
 
 mpiexec_mpt -np 1      python3 kdv_burgers.py         $CONFIG
 mpiexec_mpt -np $PROCS python3 kdv_burgers_sphere.py  $CONFIG

### 2022-07-16 adjop/kdv/main_sines.sh

@@ -1,0 +1,30 @@
+#PBS -S /bin/bash
+#PBS -l select=1:ncpus=28:mpiprocs=28:model=bro
+#PBS -l walltime=20:00:00
+#PBS -j oe
+#PBS -W group_list=s2276
+file=${0##*/}
+job_name="${file%.*}"
+#PBS -N ${job_name}
+
+export PATH=$HOME/scripts:$PATH
+deactivate
+unset PYTHONPATH
+source ~/miniconda3/etc/profile.d/conda.sh
+export PYTHONNOUSERSITE=1
+conda activate dedalus3
+# support lots of text output to stdio for analysis
+export MPI_UNBUFFERED_STDIO=true
+
+source ~/png2mp4.sh
+cd ~/scratch/dedalus/mri/adjop/kdv
+
+FILE="$(readlink -f "$0")"
+DIR="$(dirname "$(readlink -f "$0")")/"
+CONFIG="kdv_sines.cfg"
+PROCS=28
+
+mpiexec_mpt -np 1      python3 kdv_burgers.py         $CONFIG
+mpiexec_mpt -np $PROCS python3 kdv_burgers_sphere.py  $CONFIG
+mpiexec_mpt -np $PROCS python3 kdv_parallel.py        $CONFIG
+mpiexec_mpt -np 1      python3 paths_pod.py           $CONFIG

### 2022-07-16 adjop/kdv/paths_pod.py

@@ -145,13 +145,22 @@ print(pca.singular_values_)
 
 for comp_ind in range(comps):
     pod0 = pca.components_[comp_ind].reshape(iters, N)
-    pc = plt.pcolormesh(x.ravel(), np.array(range(iters)), pod0, cmap='RdBu_r', shading='gouraud', rasterized=True, clim=(-0.8, 0.8))
-    plt.colorbar(pc)
-    plt.xlim(0, Lx)
-    plt.ylim(0, iters - 1)
-    plt.xlabel(r'$x$')
-    plt.ylabel('Iteration')
-    plt.title('Descent Paths POD: SV = {:.2f}'.format(pca.singular_values_[comp_ind]))
+    f, (a0, a1) = plt.subplots(1, 2, gridspec_kw={'width_ratios': [4, 1]}, sharey=True)
+    pc = a0.pcolormesh(x.ravel(), np.array(range(iters)), pod0, cmap='PRGn', shading='gouraud', rasterized=True, clim=(-0.8, 0.8))
+    L2pod = np.sum(pod0**2, axis=1)**0.5 * 10
+    # plt.colorbar(pc, ax=[a0], orientation = 'horizontal')
+    plt.colorbar(pc, ax=[a0], location='left', pad=0.2)
+    a0.set_xlim(0, Lx)
+    a0.set_ylim(0, iters - 1)
+    a0.set_xlabel(r'$x$')
+    a0.set_ylabel('Iteration')
+    a1.plot(L2pod, np.array(range(iters)), color='black', linewidth=2)
+    a1.set_xlabel(r'$L_2$ norm')
+    a1.get_yaxis().set_visible(False)
+    a1.ticklabel_format(axis='x', style='sci')
+    # print(a1.get_yaxis().__dict__)
+    plt.setp(a1.get_xticklabels(), rotation=0, horizontalalignment='right')
+    plt.suptitle('Descent Paths POD: SV = {:.2f}'.format(pca.singular_values_[comp_ind]))
     plt.savefig(path + '/' + write_suffix + '/pod_comp{}.png'.format(comp_ind))
     logger.info('figure saved to ' + path + '/' + write_suffix + '/pod_comp{}.png'.format(comp_ind))
     plt.close()

### 2022-07-19 adjop/kdv/main.sh
diffusion descent analytical

@@ -1,6 +1,6 @@
 #PBS -S /bin/bash
 #PBS -l select=5:ncpus=40:mpiprocs=40:model=sky_ele
-#PBS -l walltime=24:00:00
+#PBS -l walltime=72:00:00
 #PBS -j oe
 #PBS -W group_list=s2276
 file=${0##*/}

### 2022-07-19 adjop/lsrtm/twomodes.py

@@ -230,39 +230,52 @@ if (not write_objectives or both):
         # print(objectives)
         pc = plt.pcolormesh(k1_coeffs.ravel(), k2_coeffs.ravel(), objectives.T, cmap='GnBu')
         cb = plt.colorbar(pc, fraction=0.046, pad=0.15)
+        plt.clim(0.0, 8e-5)
         plt.scatter([target_coeffs[0]], [target_coeffs[1]], s=60, marker='*', color='k', label = 'target')
-        cont = plt.contour(k1_coeffs.ravel(), k2_coeffs.ravel(), objectives.T, levels=[0.00001, 0.00002, 0.00004], colors=['k', 'k'])
-        plt.clabel(cont, cont.levels, inline=True, fmt=fmt, fontsize=7)
+        cont = plt.contour(k1_coeffs.ravel(), k2_coeffs.ravel(), objectives.T, levels=[0.00001, 0.00002, 0.00004, 0.00008], colors=['k', 'k'])
+        # plt.clabel(cont, cont.levels, inline=True, fmt=fmt, fontsize=7)
         epsilon = 1.0
         if plot_paths:
-            grad_ar = np.gradient(objectives)
-            from scipy import interpolate
-            k1k1, k2k2 = np.meshgrid(k1_coeffs, k2_coeffs)
-            stride = 10
-            gradinterp1 = interpolate.interp2d(k1k1[::stride, ::stride], k2k2[::stride, ::stride], grad_ar[0][::stride, ::stride].T, kind='linear')
-            gradinterp2 = interpolate.interp2d(k1k1[::stride, ::stride], k2k2[::stride, ::stride], grad_ar[1][::stride, ::stride].T, kind='linear')
-            
             sdpath = [(-0.01, -0.01)]
+
+            NN = 10000
+            linspace = np.array(range(NN))
+            tlin = np.exp(-linspace)
+            # tlin = np.linspace(0, 1, 1000000)
+            a1 = np.exp(-2*T*a*(2*np.pi * k1 / Lx)**2)
+            a2 = np.exp(-2*T*a*(2*np.pi * k2 / Lx)**2)
+            x0, y0 = sdpath[0][0], sdpath[0][1]
+            xlin = x0*tlin**(a1)
+            ylin = y0*tlin**(a2)
+            plt.plot(xlin, ylin, color='k', linestyle='--')
+
+            # grad_ar = np.gradient(objectives)
[…]

### 2022-07-20 adjop/lsrtm/lsrtm.tex
minor change
taylor series work: paper

@@
-\documentclass[longbibliography,twocolumn,amsmath,amssymb,aps,nofootinbib]{revtex4-2}
+% \documentclass[longbibliography,twocolumn,amsmath,amssymb,aps,nofootinbib]{revtex4-2}
+\documentclass[longbibliography,amsmath,amssymb,aps,nofootinbib]{revtex4-2}
 
@@
 
+% \vspace{1cm}
+% \textbf{Grad and Div}
+
+% suppose $f(x, y, z):\mathbb{R}^3 \to \mathbb{R}$
+
+% suppose $\vec{F}(x, y, z):\mathbb{R}^3 \to \mathbb{R}^3$
+% \begin{align*}
+%   \textbf{Gradient:}&\quad \grad f(x,y,z) \quad \longrightarrow \quad vector \\
+%   \textbf{Divergence:}&\quad \grad \cdot \vec{F}(x,y,z) \quad \longrightarrow \quad scalar
+% \end{align*}
+% \vspace{1cm}
+
+% \textbf{Navier-Stokes Equation}
+
+% \begin{align*}
+%   \intertext{Newton's 2nd law:}
+%   \vec{F} &= m\vec{a} \\
+%   &= \frac{d}{dt} [m\vec{u}]
+% \end{align*}
+
+% \begin{align*}
+%   \intertext{Consider a velocity $\vec{u}(x, y, t)$}
+%   \textbf{Conservation of momentum:} &\quad \partial_t \vec{u} = - \vec{u} \cdot \grad\vec{u} - \grad p + \nu \grad^2 \vec{u} + \vec{F} \tag{1}\\
+%   \text{Inertial Term:} &\quad \vec{u}\cdot\grad\vec{u} \\
+%   \text{Pressure Gradient Term:} &\quad \grad p \\
+%   \text{Diffusive (Friction) Term:} &\quad \nu \laplacian \vec{u} \\
+%   \text{Body Force:} &\quad \vec{F} \\
+%   &\; \\
+%   \textbf{Conservation of mass:} &\quad \grad \cdot \vec{u} = 0 \tag{2}\\
+%   p(x, y, t) &= f(t) \; [\sum \sum (a_n\sin(nx)b_m\sin(my))]
+% \end{align*}
+
 \clearpage
@@
   u'(x, 0) &= \sum_{k \geq 0} a_k e^{ikx}.
-  \intertext{then}
-  \davg{u'(x, T)} &= \pi \sum_{k \geq 0} \Big[\exp\Big({-2T\sum_{n \geq 1} k^{2n}  }\Big)a_k^2\Big].
+  \intertext{Let's just consider the diffusion equation ($n=1$). Then the end state is given by}
+  u'(x, T) &= \sum_{k\geq 0} a_k \exp(ikx-Tb_1k^2) \label{diffsoln}\\
+  \intertext{The nominal objective functional is given by}
+  \davg{|u'(x, T)|^2} &= \pi \sum_{k \geq 0} \Big[a_k^2\exp\Big({-2T\sum_{n \geq 1} b_n k^{2n}  }\Big)\, \Big] \label{nomobjdiff}.
+  \intertext{Equation (\ref{nomobjdiff}) gives us the elliptical heatmaps we observed previously.}
+  \intertext{Let's take a second-order spatial derivative of (\ref{diffsoln})}
+  \partial_x^2 u'(x, T) &= \pi \sum_{k\geq 0} -k^2 a_k^2 \exp(ikx-Tb_1k^2)\\
+  \intertext{In general, applying a spatial derivative of order $2m$ to (\ref{diffsoln}), yields}
+  \partial_x^{2m} u'(x, T) &= \pi \sum_{k\geq 0} (-k^2)^m a_k^2 \exp(ikx-Tb_1k^{2})\\
+  \intertext{We can leverage this fact to construct an objective functional whose levelsets are spherical. Consider}
 \end{align}
+\begin{tiny}
+\begin{align*}
+  H[u'(T)] &= \davg{|u'(x, T) - (Tb_1)\partial_x^2 u'(x, T) + \frac{(Tb_1)^2}{2!}\partial_x^4 u'(x, T) - \frac{(Tb_1)^3}{3!}\partial_x^6 u'(x, T) ...|^2} \\
+  &= \davg{\Big|\sum_{k\geq 0}a_k \exp(ikx-Tb_1k^2) + \sum_{k\geq 0}(Tb_1k^2)a_k \exp(ikx-Tb_1k^2) + \sum_{k\geq 0}\frac{(Tb_1k^2)^2}{2!}a_k \exp(ikx-Tb_1k^2) + \sum_{k\geq 0}\frac{(Tb_1k^2)^3}{3!}a_k \exp(ikx-Tb_1k^2) + ...\Big|^2} \\
+  &= \davg{|\sum_{k\geq 0}\Big[a_k \exp(ikx-Tb_1k^2)\Big(1 + (Tb_1k^2) + \frac{(Tb_1k^2)^2}{2!} + \frac{(Tb_1k^2)^3}{3!} ...\Big)\Big]|^2} \\
+  &= \davg{|\sum_{k\geq 0}\Big[ a_k \exp(ikx-Tb_1k^2) \Big( \exp(Tb_1k^2) \Big)\Big]|^2} \\
+  &= \davg{|\sum_{k\geq 0}\Big[ a_k \exp(ikx)  \Big)\Big]|^2} \\
+  &= \pi \sum_{k\geq 0} a_k^2
+  % &= \pi \sum_{k \geq 0} \Big[\exp\Big({-2T\sum_{n \geq 1} b_n k^{2n}  }\Big)\, a_k^2\Big].
+\end{align*}
+\end{tiny}
+That's gives us spherical objective levelsets. Notice what happens as we increment the number of terms in the above Taylor Series
+
+% \begin{figure*}
+%   \centering
+%   \begin{tabular}{@{}c@{}}
+%       \includegraphics[width=3.375in]{flux_sup_n.pdf}
+%   \end{tabular}
+%   \begin{tabular}{@{}c@{}}
+%       \includegraphics[width=3.375in]{EV_spectra_2ra.pdf}
+%   \end{tabular}
+%   \caption{Heat fluxes (left) and eigenvalue spectra (right) of equilibrated states with $\Ra = 2 \times 10^5$ (top) and $\Ra = 10^9$ (bottom). 
+%   The heat flux profiles are symmetric about $z=0$, so we only plot for $z<0$.
+%   We plot the advection profiles for marginally-stable modes. 
+%   At low $\Ra$, a single mode with $k_x = 1.5\pi$ is sufficient to oppose boundary layer diffusion and facilitate heat flux throughout the bulk of the domain. 
+%   At large $\Ra$, high-wavenumber modes contribute pronounced small-scale advection profiles which tightly hug the thin boundary layers. 
+%   A combination of progressively wider advection profiles is necessary to transition to bulk advection provided by the $k_x = 1.5\pi$ mode. 
+%   This bulk mode mimics the large-scale convective cells which are characteristic of Rayleigh–Bénard convection, both in qualitative structure and number of cells in the domain.
+%   The eigenfunctions which compose the advective profiles are illustrated in Appendix~\ref{sec:eigenfunctions}.}
+%   \label{fig:flux}
+% \end{figure*}
+
+\clearpage
 When the initial deviation $u'(x, 0)$ is described using a basis of fourier coefficients, its levelsets form hyperellipses, whose semi-major (semi-minor) axes are aligned with $u'(x, 0)$'s largest (smallest) fourier coefficient.
@@
 
-\begin{figure}
-  \includegraphics[width=9cm]{objtest_a0p03b0p0c0p0T3p0R0p015kt10p0kt20.png}
-  \caption{Objective function and an example gradient descent path for the diffusion equation.
+% \begin{figure}
+%   \includegraphics[width=9cm]{objtest_a0p03b0p0c0p0T3p0R0p015kt10p0kt20.png}
+%   \caption{Objective function and an example gradient descent path for the diffusion equation.
 The target simulation $U(t) = 0$ and the deviation is composed of two fourier modes, i.e.
@@
 Because the system is linear self-adjoint, each subsequent guess in the gradient descent trajectory is constrained to lie in this 2D phase space.}
-\end{figure}
+% \end{figure}
 

### 2022-07-20 adjop/lsrtm/lsrtm.tex
taylor series paper done

@@
 
-% \begin{figure*}
-%   \centering
-%   \begin{tabular}{@{}c@{}}
-%       \includegraphics[width=3.375in]{flux_sup_n.pdf}
-%   \end{tabular}
-%   \begin{tabular}{@{}c@{}}
-%       \includegraphics[width=3.375in]{EV_spectra_2ra.pdf}
-%   \end{tabular}
+\begin{figure*}[h]
+  \centering
+  \begin{tabular}{@{}c@{}}
+      \includegraphics[width=3in]{SPHRtest_Nts1a0p3b0p0c0p0T1p0R0p015kt10p0kt20.png}
+  \end{tabular}
+  \begin{tabular}{@{}c@{}}
+      \includegraphics[width=3in]{SPHRtest_Nts2a0p3b0p0c0p0T1p0R0p015kt10p0kt20.png}
+  \end{tabular}
+  \begin{tabular}{@{}c@{}}
+      \includegraphics[width=3in]{SPHRtest_Nts3a0p3b0p0c0p0T1p0R0p015kt10p0kt20.png}
+  \end{tabular}
+  \begin{tabular}{@{}c@{}}
+      \includegraphics[width=3in]{SPHRtest_Nts4a0p3b0p0c0p0T1p0R0p015kt10p0kt20.png}
+  \end{tabular}
   % \caption{Heat fluxes (left) and eigenvalue spectra (right) of equilibrated states with $\Ra = 2 \times 10^5$ (top) and $\Ra = 10^9$ (bottom). 
@@
   % \label{fig:flux}
-% \end{figure*}
+\end{figure*}
 

### 2022-07-20 adjop/lsrtm/lsrtm.tex
taylor series paper donedone

@@
   u'(x, 0) &= \sum_{k \geq 0} a_k e^{ikx}.
+\end{align}
+\clearpage
+\begin{align}
   \intertext{Let's just consider the diffusion equation ($n=1$). Then the end state is given by}
@@
   \davg{|u'(x, T)|^2} &= \pi \sum_{k \geq 0} \Big[a_k^2\exp\Big({-2T\sum_{n \geq 1} b_n k^{2n}  }\Big)\, \Big] \label{nomobjdiff}.
-  \intertext{Equation (\ref{nomobjdiff}) gives us the elliptical heatmaps we observed previously.}
+  \intertext{Equation (\ref{nomobjdiff}) gives us the elliptical heatmaps we observed previously, as shown in Figure \ref{diff_ellipse}.}
+\end{align}
+
+\begin{figure}[h]
+  \includegraphics[width=9cm]{objtest_a0p03b0p0c0p0T3p0R0p015kt10p0kt20.png}
+  \caption{Objective function and an example gradient descent path for the diffusion equation.
+The target simulation $U(t) = 0$ and the deviation is composed of two fourier modes, i.e.
+$u'(t) = a_1\sin x + a_2\sin 2x$.
+Because the system is linear self-adjoint, each subsequent guess in the gradient descent trajectory is constrained to lie in this 2D phase space.}
+  \label{diff_ellipse}
+\end{figure}
+
+\begin{align}
   \intertext{Let's take a second-order spatial derivative of (\ref{diffsoln})}
-  \partial_x^2 u'(x, T) &= \pi \sum_{k\geq 0} -k^2 a_k^2 \exp(ikx-Tb_1k^2)\\
+  \partial_x^2 u'(x, T) &= \sum_{k\geq 0} -k^2 a_k^2 \exp(ikx-Tb_1k^2)\\
   \intertext{In general, applying a spatial derivative of order $2m$ to (\ref{diffsoln}), yields}
-  \partial_x^{2m} u'(x, T) &= \pi \sum_{k\geq 0} (-k^2)^m a_k^2 \exp(ikx-Tb_1k^{2})\\
+  \partial_x^{2m} u'(x, T) &= \sum_{k\geq 0} (-k^2)^m a_k^2 \exp(ikx-Tb_1k^{2})\\
   \intertext{We can leverage this fact to construct an objective functional whose levelsets are spherical. Consider}
@@
 \end{tiny}
-That's gives us spherical objective levelsets. Notice what happens as we increment the number of terms in the above Taylor Series
+That gives us spherical objective levelsets. Notice what happens as we increment the number of terms in the above Taylor Series
 
@@
 Accordingly, the path of steepest descent is only aligned with $u'(x, 0)$ when $u'(x, 0)$ is composed of a single fourier mode. Such systems are self-adjoint, meaning $L[u] = L^{\dagger}[u]$
-
-% \begin{figure}
-%   \includegraphics[width=9cm]{objtest_a0p03b0p0c0p0T3p0R0p015kt10p0kt20.png}
-%   \caption{Objective function and an example gradient descent path for the diffusion equation.
-The target simulation $U(t) = 0$ and the deviation is composed of two fourier modes, i.e.
-$u'(t) = a_1\sin x + a_5\sin 5x$.
-Because the system is linear self-adjoint, each subsequent guess in the gradient descent trajectory is constrained to lie in this 2D phase space.}
-% \end{figure}
 

### 2022-07-20 adjop/lsrtm/twomodes.py
taylor series limits plotted

@@ -71,7 +71,7 @@ kt1_str = str(target_coeffs[0]).replace('.', 'p')
 kt2_str = str(target_coeffs[1]).replace('.', 'p')
 R_str = str(R).replace('.', 'p')
 
-objectives_str = path + '/objectives_a' + a_str + 'b' + b_str + 'c' + c_str + 'T' + T_str + 'R' + R_str + 'kt1' + kt1_str + 'kt2' + kt2_str + '.txt'
+objectives_str = path + '/objectives_Nts' + str(Nts) + 'a' + a_str + 'b' + b_str + 'c' + c_str + 'T' + T_str + 'R' + R_str + 'kt1' + kt1_str + 'kt2' + kt2_str + '.txt'
 save_dir = path + '/SPHRtest_Nts' + str(Nts) + 'a' + a_str + 'b' + b_str + 'c' + c_str + 'T' + T_str + 'R' + R_str + 'kt1' + kt1_str + 'kt2' + kt2_str + '.png'
 
 # Simulation Parameters
@@ -257,12 +257,17 @@ if (not write_objectives or both):
             # tlin = np.linspace(0, 1, 1000000)
             a1 = np.exp(-2*T*a*(2*np.pi * k1 / Lx)**2)
             a2 = np.exp(-2*T*a*(2*np.pi * k2 / Lx)**2)
+            if Nts > 1:
+                for nt in range(1, Nts):
+                    a1 += (T*a*k1**2)**nt*np.exp(-2*T*a*(2*np.pi * k1 / Lx)**2)
+                    a2 += (T*a*k2**2)**nt*np.exp(-2*T*a*(2*np.pi * k2 / Lx)**2)
+
             x0, y0 = sdpath[0][0], sdpath[0][1]
             xlin = x0*tlin**(a1)
             ylin = y0*tlin**(a2)
-            
-            # plt.plot(xlin, ylin, color='k', linestyle='--')
-            # plt.scatter([sdpath[-1][0]], [sdpath[-1][1]], s=60, marker='o', color='k', label = 'guess')
+
+            plt.plot(xlin, ylin, color='k', linestyle='--')
+            plt.scatter([sdpath[-1][0]], [sdpath[-1][1]], s=60, marker='o', color='k', label = 'guess')
 
             # grad_ar = np.gradient(objectives)
             # from scipy import interpolate

### 2022-07-21 adjop/lsrtm/lsrtm.tex
not sure
undiffuse section

@@
 \begin{align}
-  \intertext{Let's just consider the diffusion equation ($n=1$). Then the end state is given by}
+  \intertext{Let's just consider the diffusion equation ($n=1$) with diffusivity $b_1$. The end state is given by}
   u'(x, T) &= \sum_{k\geq 0} a_k \exp(ikx-Tb_1k^2) \label{diffsoln}\\
-  \intertext{The nominal objective functional is given by}
+  \intertext{and the nominal objective functional is given by}
   \davg{|u'(x, T)|^2} &= \pi \sum_{k \geq 0} \Big[a_k^2\exp\Big({-2T\sum_{n \geq 1} b_n k^{2n}  }\Big)\, \Big] \label{nomobjdiff}.
@@
   \partial_x^{2m} u'(x, T) &= \sum_{k\geq 0} (-k^2)^m a_k^2 \exp(ikx-Tb_1k^{2})\\
-  \intertext{We can leverage this fact to construct an objective functional whose levelsets are spherical. Consider}
+  \intertext{We can leverage this fact to construct an objective functional whose levelsets are spherical.} 
+  \intertext{Consider the following objective functional}
 \end{align}
+\vspace{-1cm}
 \begin{tiny}
@@
   % \label{fig:flux}
+\end{figure*}
+
+\begin{figure*}[h]
+  \centering
+  \includegraphics[width=4.5in]{SPHRtest_Nts5a0p3b0p0c0p0T1p0R0p015kt10p0kt20.png}
 \end{figure*}

### 2022-07-21 adjop/Undiffuse.py
pod 4001

@@ -1,0 +1,58 @@
+from dedalus.core import SpectralOperator1D
+from dedalus.tools.dispatch import MultiClass
+
+class Undiffuse(SpectralOperator1D, metaclass=MultiClass):
+    """
+    Differentiation along one dimension.
+
+    Parameters
+    ----------
+    operand : number or Operand object
+    space : Space object
+
+    """
+
+    name = "Diff"
+
+    def __init__(self, operand, coord, out=None):
+        super().__init__(operand, out=out)
+        # SpectralOperator requirements
+        self.coord = coord
+        self.input_basis = operand.domain.get_basis(coord)
+        self.output_basis = self._output_basis(self.input_basis)
+        self.first_axis = coord.axis
+        self.last_axis = coord.axis
+        self.axis = coord.axis
+        # LinearOperator requirements
+        self.operand = operand
+        # FutureField requirements
+        self.domain = operand.domain.substitute_basis(self.input_basis, self.output_basis)
+        self.tensorsig = operand.tensorsig
+        self.dtype = operand.dtype
+
+    @classmethod
+    def _check_args(cls, operand, coord, out=None):
+        # Dispatch by operand basis
+        if isinstance(operand, Operand):
+            basis = operand.domain.get_basis(coord)
+            if isinstance(basis, cls.input_basis_type):
+                return True
+        return False
+
+    def new_operand(self, operand, **kw):
+        return Differentiate(operand, self.coord, **kw)
+
+    @staticmethod
+    def _output_basis(input_basis):
+        # Subclasses must implement
+        raise NotImplementedError()
+
+    def __str__(self):
+        return 'd{!s}({!s})'.format(self.coord.name, self.operand)
+
+    def _expand_multiply(self, operand, vars):
+        """Expand over multiplication."""
+        args = operand.args
+        # Apply product rule to factors
+        partial_diff = lambda i: prod([self.new_operand(arg) if i==j else arg for j,arg in enumerate(args)])
+        return sum((partial_diff(i) for i in range(len(args))))

### 2022-07-21 adjop/test_undiffuse.py

@@ -1,0 +1,1 @@
+import Undiffuse
